---
title: "EOL_LF"
author: "samir"
date: "2023-09-04"
output: html_document
---

```{r libraries, echo=FALSE, warning=FALSE, message=FALSE}

library("readr")
library("viridis")
library("scales")
library("lubridate")
library("gsignal")
library("FactoMineR")
library("factoextra")
library("gplots")
library("heatwaveR")
library("dplyr")
library("imputeTS")
library("Metrics")
library("tidyr")
library("tseries")
library("forecast")
library("tsDyn")
library("tsoutliers")
library("seacarb")
library("lmodel2")
library("lmtest")
library("ggExtra")
library("corrplot")
library("ggcorrplot")
library("marelac")
library("spectral")
library("signal")
library("cowplot")
library("kableExtra")
library("EnvStats")
library(xts)
library(dygraphs)
```

<!-- # Point B : surface (2007-2022)  -->

<!-- Utilisation of : na_seadec() function   -->
<!-- = Removes the seasonal component from the time series, performs imputation on the deseasonalized series and afterwards adds the seasonal component again.   -->


<!-- ## Temperature & salinity -->

<!-- ```{r interpolation temp & salinity, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7,fig.height=5} -->

<!-- SOMLIT_pcO2 <- read_delim("./Data/Data_PointB/DATA_SOMLIT_07-22_impute_0m.csv",  -->
<!--     delim = ";", escape_double = FALSE, trim_ws = TRUE) %>%  -->
<!--   dplyr::filter(datetime <= "2022-12-31") -->

<!-- impute_temp <- na_seadec(SOMLIT_pcO2$temperature, algorithm = "interpolation", find_frequency = F) -->
<!-- impute_sal <- na_seadec(SOMLIT_pcO2$salinity, algorithm = "interpolation", find_frequency = F) -->

<!-- # plot  -->
<!-- ggplot_na_imputations(x_with_na = SOMLIT_pcO2$temperature, -->
<!--                           x_with_imputations = impute_temp, -->
<!--                           color_imputations = "gold") -->


<!-- SOMLIT_pcO2$temperature[is.na(SOMLIT_pcO2$pCO2_atmos)] <- "NA" -->

<!-- SOMLIT_pcO2 %>%  -->
<!--   dplyr::filter(temperature == "NA") %>%  -->
<!--   dplyr::group_by(Year) %>%  -->
<!--   dplyr::summarise(number_of_NA = n()) %>%  -->
<!--   ggplot(aes(x = Year, y = number_of_NA))  + -->
<!--   geom_segment(aes(x = Year, xend = Year, y = 0, yend = number_of_NA),  -->
<!--                color = "grey", linewidth = 2) + -->
<!--   geom_point(size = 5, fill = "#689D71", shape = 21, color = "black") + -->
<!--   coord_flip() + -->
<!--   scale_y_continuous(name = "Nombre de données manquantes T", limits = c(0,15), breaks = seq(0, 15, 1)) +  -->
<!--   scale_x_discrete(name="", limits=c("2007":"2021")) -->

<!-- SOMLIT_pcO2 <- SOMLIT_pcO2 %>%  -->
<!--   mutate(pCO2_atmos = as.numeric(pCO2_atmos)) -->


<!-- # test filling date with complete() -->
<!-- test <- SOMLIT_pcO2 %>% complete(datetime = seq.Date(min(datetime), max(datetime), by="week")) -->
<!-- test2 <- SOMLIT_pcO2 %>% complete(datetime ) -->

<!-- #ggplot_na_gapsize(SOMLIT_pcO2$pCO2_atmos) -->


<!-- # ggplot_na_imputations(SOMLIT_pcO2$pCO2_atmos, SOMLIT_pcO2$impute_pCO2_atmos, x_axis_labels = SOMLIT_pcO2$datetime, xlab = "",  -->
<!-- #                       ylab=expression(paste(italic(p), CO[2], " atmosphérique ", "(atm)")),  -->
<!-- #                       title="", color_lines = "lightslategray", -->
<!-- #                       color_points = "steelblue", size_points = 1.2, size_imputations = 1.7) + -->
<!-- #   scale_x_date(date_breaks = "2 years", date_labels = "%Y")  -->
<!-- #    -->



<!-- ``` -->

# Data

```{r importation data DATA DIC + Oxy , echo=FALSE, message=FALSE, warning=FALSE}
#period : jan. 2007 - dec. 2022
datadic <- read_delim("./Data/data_PointB/DATA_SOMLIT_07-22_impute_0m_deltas_pCO2_and_fluxes_CO2.csv", 
    delim = ";", escape_double = FALSE, trim_ws = TRUE)
datadic <- datadic %>%
  rename(pH_spec_insitu=pH_insi, pH_spec_25 = pH_s_somlit_temp_25) %>% 
  dplyr::select(-c("Year", "Month", "Year.Month"))
datadic$datetime <- dmy(datadic$datetime)

datadic2007_2009 <- read_delim("./Data/data_PointB/SOMLIT_MAJ_Samir.csv", 
    delim = ",", escape_double = FALSE, trim_ws = TRUE)
datadic2007_2009 <- datadic2007_2009 %>%
  rename(datetime=sampling_date) %>% 
  dplyr::filter(depth == 1 & datetime < "2009-01-01") %>% 
           dplyr::select(-c("depth", "CSC_flag"))
datadic2007_2009$datetime <- ymd(datadic2007_2009$datetime)
# bind
datadic <- bind_rows(datadic, datadic2007_2009)
datadic <- dplyr::arrange(datadic,datetime)


# Add Somlit data OXY...
oxy <- read_delim("./Data/data_PointB/DATA_SOMLIT_02-22_oxy_CLEAN_impute_0m_AOU.csv", 
    delim = ";", escape_double = FALSE, trim_ws = TRUE)
oxy$datetime <- as.Date(oxy$datetime)

datadic <- left_join(datadic, oxy %>% dplyr::select(-c("PROF_NUM")), by="datetime")

 datadic <- datadic %>% 
  complete(datetime = seq.Date(min(datetime), max(datetime), by="week"))

# Add NA row with missing weeks (missing tuesdays). After we have to filter because it added row on a week where data was already there but on mondays or wednesday...
 datadic <-  datadic %>% 
   arrange(datetime) %>%
  mutate(diff_decale = datetime - lag(datetime, default = first(datetime)),
         diff_day=as.numeric(lead(datetime,default = last(datetime))-datetime)) %>% 
  select(diff_day, diff_decale, everything())

 datadic<- datadic[!( datadic$diff_day=="1" &  datadic$diff_decale=="7"),]
 datadic<- datadic[!( datadic$diff_day=="7" &  datadic$diff_decale=="1"),]
 datadic<- datadic[!( datadic$diff_day=="2" &  datadic$diff_decale=="7"),]
 datadic<- datadic[!( datadic$diff_day=="3" &  datadic$diff_decale=="7"),]

 datadic<- datadic %>% 
  dplyr::filter((datetime!="2011-11-08" &  datetime!="2011-11-01"& datetime!="2009-11-03"& datetime!="2020-09-08"& datetime!="2007-05-08"& datetime!="2012-05-01"& datetime!="2013-01-01"& datetime!="2013-11-26"& datetime!="2014-11-04"& datetime!="2014-11-11"& datetime!="2016-11-01"& datetime!="2018-01-09"& datetime!="2018-01-16"& datetime!="2018-03-13"& datetime!="2018-05-08"& datetime!="2020-02-11"& datetime!="2020-09-08" & datetime!="2008-09-23" & datetime!= "2009-01-20"& datetime!= "2012-05-08")) %>% 
  dplyr::distinct(datetime, .keep_all = TRUE) %>% 
   rename(pH_spec_insi= pH_s_somlit_temp_insi)

 # fill with outpu_somlit because pH is not full
 output_somlit <- read_delim("../PointB-git/output_jp/data/output_somlit.csv", 
    delim = ",", escape_double = FALSE, trim_ws = TRUE)
  output_somlit <- output_somlit %>% 
    dplyr::filter(depth ==0) %>% 
    dplyr::rename(datetime=sampling_date)
  
  # remove pH_spec_25 and pH_spec_insitu from SOMLIT because not full. MErge them from output_somlit.
  datadic <- datadic %>% 
    select(-c(pH_spec_25, pH_spec_insi, pH_insi))
 datadic <- left_join(datadic, output_somlit %>% dplyr::select(datetime, pH_spec_25), by="datetime")
 
SOMLIT<- datadic

#Plot
 diff_xts <- dplyr::select(SOMLIT, datetime, pH_spec_25,pH_spec_insitu,pH_18,pH_calc)
diff_xts <- as.xts(diff_xts, order.by=diff_xts$datetime)
dygraph(diff_xts, main = "", ylab = "") %>%
  dySeries("pH_spec_25", drawPoints = TRUE, color = "blue" ,pointSize=2, label = "pH_spec_25")%>%
  dySeries("pH_spec_insitu", drawPoints = TRUE, color = "cyan", pointSize=2, label = "pH_spec_insitu")%>%
    dySeries("pH_18", drawPoints = TRUE, color = "darkblue", pointSize=2, label = "pH_18")%>%
  dySeries("pH_calc", drawPoints = TRUE, color = "orange", pointSize=2, label = "pH_calc")%>%
dyOptions(drawGrid = TRUE,
    drawPoints = TRUE,
    useDataTimezone = TRUE)%>%
  dyRangeSelector(height = 30)

 diff_xts <- dplyr::select(SOMLIT, datetime, NpCO2,TpCO2,pCO2,pCO2_w)
diff_xts <- as.xts(diff_xts, order.by=diff_xts$datetime)
dygraph(diff_xts, main = "", ylab = "") %>%
  dySeries("pCO2", drawPoints = TRUE, color = "blue" ,pointSize=2, label = "pCO2")%>%
  dySeries("pCO2_w", drawPoints = TRUE, color = "cyan", pointSize=2, label = "pCO2_w")%>%
    dySeries("NpCO2", drawPoints = TRUE, color = "darkblue", pointSize=2, label = "NpCO2")%>%
  dySeries("TpCO2", drawPoints = TRUE, color = "orange", pointSize=2, label = "TpCO2")%>%
dyOptions(drawGrid = TRUE,
    drawPoints = TRUE,
    useDataTimezone = TRUE)%>%
  dyRangeSelector(height = 30)

 diff_xts <- dplyr::select(SOMLIT, datetime, ta,dic)
diff_xts <- as.xts(diff_xts, order.by=diff_xts$datetime)
dygraph(diff_xts, main = "", ylab = "") %>%
  dySeries("ta", drawPoints = TRUE, color = "blue" ,pointSize=2, label = "ta")%>%
  dySeries("dic", drawPoints = TRUE, color = "red", pointSize=2, label = "dic")%>%
dyOptions(drawGrid = TRUE,
    drawPoints = TRUE,
    useDataTimezone = TRUE)%>%
  dyRangeSelector(height = 30)
```


# Anomalies method

## By months
  
```{r calcul monthly means, echo=FALSE, warning=FALSE, message=FALSE}
monthly_means <- ungroup(SOMLIT) %>% 
  #mutate(Month = format(datetime, format="%m")) %>% 
  group_by(month) %>%
  summarise(
    sal_month = mean(salinity, na.rm = TRUE),
    sd_sal = sd(salinity, na.rm = TRUE),
    temp_month = mean(temperature, na.rm = TRUE),
    sd_temp = sd(temperature, na.rm = TRUE),
    oxy_month = mean(oxy_umol_kg, na.rm = TRUE),
    sd_oxy = sd(oxy_umol_kg, na.rm= TRUE),
    aou_month = mean(AOU, na.rm = TRUE),
    sd_aou = sd(AOU, na.rm= TRUE),
    ta_month = mean(ta, na.rm = TRUE),
    sd_ta = sd(ta, na.rm= TRUE),
    dic_month = mean(dic, na.rm = TRUE),
    sd_dic = sd(dic, na.rm= TRUE),
    phinsi_month = mean(pH_spec_insitu, na.rm = TRUE),
    sd_phinsi = sd(pH_spec_insitu, na.rm= TRUE),
     phspec25_month = mean(pH_spec_25, na.rm = TRUE),
    sd_phspec25 = sd(pH_spec_25, na.rm= TRUE), 
    npco2_month = mean(NpCO2, na.rm = TRUE),
    sd_npco2 = sd(NpCO2, na.rm= TRUE),
    tpco2_month = mean(TpCO2, na.rm = TRUE),
    sd_tpco2 = sd(TpCO2, na.rm= TRUE),
    pco2w_month = mean(pCO2_w, na.rm = TRUE),
    sd_pco2w = sd(pCO2_w, na.rm= TRUE))
```

```{r regression analysis SOMLIT, echo=FALSE, warning=FALSE, message=FALSE}

anomalies_SOM_month <- left_join(ungroup(SOMLIT), monthly_means, by = "month") %>%
  mutate(sal_ano = salinity - sal_month,
         temp_ano = temperature - temp_month,
         oxy_ano = oxy_umol_kg - oxy_month,
         aou_ano = AOU - aou_month,
          ta_ano = ta - ta_month,
          dic_ano = dic - dic_month,
          phinsi_ano = pH_spec_insitu - phinsi_month,
          phspec25_ano = pH_spec_25 - phspec25_month,
          npco2_ano = NpCO2 - npco2_month,
          tpco2_ano = TpCO2 - tpco2_month,
         pco2w_ano = pCO2_w - pco2w_month)

# Regression and table anomalies
var_list <- c("temp_ano", "sal_ano","oxy_ano", "aou_ano", "ta_ano", "dic_ano", "phinsi_ano","phspec25_ano","npco2_ano", "tpco2_ano","pco2w_ano")

lms <- lapply(var_list, function(x) {
  summary(lm(substitute(i ~ decimal_date(datetime), list(i = as.name(x))), 
             data = anomalies_SOM_month))
})
reg_raw <- NULL
for (i in 1:length(var_list)) {
  #one loops through all anomalies
  # calculate probability of fstatistic because it cannot be extracted from lms above
  # see http://stats.stackexchange.com/questions/92824/extracting-the-model-p-value-for-a-multiple-regression-in-r
  # slope returned by lm is per second
  prob <-
    pf(lms[[i]]$fstatistic[1],
       lms[[i]]$fstatistic[2],
       lms[[i]]$fstatistic[3],
       lower.tail = FALSE)
  reg_raw <-
    rbind(reg_raw, as.numeric(
      c(
        lms[[i]]$coefficients[2, 1],
        lms[[i]]$coefficients[2, 2],
        lms[[i]]$coefficients[2, 4],
        lms[[i]]$coefficients[1, 1],
        lms[[i]]$coefficients[1, 2],
        lms[[i]]$coefficients[1, 4],
        lms[[i]]$fstatistic[1],
        lms[[i]]$fstatistic[3],
        lms[[i]]$r.squared,
        prob
      )
    ))
}
colnames(reg_raw) <-
  c("Slope",
    "SE Slope",
    "P Slope",
    "Intercept",
    "SE int.",
    "P int.",
    "F",
    "df",
    "R2",
    "P value")

row.names(reg_raw) <- var_list
reg_anomalies_raw <- reg_raw
kable(reg_anomalies_raw, caption="Time series anomaly regression analyses per month", format = "pipe", align='l')

##RAPPORT##
# reg_anomalies_raw_rapport <- as.data.frame(reg_anomalies_raw)
# reg_anomalies_raw_rapport <- reg_anomalies_raw_rapport %>% 
#   mutate(Variables = c("T (°C)", "S"))
# 
# reg_anomalies_raw_rapport <- reg_anomalies_raw_rapport %>% 
#   mutate( 
#   `Pente ± sd` = paste(round(reg_anomalies_raw_rapport$Slope, digits=3), "±", round(reg_anomalies_raw_rapport$`SE Slope`, digits=4)), 
#   `P pente` = reg_anomalies_raw_rapport$`P Slope`, 
#   `Int. ± sd` = paste(round(reg_anomalies_raw_rapport$Intercept, digits=2), "±", round(reg_anomalies_raw_rapport$`SE int.`, digits=2)),
#   `P Int.` = reg_anomalies_raw_rapport$`P int.`, 
#   `R^2^` = round(reg_anomalies_raw_rapport$R2, digits=4))
# 
# 
# reg_anomalies_raw_rapport <- reg_anomalies_raw_rapport[,c(11:16)]
# 
# kable(reg_anomalies_raw_rapport, caption="Time series anomaly regression analyses (EOL)", format = "pipe", align='l')

```

### Temperature

```{r SOM month - plot anomalies + trends - temperature, echo=FALSE, warning=FALSE, message=FALSE}
slope <- reg_anomalies_raw[1,1]
std <- reg_anomalies_raw[1,2]
pv <-reg_anomalies_raw[1,10]

plot_ano_temp_SOM_month <- ggplot(data = anomalies_SOM_month, aes(x = datetime, y = temp_ano), na.rm=TRUE) +
  #scale_x_datetime(date_breaks="2 year", date_minor_breaks="1 years", labels = date_format("%Y")) +
  geom_point(colour="blue", na.rm=TRUE, size=0.65) +
  geom_smooth(method=lm, colour="black", fill="grey", linewidth=0.6, na.rm=TRUE) +
  labs(title="SOMLIT - Temperature anomalies + trend",x="", y="Temperature (°C)")+
  annotate(geom="text", x=as.POSIXct("2019-12-03 01:00:00"), y=5.5, label=paste0( "slope: ",  slope," ± ", std, ", P-value: ", pv), color="black")

plot_ano_temp_SOM_month
```


Other parameters to come.






# Other methods:

**By autocorrelation + decompose function or Mann-Kendall : not possible because of gaps**

--> Filling gaps.



## Moving average

### Temperature

```{r moving average SOMLIT T, echo=FALSE, warning=FALSE, message=FALSE}

# data_somlit_ts <- read_delim("./Data/data_PointB/DATA_SOMLIT_07-22_impute_0m.csv",
#     delim = ";", escape_double = FALSE, trim_ws = TRUE) %>%
#   dplyr::filter(datetime >= "2014-01-07") %>%
#   mutate(Month = format(datetime, format="%m"))

#Transfome data into TS, frequency : 365.25/7 = 52.17857
#ts_temp_som <- ts(SOMLIT$temp_SOM, start = c(2018,1), end = c(2022,52), freq = 52)
ts_temp_som <- ts(SOMLIT$temperature, freq = 52)

# look at the NA in the ts before filling
ggplot_na_distribution(ts_temp_som )

# filling datat(imputation) with na_seadec() = Removes the seasonal component from the time series, performs imputation on the deseasonalized series and afterwards adds the seasonal component again.
ts_seadec <-  na_seadec(ts_temp_som, algorithm = "ma",find_frequency = TRUE)  
# plot 
ggplot_na_imputations(x_with_na = ts_temp_som,
                          x_with_imputations = ts_seadec,
                          color_imputations = "gold")
 
 
#decompose on filled data : ts_seadec
ts_temp_som_decomp <- decompose(ts_seadec, type = "additive")

autoplot(ts_temp_som_decomp) +
  xlab(' ')

#remove seasonality :
ts_temp_som_less_season <- ts_temp_som_decomp$x - ts_temp_som_decomp$seasonal

plot_ts_temp_som_less_season <- autoplot(ts_temp_som_less_season) +
  xlab("") + ylab("Temp. (°C)") +
  ggtitle("Temperature time series at Pt B (without seasonality)") +
  geom_smooth(method=lm)

#regression values :
reg_val_ts_temp_som_less_season <- as.data.frame(plot_ts_temp_som_less_season$data)

model_ts_temp_less_season <- lm(formula=y~x, data=reg_val_ts_temp_som_less_season)
summary(model_ts_temp_less_season)


#plot with slope value :
plot_ts_temp_som_less_season <- autoplot(ts_temp_som_less_season) +
  xlab("") + ylab("Temp. (°C)") +
  ggtitle("Temperature time series at Pt B (without seasonality)") +
  geom_smooth(method=lm) +
  annotate(geom="text", x=5, y=24, label="slope :  0.07 ± 0.016 (P = 2.99e-06, significative)", color="black")

plot_ts_temp_som_less_season
```

Slope is `r round(model_ts_temp_less_season$coefficients[[2]],3)`.

### Salinity

```{r moving average SOMLIT S, echo=FALSE, warning=FALSE, message=FALSE}

# data_somlit_ts <- read_delim("./Data/data_PointB/DATA_SOMLIT_07-22_impute_0m.csv",
#     delim = ";", escape_double = FALSE, trim_ws = TRUE) %>%
#   dplyr::filter(datetime >= "2014-01-07") %>%
#   mutate(Month = format(datetime, format="%m"))

#Transfome data into TS, frequency : 365.25/7 = 52.17857
ts_sal_som <- ts(SOMLIT$salinity, freq = 52)

# look at the NA in the ts before filling
ggplot_na_distribution(ts_sal_som )

# filling datat(imputation) with na_seadec() = Removes the seasonal component from the time series, performs imputation on the deseasonalized series and afterwards adds the seasonal component again.
ts_seadec <-  na_seadec(ts_sal_som, algorithm = "ma",find_frequency = TRUE)  
# plot 
ggplot_na_imputations(x_with_na = ts_sal_som,
                          x_with_imputations = ts_seadec,
                          color_imputations = "gold")
 
 
#decompose on filled data : ts_seadec
ts_sal_som_decomp <- decompose(ts_seadec, type = "additive")

autoplot(ts_sal_som_decomp) +
  xlab(' ')

#remove seasonality :
ts_sal_som_less_season <- ts_sal_som_decomp$x - ts_sal_som_decomp$seasonal

plot_ts_sal_som_less_season <- autoplot(ts_sal_som_less_season) +
  xlab("") + ylab("Salinity") +
  ggtitle("Salinity time series at Pt B (without seasonality)") +
  geom_smooth(method=lm)

#regression values :
reg_val_ts_sal_som_less_season <- as.data.frame(plot_ts_sal_som_less_season$data)

model_ts_sal_less_season <- lm(formula=y~x, data=reg_val_ts_sal_som_less_season)
summary(model_ts_sal_less_season)


#plot with slope value :
plot_ts_sal_som_less_season <- autoplot(ts_sal_som_less_season) +
  xlab("") + ylab("Salinity") +
  ggtitle("Salinity time series at Pt B (without seasonality)") +
  geom_smooth(method=lm) +
  annotate(geom="text", x=10, y=36, label="slope :  -0.0001 ± 0.002 (P = 0.9, non-significative)", color="black")

plot_ts_sal_som_less_season
```

Slope is `r round(model_ts_sal_less_season$coefficients[[2]],3)`.

### Oxygen

```{r moving average SOMLIT oxy, echo=FALSE, warning=FALSE, message=FALSE}

# data_somlit_ts <- read_delim("./Data/data_PointB/DATA_SOMLIT_07-22_impute_0m.csv",
#     delim = ";", escape_double = FALSE, trim_ws = TRUE) %>%
#   dplyr::filter(datetime >= "2014-01-07") %>%
#   mutate(Month = format(datetime, format="%m"))

#Transfome data into TS, frequency : 365.25/7 = 52.17857
ts_oxy_som <- ts(SOMLIT$oxy_umol_kg, freq = 52)

# look at the NA in the ts before filling
ggplot_na_distribution(ts_oxy_som )

# filling datat(imputation) with na_seadec() = Removes the seasonal component from the time series, performs imputation on the deseasonalized series and afterwards adds the seasonal component again.
ts_seadec <-  na_seadec(ts_oxy_som, algorithm = "ma",find_frequency = TRUE)  
# plot 
ggplot_na_imputations(x_with_na = ts_oxy_som,
                          x_with_imputations = ts_seadec,
                          color_imputations = "gold")
 
 
#decompose on filled data : ts_seadec
ts_oxy_som_decomp <- decompose(ts_seadec, type = "additive")

autoplot(ts_oxy_som_decomp) +
  xlab(' ')

#remove seasonality :
ts_oxy_som_less_season <- ts_oxy_som_decomp$x - ts_oxy_som_decomp$seasonal

plot_ts_oxy_som_less_season <- autoplot(ts_oxy_som_less_season) +
  xlab("") + ylab("Oxygen (µmol kg") +
  ggtitle("Oxygen time series at Pt B (without seasonality)") +
  geom_smooth(method=lm)

#regression values :
reg_val_ts_oxy_som_less_season <- as.data.frame(plot_ts_oxy_som_less_season$data)

model_ts_oxy_less_season <- lm(formula=y~x, data=reg_val_ts_oxy_som_less_season)
summary(model_ts_oxy_less_season)


#plot with slope value :
plot_ts_oxy_som_less_season <- autoplot(ts_oxy_som_less_season) +
  xlab("") + ylab("Oxygen (umol kg") +
  ggtitle("Oxygen time series at Pt B (without seasonality)") +
  geom_smooth(method=lm) +
  annotate(geom="text", x=10, y=150, label="slope =  0.46 ± 0.15 (P = 0.001 (**), significative)", color="black")

plot_ts_oxy_som_less_season
```

Slope is `r round(model_ts_oxy_less_season$coefficients[[2]],3)`.

### AOU

```{r moving average SOMLIT aou, echo=FALSE, warning=FALSE, message=FALSE}

ts_aou_som <- ts(SOMLIT$AOU, freq = 52)

# look at the NA in the ts before filling
ggplot_na_distribution(x= SOMLIT$AOU, x_axis_labels= SOMLIT$datetime )

# filling datat(imputation) with na_seadec() = Removes the seasonal component from the time series, performs imputation on the deseasonalized series and afterwards adds the seasonal component again.
ts_seadec_aou <-  na_seadec(ts_aou_som, algorithm = "ma",find_frequency = TRUE)  

# plot 
ggplot_na_imputations(x_with_na = SOMLIT$AOU,
                          x_with_imputations = ts_seadec_aou,
                          color_imputations = "gold")
###########
#decompose on filled data : ts_seadec_aou
ts_aou_som_decomp <- decompose(ts_seadec_aou, type = "additive")

autoplot(ts_aou_som_decomp) +
  xlab(' ')

#remove seasonality :
ts_aou_som_less_season <- ts_aou_som_decomp$x - ts_aou_som_decomp$seasonal

plot_ts_aou_som_less_season <- autoplot(ts_aou_som_less_season) +
  xlab("") + ylab("aou") +
  ggtitle("AOU time series at SOMLIT (without seasonality)") +
  geom_smooth(method=lm)

#regression values :
reg_val_ts_aou_som_less_season <- as.data.frame(plot_ts_aou_som_less_season$data)

model_ts_aou_less_season <- lm(formula=y~x, data=reg_val_ts_aou_som_less_season)
summary(model_ts_aou_less_season)


#plot with slope value :
slope <- round(model_ts_aou_less_season$coefficients[[2]],4)
std <- round(sqrt(diag(vcov(model_ts_aou_less_season)))[[2]], 5)
pv <-summary(model_ts_aou_less_season)$coefficients[[8]]

plot_ts_aou_som_less_season <- autoplot(ts_aou_som_less_season) +
  xlab("") + ylab("aou (°C)") +
  ggtitle("AOU time series at SOMLIT (without seasonality)") +
  geom_smooth(method=lm) +
  annotate(geom="text", x=100, y=100, label=paste0( "slope: ",  slope," ± ", std, ", P-value: ", pv), color="black")

plot_ts_aou_som_less_season
```

Slope is `r round(model_ts_aou_less_season$coefficients[[2]],3)`.

### pH spectro

```{r moving average SOMLIT ph, echo=FALSE, warning=FALSE, message=FALSE}

# subset dataframe to start in 2015 with pH, we remove 2014 which is not full
SOMLIT_ph_spec <- SOMLIT %>% 
  dplyr::filter(datetime >= "2015-01-01")

# Periode shorter : 
ts_ph_som <- ts(SOMLIT_ph_spec$pH_spec_insitu,  frequency = 52)
frequency(ts_ph_som)
# look at the NA in the ts before filling
ggplot_na_distribution(ts_ph_som ) + coord_cartesian(ylim=c(7.85, 8.16))

# filling datat(imputation) with na_seadec() = Removes the seasonal component from the time series, performs imputation on the deseasonalized series and afterwards adds the seasonal component again.
ts_seadec_ph <-  na_seadec(ts_ph_som, algorithm = "ma",find_frequency = TRUE)  
# plot 
ggplot_na_imputations(x_with_na = SOMLIT_ph_spec$pH_spec_insitu,
                          x_with_imputations = ts_seadec_ph,
                          color_imputations = "gold")
###########
#decompose on filled data : ts_seadec_ph
ts_ph_som_decomp <- decompose(ts_seadec_ph, type = "additive")

autoplot(ts_ph_som_decomp) +
  xlab(' ')

#remove seasonality :
ts_ph_som_less_season <- ts_ph_som_decomp$x - ts_ph_som_decomp$seasonal

plot_ts_ph_som_less_season <- autoplot(ts_ph_som_less_season) +
  xlab("") + ylab("ph") +
  ggtitle("pH time series at SOMLIT (without seasonality)") +
  geom_smooth(method=lm)

#regression values :
reg_val_ts_ph_som_less_season <- as.data.frame(plot_ts_ph_som_less_season$data)

model_ts_ph_less_season <- lm(formula=y~x, data=reg_val_ts_ph_som_less_season)
summary(model_ts_ph_less_season)


#plot with slope value :
# library(broom)
# tab <- broom::tidy(lm_fit)
# broom::tidy(model_ts_ph_less_season) %>%
#   pull(p.value)
# broom::tidy(model_ts_ph_less_season) %>%
#   dplyr::filter(term=="x") %>%
#   pull(p.value)

slope <- round(model_ts_ph_less_season$coefficients[[2]],4)
std <- round(sqrt(diag(vcov(model_ts_ph_less_season)))[[2]], 5)
pv <-summary(model_ts_ph_less_season)$coefficients[[8]]

plot_ts_ph_som_less_season <- autoplot(ts_ph_som_less_season) +
  xlab("") + ylab("pH") +
  ggtitle("pH spec time series (2015-2022) at SOMLIT (without seasonality)") +
  geom_smooth(method=lm) +
  annotate(geom="text", x=4, y=8.15, label=paste0( "slope: ",  slope," ± ", std, ", P-value: ", pv), color="black")

plot_ts_ph_som_less_season
```

Slope is `r round(model_ts_ph_less_season$coefficients[[2]],4)`.

### pH spect 25°C

```{r moving average SOMLIT ph spec 25, echo=FALSE, warning=FALSE, message=FALSE}

# subset dataframe to start in 2015 with pH, we remove 2014 which is not full
SOMLIT_ph_spec25 <- SOMLIT %>% 
  dplyr::filter(datetime >= "2016-07-12") %>%   #"2016-07-12
  dplyr::select(pH_spec_25,pH_spec_insitu, everything())

# Periode shorter : 
ts_ph25_som <- ts(SOMLIT_ph_spec25$pH_spec_25,  frequency = 52)
# look at the NA in the ts before filling
ggplot_na_distribution(ts_ph25_som ) + coord_cartesian(ylim=c(7.8, 8.06))

# filling datat(imputation) with na_seadec() = Removes the seasonal component from the time series, performs imputation on the deseasonalized series and afterwards adds the seasonal component again.
ts_seadec_ph25 <-  na_seadec(ts_ph25_som, algorithm = "ma",find_frequency = TRUE)  
# plot 
ggplot_na_imputations(x_with_na = SOMLIT_ph_spec25$pH_spec_25,
                          x_with_imputations = ts_seadec_ph25,
                          color_imputations = "gold")
###########
#decompose on filled data : ts_seadec_ph18
ts_ph25_som_decomp <- decompose(ts_seadec_ph25, type = "additive")

autoplot(ts_ph25_som_decomp) +
  xlab(' ')

#remove seasonality :
ts_ph25_som_less_season <- ts_ph25_som_decomp$x - ts_ph25_som_decomp$seasonal

plot_ts_ph25_som_less_season <- autoplot(ts_ph25_som_less_season) +
  xlab("") + ylab("ph") +
  ggtitle("pH25 time series at SOMLIT (without seasonality)") +
  geom_smooth(method=lm)

#regression values :
reg_val_ts_ph25_som_less_season <- as.data.frame(plot_ts_ph25_som_less_season$data)

model_ts_ph25_less_season <- lm(formula=y~x, data=reg_val_ts_ph25_som_less_season)
summary(model_ts_ph25_less_season)


#plot with slope value :
# library(broom)
# tab <- broom::tidy(lm_fit)
# broom::tidy(model_ts_ph_less_season) %>%
#   pull(p.value)
# broom::tidy(model_ts_ph_less_season) %>%
#   dplyr::filter(term=="x") %>%
#   pull(p.value)

slope <- round(model_ts_ph25_less_season$coefficients[[2]],4)
std <- round(sqrt(diag(vcov(model_ts_ph25_less_season)))[[2]], 5)
pv <-summary(model_ts_ph25_less_season)$coefficients[[8]]

plot_ts_ph25_som_less_season <- autoplot(ts_ph25_som_less_season) +
  xlab("") + ylab("pH") +
  ggtitle("pH 25 spec time series (2016-2022) at SOMLIT (without seasonality)") +
  geom_smooth(method=lm) +
  annotate(geom="text", x=4, y=8.02, label=paste0( "slope: ",  slope," ± ", std, ", P-value: ", pv), color="black")

plot_ts_ph25_som_less_season
```

Slope is `r round(model_ts_ph25_less_season$coefficients[[2]],4)`.

### AT

- Data before 2022-09-28.

```{r moving average SOMLIT AT, echo=FALSE, warning=FALSE, message=FALSE}

# subset dataframe to start in 2015 with pH, we remove 2014 which is not full
SOMLIT_atct <- SOMLIT %>% 
  dplyr::filter(datetime <= "2022-09-28")

# Periode shorter : 
ts_at_som <- ts(SOMLIT_atct$ta,  frequency = 52)
# look at the NA in the ts before filling
ggplot_na_distribution(ts_at_som ) + coord_cartesian(ylim=c(2430, 2600))

# filling datat(imputation) with na_seadec() = Removes the seasonal component from the time series, performs imputation on the deseasonalized series and afterwards adds the seasonal component again.
ts_seadec_ta <-  na_seadec(ts_at_som, algorithm = "ma",find_frequency = TRUE)  
# plot 
ggplot_na_imputations(x_with_na = SOMLIT_atct$ta,
                          x_with_imputations = ts_seadec_ta,
                          color_imputations = "gold")
###########
#decompose on filled data : ts_seadec_ph18
ts_ta_som_decomp <- decompose(ts_seadec_ta, type = "additive")

autoplot(ts_ta_som_decomp) +
  xlab(' ')

#remove seasonality :
ts_ta_som_less_season <- ts_ta_som_decomp$x - ts_ta_som_decomp$seasonal

plot_ts_ta_som_less_season <- autoplot(ts_ta_som_less_season) +
  xlab("") + ylab("TA") +
  ggtitle("TA time series at SOMLIT (without seasonality)") +
  geom_smooth(method=lm)

#regression values :
reg_val_ts_ta_som_less_season <- as.data.frame(plot_ts_ta_som_less_season$data)

model_ts_ta_less_season <- lm(formula=y~x, data=reg_val_ts_ta_som_less_season)
summary(model_ts_ta_less_season)


#plot with slope value :
# library(broom)
# tab <- broom::tidy(lm_fit)
# broom::tidy(model_ts_ph_less_season) %>%
#   pull(p.value)
# broom::tidy(model_ts_ph_less_season) %>%
#   dplyr::filter(term=="x") %>%
#   pull(p.value)

slope <- round(model_ts_ta_less_season$coefficients[[2]],4)
std <- round(sqrt(diag(vcov(model_ts_ta_less_season)))[[2]], 5)
pv <-summary(model_ts_ta_less_season)$coefficients[[8]]

plot_ts_ta_som_less_season <- autoplot(ts_ta_som_less_season) +
  xlab("") + ylab("pH") +
  ggtitle("TA time series (2016-2022) at SOMLIT (without seasonality)") +
  geom_smooth(method=lm) +
  annotate(geom="text", x=10, y=2490, label=paste0( "slope: ",  slope," ± ", std, ", P-value: ", pv), color="black")

plot_ts_ta_som_less_season
```

### DIC

- Data before 2022-09-29.

```{r moving average SOMLIT dic, echo=FALSE, warning=FALSE, message=FALSE}

# Periode shorter : 
ts_dic_som <- ts(SOMLIT_atct$dic,  frequency = 52)
# look at the NA in the ts before filling
ggplot_na_distribution(ts_dic_som ) + coord_cartesian(ylim=c(2180, 2350))

# filling datat(imputation) with na_seadec() = Removes the seasonal component from the time series, performs imputation on the deseasonalized series and afterwards adds the seasonal component again.
ts_seadec_dic <-  na_seadec(ts_dic_som, algorithm = "ma",find_frequency = TRUE)  
# plot 
ggplot_na_imputations(x_with_na = SOMLIT_atct$dic,
                          x_with_imputations = ts_seadec_dic,
                          color_imputations = "gold")
###########
#decompose on filled data : ts_seadec_ph18
ts_dic_som_decomp <- decompose(ts_seadec_dic, type = "additive")

autoplot(ts_dic_som_decomp) +
  xlab(' ')

#remove seasonality :
ts_dic_som_less_season <- ts_dic_som_decomp$x - ts_dic_som_decomp$seasonal

plot_ts_dic_som_less_season <- autoplot(ts_dic_som_less_season) +
  xlab("") + ylab("dic") +
  ggtitle("dic time series at SOMLIT (without seasonality)") +
  geom_smooth(method=lm)

#regression values :
reg_val_ts_dic_som_less_season <- as.data.frame(plot_ts_dic_som_less_season$data)

model_ts_dic_less_season <- lm(formula=y~x, data=reg_val_ts_dic_som_less_season)
summary(model_ts_dic_less_season)


#plot with slope value :
# library(broom)
# tab <- broom::tidy(lm_fit)
# broom::tidy(model_ts_ph_less_season) %>%
#   pull(p.value)
# broom::tidy(model_ts_ph_less_season) %>%
#   dplyr::filter(term=="x") %>%
#   pull(p.value)

slope <- round(model_ts_dic_less_season$coefficients[[2]],4)
std <- round(sqrt(diag(vcov(model_ts_dic_less_season)))[[2]], 5)
pv <-summary(model_ts_dic_less_season)$coefficients[[8]]

plot_ts_dic_som_less_season <- autoplot(ts_dic_som_less_season) +
  xlab("") + ylab("pH") +
  ggtitle("dic time series (2016-2022) at SOMLIT (without seasonality)") +
  geom_smooth(method=lm) +
  annotate(geom="text", x=10, y=2330, label=paste0( "slope: ",  slope," ± ", std, ", P-value: ", pv), color="black")

plot_ts_dic_som_less_season
```

Slope is `r round(model_ts_dic_less_season$coefficients[[2]],4)`.

### pCO2

- pCO2 from 2007-01 to 2008-12.
- pCO2_w from 2009-01 to 2022-09.
- pCO2_atmos from 2009-01 to 2021-12
- NpCO2 from 2009-01 to 2022-09.
- TpCO2 from 2009-01 to 2022-12.

```{r moving average SOMLIT npco2, echo=FALSE, warning=FALSE, message=FALSE}
# Periode shorter : 
ts_npco2_som <- ts(dplyr::select(SOMLIT %>% dplyr::filter(datetime > "2009-01-01" & datetime< "2021-12-31"), NpCO2),  frequency = 52)
# look at the NA in the ts before filling
ggplot_na_distribution(ts_npco2_som ) + coord_cartesian(ylim=c(300, 510))

# filling datat(imputation) with na_seadec() = Removes the seasonal component from the time series, performs imputation on the deseasonalized series and afterwards adds the seasonal component again.
ts_seadec_npco2 <-  na_seadec(ts_npco2_som, algorithm = "ma",find_frequency = TRUE)  
# plot 
ggplot_na_imputations(x_with_na = dplyr::select(SOMLIT %>% dplyr::filter(datetime > "2009-01-01" & datetime< "2021-12-31"), NpCO2),
                          x_with_imputations = ts_seadec_npco2,
                          color_imputations = "gold")
###########
#decompose on filled data : ts_seadec_ph18
ts_npco2_som_decomp <- decompose(ts_seadec_npco2, type = "additive")

autoplot(ts_npco2_som_decomp) +
  xlab(' ')

#remove seasonality :
ts_npco2_som_less_season <- ts_npco2_som_decomp$x - ts_npco2_som_decomp$seasonal

plot_ts_npco2_som_less_season <- autoplot(ts_npco2_som_less_season) +
  xlab("") + ylab("NpCO2") +
  ggtitle("NpCO2 time series at SOMLIT (without seasonality)") +
  geom_smooth(method=lm)

#regression values :
reg_val_ts_npco2_som_less_season <- as.data.frame(plot_ts_npco2_som_less_season$data)

model_ts_npco2_less_season <- lm(formula=y~x, data=reg_val_ts_npco2_som_less_season)
summary(model_ts_npco2_less_season)


#plot with slope value :
# library(broom)
# tab <- broom::tidy(lm_fit)
# broom::tidy(model_ts_ph_less_season) %>%
#   pull(p.value)
# broom::tidy(model_ts_ph_less_season) %>%
#   dplyr::filter(term=="x") %>%
#   pull(p.value)

slope <- round(model_ts_npco2_less_season$coefficients[[2]],4)
std <- round(sqrt(diag(vcov(model_ts_npco2_less_season)))[[2]], 5)
pv <-summary(model_ts_npco2_less_season)$coefficients[[8]]

plot_ts_npco2_som_less_season <- autoplot(ts_npco2_som_less_season) +
  xlab("") + ylab("NpCO2") +
  ggtitle("NpCO2 time series (2009-01 to 2021-12) at SOMLIT (without seasonality)") +
  geom_smooth(method=lm) +
  annotate(geom="text", x=10, y=330, label=paste0( "slope: ",  slope," ± ", std, ", P-value: ", pv), color="black")

plot_ts_npco2_som_less_season
```

Slope is `r round(model_ts_npco2_less_season$coefficients[[2]],4)`.

```{r moving average SOMLIT tpco2, echo=FALSE, warning=FALSE, message=FALSE}
# Periode shorter : 
ts_tpco2_som <- ts(dplyr::select(SOMLIT %>% dplyr::filter(datetime > "2009-01-01" & datetime< "2022-12-31"), TpCO2),  frequency = 52)
# look at the NA in the ts before filling
ggplot_na_distribution(ts_tpco2_som ) + coord_cartesian(ylim=c(300, 620))

# filling datat(imputation) with na_seadec() = Removes the seasonal component from the time series, performs imputation on the deseasonalized series and afterwards adds the seasonal component again.
ts_seadec_tpco2 <-  na_seadec(ts_tpco2_som, algorithm = "ma",find_frequency = TRUE)  
# plot 
ggplot_na_imputations(x_with_na = dplyr::select(SOMLIT %>% dplyr::filter(datetime > "2009-01-01" & datetime< "2022-12-31"), TpCO2),
                          x_with_imputations = ts_seadec_tpco2,
                          color_imputations = "gold")
###########
#decompose on filled data : ts_seadec_ph18
ts_tpco2_som_decomp <- decompose(ts_seadec_tpco2, type = "additive")

autoplot(ts_tpco2_som_decomp) +
  xlab(' ')

#remove seasonality :
ts_tpco2_som_less_season <- ts_tpco2_som_decomp$x - ts_tpco2_som_decomp$seasonal

plot_ts_tpco2_som_less_season <- autoplot(ts_tpco2_som_less_season) +
  xlab("") + ylab("TpCO2") +
  ggtitle("TpCO2 time series at SOMLIT (without seasonality)") +
  geom_smooth(method=lm)

#regression values :
reg_val_ts_tpco2_som_less_season <- as.data.frame(plot_ts_tpco2_som_less_season$data)

model_ts_tpco2_less_season <- lm(formula=y~x, data=reg_val_ts_tpco2_som_less_season)
summary(model_ts_tpco2_less_season)


#plot with slope value :
# library(broom)
# tab <- broom::tidy(lm_fit)
# broom::tidy(model_ts_ph_less_season) %>%
#   pull(p.value)
# broom::tidy(model_ts_ph_less_season) %>%
#   dplyr::filter(term=="x") %>%
#   pull(p.value)

slope <- round(model_ts_tpco2_less_season$coefficients[[2]],4)
std <- round(sqrt(diag(vcov(model_ts_tpco2_less_season)))[[2]], 5)
pv <-summary(model_ts_tpco2_less_season)$coefficients[[8]]

plot_ts_tpco2_som_less_season <- autoplot(ts_tpco2_som_less_season) +
  xlab("") + ylab("TpCO2") +
  ggtitle("TpCO2 time series (2009-01 to 2022-12) at SOMLIT (without seasonality)") +
  geom_smooth(method=lm) +
  annotate(geom="text", x=10, y=330, label=paste0( "slope: ",  slope," ± ", std, ", P-value: ", pv), color="black")

plot_ts_tpco2_som_less_season
```

Slope is `r round(model_ts_tpco2_less_season$coefficients[[2]],4)`.

## Mann-Kendall

- The function kendall.SeasonalTrendTest returns estimated values of Kendall's τ, the slope, and the intercept for each season, as well as a single estimate for each of these three quantities combined over all seasons. The overall estimate of τ is the weighted average of the p seasonal τ's

- The overall estimate of slope is the median of all two-point slopes computed within each season.

- The overall estimate of intercept is the median of the p seasonal estimates of intercept.

--------

- The Kendall Tau, or Kendall rank correlation coefficient, measures the monotony of the slope. Kendall's Tau varies between -1 and 1; it is positive when the trend increases and negative when the trend decreases.

- The Sen slope, which estimates the overall slope of the time series. This slope corresponds to the median of all the slopes calculated between each pair of points in the series.

- The significance, which represents the threshold for which the hypothesis that there is no trend is accepted. The trend is statistically significant when the p-value is less than 0.05.

Results:

### Temperature

```{r mann kendall SOM temp, echo=FALSE, warning=FALSE, message=FALSE}
FFtemp_som <- kendallSeasonalTrendTest(impute_temp ~ month(datetime) + year(datetime), data = SOMLIT)
FFtemp_som$estimate
```

### Salinity

```{r mann kendall SOM sal, echo=FALSE, warning=FALSE, message=FALSE}
FFsal_som <- kendallSeasonalTrendTest(impute_sal ~ month(datetime) + year(datetime), data = SOMLIT)
FFsal_som$estimate
```

### Oxygen

```{r mann kendall SOM oxy, echo=FALSE, warning=FALSE, message=FALSE}
FFoxy_som <- kendallSeasonalTrendTest(oxy_umol_kg ~ month(datetime) + year(datetime), data = SOMLIT)
FFoxy_som$estimate
```

### AOU

```{r mann kendall SOM aou, echo=FALSE, warning=FALSE, message=FALSE}
FFaou_som <- kendallSeasonalTrendTest(AOU ~ month(datetime) + year(datetime), data = SOMLIT)
FFaou_som$estimate
```

### pH spectro

```{r mann kendall SOM ph, echo=FALSE, warning=FALSE, message=FALSE}
FFph_som <- kendallSeasonalTrendTest(pH_spec_insitu ~ month(datetime) + year(datetime), data = SOMLIT)
FFph_som$estimate
```

### pH spectro 25°C

```{r mann kendall SOM ph25, echo=FALSE, warning=FALSE, message=FALSE}
FFph25_som <- kendallSeasonalTrendTest(pH_spec_25 ~ month(datetime) + year(datetime), data = SOMLIT)
FFph25_som$estimate
```

### TA

```{r mann kendall SOM ta, echo=FALSE, warning=FALSE, message=FALSE}
FFta_som <- kendallSeasonalTrendTest(ta ~ month(datetime) + year(datetime), data = SOMLIT)
FFta_som$estimate
```

### DIC

```{r mann kendall SOM dic, echo=FALSE, warning=FALSE, message=FALSE}
FFdic_som <- kendallSeasonalTrendTest(dic ~ month(datetime) + year(datetime), data = SOMLIT)
FFdic_som$estimate
```


### NpCO2

```{r mann kendall SOM NpCO2, echo=FALSE, warning=FALSE, message=FALSE}
FFnpco2_som <- kendallSeasonalTrendTest(NpCO2 ~ month(datetime) + year(datetime), data = SOMLIT)
FFnpco2_som$estimate
```

### TpCO2

```{r mann kendall SOM TpCO2, echo=FALSE, warning=FALSE, message=FALSE}
FFtpco2_som <- kendallSeasonalTrendTest(TpCO2 ~ month(datetime) + year(datetime), data = SOMLIT)
FFtpco2_som$estimate
```
