---
title: "EOL_HF"
author: "samir"
date: '`r format(Sys.time(), "%d %B %Y %H:%M")`'

output:
    rmdformats::robobook:
     self_contained: true
     thumbnails: false
     lightbox: true
     gallery: true
     highlight: pygments
---

```{r libraries, echo=FALSE, warning=FALSE, message=FALSE}

library("readr")
library("viridis")
library("scales")
library("lubridate")
library("gsignal")
library("FactoMineR")
library("factoextra")
library("gplots")
library("heatwaveR")
library("dplyr")
library("imputeTS")
library("Metrics")
library("tidyr")
library("tseries")
library("forecast")
library("tsDyn")
library("tsoutliers")
library("seacarb")
library("lmodel2")
library("lmtest")
library("ggExtra")
library("corrplot")
library("ggcorrplot")
library("marelac")
library("spectral")
library("signal")
library("cowplot")
library("kableExtra")
library("EnvStats")

```

```{r importation Data EOL raw, echo=FALSE, warning=FALSE, message=FALSE}

#period : septembre 2013-octobre 2022
EOL_RAW <- read_delim("./Data/data_EOL/EOL_raw_13-23.csv", 
    delim = ",", escape_double = FALSE, trim_ws = TRUE)  

EOL_RAW <- EOL_RAW %>% 
  rename(temperature = temp_eol_c, salinity = sal_eol_psu, oxygen = oxy_umol_kg) %>% 
  unite(col = "datetime","datetime", "hour", sep = " ", remove = TRUE)
EOL_RAW$datetime <- dmy_hms(EOL_RAW$datetime) 

EOL_RAW <- EOL_RAW %>% 
  dplyr::filter(datetime >= "2018-01-01 00:00:00" & datetime <= "2022-12-31 23:59:59")

```

# EOL Raw data: January 2018 to December 2022

Gaps:

- 21 Mar to 17 May 2019.
- 22 to 24 Nov 2022

NAs:

- No NAs for T
- NAs for S, O, AOU
  
```{r visualisation data EOL, echo=FALSE, warning=FALSE, message=FALSE, fig.width=9,fig.height=5}

# calcul OSAT (ml/kg) + calcul AOU EOL umol/kg :
EOL_RAW <- EOL_RAW %>% 
  mutate(T1 = (temperature + 273.15) / 100,
         OSAT = exp((((-173.9894) + 255.5907/T1 + 146.4813*log(T1) - 22.2040*T1)) + salinity * (-0.037362 + T1 * (0.016504 - 0.0020564 * T1))),
         AOU = OSAT - oxygen)

#neg_AOU_EOL = EOL_RAW$oxygen - OSAT

plot_EOL_RAW_temp <- EOL_RAW %>% 
  ggplot() +
  aes(x=datetime, y=temperature) +
  geom_line() +
  scale_x_datetime(name="", date_labels = "%d-%Y") +
  scale_y_continuous(name="Temperature (°C)")

plot_EOL_RAW_sal <- EOL_RAW %>% 
  ggplot() +
  aes(x=datetime, y=salinity) +
  geom_line() +
  scale_x_datetime(name="", date_labels = "%d-%Y") +
  scale_y_continuous(name="Salinity")

plot_EOL_RAW_oxy <- EOL_RAW %>% 
  ggplot() +
  aes(x=datetime, y=oxygen) +
  geom_line() +
  scale_x_datetime(name="", date_labels = "%d-%Y") +
  scale_y_continuous(name=expression(paste("Oxygen (μmol ", kg^-1, ")")))

plot_EOL_RAW_AOU <- EOL_RAW %>% 
  ggplot() +
  aes(x=datetime, y=AOU) +
  geom_line() +
  scale_x_datetime(name="", date_labels = "%d-%Y") +
  scale_y_continuous(name=expression(paste("AOU (μmol ", kg^-1, ")")))

 plot_grid(plot_EOL_RAW_temp, 
            plot_EOL_RAW_sal,
            plot_EOL_RAW_oxy,
            plot_EOL_RAW_AOU,
            align='vh', ncol=2)

# visualisation of gaps
 
ggplot_na_distribution(x= EOL_RAW$temperature, x_axis_labels= EOL_RAW$datetime )
ggplot_na_distribution(x= EOL_RAW$salinity, x_axis_labels= EOL_RAW$datetime )
ggplot_na_distribution(x= EOL_RAW$oxygen, x_axis_labels= EOL_RAW$datetime )
ggplot_na_distribution(x= EOL_RAW$AOU, x_axis_labels= EOL_RAW$datetime )

```

# Anomalies method

## By months
  
```{r calcul monthly means, echo=FALSE, warning=FALSE, message=FALSE}
monthly_means <- ungroup(EOL_RAW) %>% 
  #mutate(Month = format(datetime, format="%m")) %>% 
  group_by(Month) %>%
  summarise(
    sal_month = mean(salinity, na.rm = TRUE),
    sd_sal = sd(salinity, na.rm = TRUE),
    temp_month = mean(temperature, na.rm = TRUE),
    sd_temp = sd(temperature, na.rm = TRUE),
    oxy_month = mean(oxygen, na.rm = TRUE),
    sd_oxy = sd(oxygen, na.rm= TRUE),
    aou_month = mean(AOU, na.rm = TRUE),
    sd_aou = sd(AOU, na.rm= TRUE))
```

```{r regression analysis, echo=FALSE, warning=FALSE, message=FALSE}

anomalies_EOL_month <- left_join(ungroup(EOL_RAW), monthly_means, by = "Month") %>%
  mutate(sal_ano = salinity - sal_month,
         temp_ano = temperature - temp_month,
         oxy_ano = oxygen - oxy_month,
         aou_ano = AOU - aou_month)

# Regression and table anomalies
var_list <- c("temp_ano", "sal_ano","oxy_ano", "aou_ano")

lms <- lapply(var_list, function(x) {
  summary(lm(substitute(i ~ decimal_date(datetime), list(i = as.name(x))), 
             data = anomalies_EOL_month))
})
reg_raw <- NULL
for (i in 1:length(var_list)) {
  #one loops through all anomalies
  # calculate probability of fstatistic because it cannot be extracted from lms above
  # see http://stats.stackexchange.com/questions/92824/extracting-the-model-p-value-for-a-multiple-regression-in-r
  # slope returned by lm is per second
  prob <-
    pf(lms[[i]]$fstatistic[1],
       lms[[i]]$fstatistic[2],
       lms[[i]]$fstatistic[3],
       lower.tail = FALSE)
  reg_raw <-
    rbind(reg_raw, as.numeric(
      c(
        lms[[i]]$coefficients[2, 1],
        lms[[i]]$coefficients[2, 2],
        lms[[i]]$coefficients[2, 4],
        lms[[i]]$coefficients[1, 1],
        lms[[i]]$coefficients[1, 2],
        lms[[i]]$coefficients[1, 4],
        lms[[i]]$fstatistic[1],
        lms[[i]]$fstatistic[3],
        lms[[i]]$r.squared,
        prob
      )
    ))
}
colnames(reg_raw) <-
  c("Slope",
    "SE Slope",
    "P Slope",
    "Intercept",
    "SE int.",
    "P int.",
    "F",
    "df",
    "R2",
    "P value")

row.names(reg_raw) <- var_list
reg_anomalies_raw <- reg_raw
kable(reg_anomalies_raw, caption="Time series anomaly regression analyses per month", format = "pipe", align='l')

##RAPPORT##
# reg_anomalies_raw_rapport <- as.data.frame(reg_anomalies_raw)
# reg_anomalies_raw_rapport <- reg_anomalies_raw_rapport %>% 
#   mutate(Variables = c("T (°C)", "S"))
# 
# reg_anomalies_raw_rapport <- reg_anomalies_raw_rapport %>% 
#   mutate( 
#   `Pente ± sd` = paste(round(reg_anomalies_raw_rapport$Slope, digits=3), "±", round(reg_anomalies_raw_rapport$`SE Slope`, digits=4)), 
#   `P pente` = reg_anomalies_raw_rapport$`P Slope`, 
#   `Int. ± sd` = paste(round(reg_anomalies_raw_rapport$Intercept, digits=2), "±", round(reg_anomalies_raw_rapport$`SE int.`, digits=2)),
#   `P Int.` = reg_anomalies_raw_rapport$`P int.`, 
#   `R^2^` = round(reg_anomalies_raw_rapport$R2, digits=4))
# 
# 
# reg_anomalies_raw_rapport <- reg_anomalies_raw_rapport[,c(11:16)]
# 
# kable(reg_anomalies_raw_rapport, caption="Time series anomaly regression analyses (EOL)", format = "pipe", align='l')

```

### Temperature

```{r EOL month - plot anomalies + trends - temperature, echo=FALSE, warning=FALSE, message=FALSE}

plot_ano_temp_EOL_month <- ggplot(data = anomalies_EOL_month, aes(x = datetime, y = temp_ano), na.rm=TRUE) +
  #scale_x_datetime(date_breaks="2 year", date_minor_breaks="1 years", labels = date_format("%Y")) +
  geom_point(colour="blue", na.rm=TRUE, size=0.65) +
  geom_smooth(method=lm, colour="black", fill="grey", linewidth=0.6, na.rm=TRUE) +
  labs(title="EOL - Temperature anomalies + trend",x="", y="Temperature (°C)")+
  annotate(geom="text", x=as.POSIXct("2019-12-03 01:00:00"), y=5.5, label="slope : 0.032 ± 0.004 (P: 2.316919e-15 - significative)", color="black")

plot_ano_temp_EOL_month
```

### Salinity

```{r EOL month - plot anomalies + trends - salinity, echo=FALSE, warning=FALSE, message=FALSE}

plot_ano_sal_EOL_month <- ggplot(data = anomalies_EOL_month, aes(x = datetime, y = sal_ano), na.rm=TRUE) +
  #scale_x_datetime(date_breaks="2 year", date_minor_breaks="1 years", labels = date_format("%Y")) +
  geom_point(colour="blue", na.rm=TRUE, size=0.65) +
  geom_smooth(method=lm, colour="black", fill="grey", linewidth=0.6, na.rm=TRUE) +
  labs(title="EOL - Salinity anomalies + trend ",x="", y="Salinity") +
  annotate(geom="text", x=as.POSIXct("2019-12-03 01:00:00"), y=1, label="slope : 0.026 ± 0.0007 (P: 2.728093e-263 - significative)", color="black")

plot_ano_sal_EOL_month

```

### Oxygen 

```{r EOL month - plot anomalies + trends - oxygen, echo=FALSE, warning=FALSE, message=FALSE}

plot_ano_oxy_EOL_month <- ggplot(data = anomalies_EOL_month, aes(x = datetime, y = oxy_ano), na.rm=TRUE) +
  #scale_x_datetime() +
  geom_point(colour="blue", na.rm=TRUE, size=0.65) + 
  geom_smooth(method=lm, colour="black", fill="grey", linewidth=0.6, na.rm=TRUE) +
  labs(title="EOL - Oxygen anomalies + trend",x="", y=expression(paste("Oxygen (μmol ", kg^-1, ")")))+
  annotate(geom="text", x=as.POSIXct("2020-01-16 01:00:00"), y=-30, label="slope : -1.46 ± 0.027 (P: 0 - siginficative)" , color="black")
  
plot_ano_oxy_EOL_month
```

### AOU 

```{r EOL month - plot anomalies + trends - AOU, echo=FALSE, warning=FALSE, message=FALSE}
plot_ano_AOU_EOL_month <- ggplot(data = anomalies_EOL_month, aes(x = datetime, y = aou_ano), na.rm=TRUE) +
  #scale_x_datetime() +
  geom_point(colour="blue", na.rm=TRUE, size=0.65) + 
  geom_smooth(method=lm, colour="black", fill="grey", linewidth=0.6, na.rm=TRUE) +
  labs(title="EOL - AOU anomalies + trend",x="", y=expression(paste("AOU (μmol ", kg^-1, ")")))+
  annotate(geom="text", x=as.POSIXct("2019-03-16 01:00:00"), y=28, label="slope : 1.35 ± 0.023 P: 0 - significative", color="black")

plot_ano_AOU_EOL_month
```

## By days
  
```{r calcul day means, echo=FALSE, warning=FALSE, message=FALSE}
#Add julian days
EOL_RAW$Day<- yday(EOL_RAW$datetime)
day_means <- ungroup(EOL_RAW) %>% 
  #mutate(Month = format(datetime, format="%m")) %>% 
  group_by(Day) %>%
  summarise(
    sal_day = mean(salinity, na.rm = TRUE),
    sd_sal_day = sd(salinity, na.rm = TRUE),
    temp_day = mean(temperature, na.rm = TRUE),
    sd_temp_day = sd(temperature, na.rm = TRUE),
    oxy_day = mean(oxygen, na.rm = TRUE),
    sd_oxy_day = sd(oxygen, na.rm= TRUE),
    aou_day = mean(AOU, na.rm = TRUE),
    sd_aou_day = sd(AOU, na.rm= TRUE))
```

```{r regression analysis days, echo=FALSE, warning=FALSE, message=FALSE}

anomalies_EOL_day <- left_join(ungroup(EOL_RAW), day_means, by = "Day") %>%
  mutate(sal_ano_day = salinity - sal_day,
         temp_ano_day = temperature - temp_day,
         oxy_ano_day = oxygen - oxy_day,
         aou_ano_day = AOU - aou_day)

# Regression and table anomalies
var_list <- c("temp_ano_day", "sal_ano_day","oxy_ano_day", "aou_ano_day")

lms <- lapply(var_list, function(x) {
  summary(lm(substitute(i ~ decimal_date(datetime), list(i = as.name(x))), 
             data = anomalies_EOL_day))
})
reg_raw <- NULL
for (i in 1:length(var_list)) {
  #one loops through all anomalies
  # calculate probability of fstatistic because it cannot be extracted from lms above
  # see http://stats.stackexchange.com/questions/92824/extracting-the-model-p-value-for-a-multiple-regression-in-r
  # slope returned by lm is per second
  prob <-
    pf(lms[[i]]$fstatistic[1],
       lms[[i]]$fstatistic[2],
       lms[[i]]$fstatistic[3],
       lower.tail = FALSE)
  reg_raw <-
    rbind(reg_raw, as.numeric(
      c(
        lms[[i]]$coefficients[2, 1],
        lms[[i]]$coefficients[2, 2],
        lms[[i]]$coefficients[2, 4],
        lms[[i]]$coefficients[1, 1],
        lms[[i]]$coefficients[1, 2],
        lms[[i]]$coefficients[1, 4],
        lms[[i]]$fstatistic[1],
        lms[[i]]$fstatistic[3],
        lms[[i]]$r.squared,
        prob
      )
    ))
}
colnames(reg_raw) <-
  c("Slope",
    "SE Slope",
    "P Slope",
    "Intercept",
    "SE int.",
    "P int.",
    "F",
    "df",
    "R2",
    "P value")

row.names(reg_raw) <- var_list
reg_anomalies_raw_day <- reg_raw

kable(reg_anomalies_raw_day, caption="Time series anomaly regression analyses per day", format = "pipe", align='l')
```
### Temperature

```{r EOL day - plot anomalies + trends - temperature, echo=FALSE, warning=FALSE, message=FALSE}

plot_ano_temp_EOL_day <- ggplot(data = anomalies_EOL_day, aes(x = datetime, y = temp_ano_day), na.rm=TRUE) +
  #scale_x_datetime(date_breaks="2 year", date_minor_breaks="1 years", labels = date_format("%Y")) +
  geom_point(colour="blue", na.rm=TRUE, size=0.65) +
  geom_smooth(method=lm, colour="black", fill="grey", linewidth=0.6, na.rm=TRUE) +
  labs(title="EOL - Temperature anomalies + trend",x="", y="Temperature (°C)")+
  annotate(geom="text", x=as.POSIXct("2019-12-03 01:00:00"), y=5.5, label="slope : 0.034 ± 0.003 (P: 1.842844e-29 - significative)", color="black")

plot_ano_temp_EOL_day
```

### Salinity

```{r EOL day - plot anomalies + trends - salinity, echo=FALSE, warning=FALSE, message=FALSE}

plot_ano_sal_EOL_day <- ggplot(data = anomalies_EOL_day, aes(x = datetime, y = sal_ano_day), na.rm=TRUE) +
  #scale_x_datetime(date_breaks="2 year", date_minor_breaks="1 years", labels = date_format("%Y")) +
  geom_point(colour="blue", na.rm=TRUE, size=0.65) +
  geom_smooth(method=lm, colour="black", fill="grey", linewidth=0.6, na.rm=TRUE) +
  labs(title="EOL - Salinity anomalies + trend",x="", y="Salinity (°C)")+
  annotate(geom="text", x=as.POSIXct("2019-12-03 01:00:00"), y=1, label="slope : 0.025 ± 0.0007 (P: 3.362084e-280 - significative)", color="black")

plot_ano_sal_EOL_day
```

### Oxygen

```{r EOL day - plot anomalies + trends - oxygen, echo=FALSE, warning=FALSE, message=FALSE}

plot_ano_oxy_EOL_day <- ggplot(data = anomalies_EOL_day, aes(x = datetime, y = oxy_ano_day), na.rm=TRUE) +
  #scale_x_datetime() +
  geom_point(colour="blue", na.rm=TRUE, size=0.65) + 
  geom_smooth(method=lm, colour="black", fill="grey", linewidth=0.6, na.rm=TRUE) +
  labs(title="EOL - Oxygen anomalies + trend",x="", y=expression(paste("Oxygen (μmol ", kg^-1, ")")))+
  annotate(geom="text", x=as.POSIXct("2020-01-16 01:00:00"), y=-30, label="slope : -1.46 ± 0.024 (P: 0 - siginficative)" , color="black")

plot_ano_oxy_EOL_day
```

### AOU 

```{r EOL day - plot anomalies + trends - AOU, echo=FALSE, warning=FALSE, message=FALSE}
plot_ano_AOU_EOL_day <- ggplot(data = anomalies_EOL_day, aes(x = datetime, y = aou_ano_day), na.rm=TRUE) +
  #scale_x_datetime() +
  geom_point(colour="blue", na.rm=TRUE, size=0.65) + 
  geom_smooth(method=lm, colour="black", fill="grey", linewidth=0.6, na.rm=TRUE) +
  labs(title="EOL - AOU anomalies + trend ",x="", y=expression(paste("AOU (μmol ", kg^-1, ")")))+
  annotate(geom="text", x=as.POSIXct("2019-03-16 01:00:00"), y=28, label="slope : 1.36 ± 0.023 P: 0 - significative", color="black")

plot_ano_AOU_EOL_month
```

Results from anomaly methods (month and day) are similars with a slight improvement in SD.


<!-- ## Daily dataset (8AM-10AM from January 2018 to December 2022) - Temperature and Salinity -->

<!-- Missing values : -->
<!-- - april 2019 -->

<!-- ```{r importation Data EOL 2 - daily, echo=FALSE, warning=FALSE, message=FALSE} -->

<!-- # Modifier en partant du fichier EOL_raw -->
<!-- EOL_daily <- read_delim("./Data/data_EOL/EOL_13-23_1perday_morning.csv", -->
<!--     delim = ";", escape_double = FALSE, trim_ws = TRUE) %>% -->
<!--   rename(temperature=temp_eol_c, salinity=sal_eol_psu) %>% -->
<!--   mutate(datetime = as.POSIXct(datetime)) %>% -->
<!--   dplyr::filter(datetime >= "2018-01-01 00:00:00" & datetime <= "2022-12-31 23:59:00") -->

<!-- ``` -->

<!-- ```{r EOL 2 - calcul monthly means - temperature & salinity, echo=TRUE, warning=FALSE, message=FALSE} -->

<!-- monthly_means_EOL2 <- ungroup(EOL_daily) %>% -->
<!--   mutate(Month = format(datetime, format="%m")) %>% -->
<!--   group_by(Month) %>% -->
<!--   summarise( -->
<!--     sal_month = mean(salinity, na.rm = TRUE), -->
<!--     temp_month = mean(temperature, na.rm = TRUE)) -->

<!-- monthly_means_EOL2 -->

<!-- ``` -->

<!-- ```{r EOL 2 - regression analysis - temperature & salinity, echo=FALSE, warning=FALSE, message=FALSE} -->

<!-- anomalies_EOL_2 <- left_join(ungroup(EOL_daily), monthly_means_EOL2, by = "Month") %>% -->
<!--   mutate(sal_ano = salinity - sal_month, -->
<!--          temp_ano = temperature - temp_month) -->

<!-- # Regression and table anomalies -->
<!-- var_list <- -->
<!--   c( -->
<!--     "sal_ano", -->
<!--     "temp_ano") -->

<!-- lms <- lapply(var_list, function(x) { -->
<!--   summary(lm(substitute(i ~ decimal_date(datetime), list(i = as.name(x))), -->
<!--              data = anomalies_EOL_2)) -->
<!-- }) -->
<!-- reg_EOL_2 <- NULL -->
<!-- for (i in 1:length(var_list)) { -->
<!--   #one loops through all anomalies -->
<!--   # calculate probability of fstatistic because it cannot be extracted from lms above -->
<!--   # see http://stats.stackexchange.com/questions/92824/extracting-the-model-p-value-for-a-multiple-regression-in-r -->
<!--   # slope returned by lm is per second -->
<!--   prob <- -->
<!--     pf(lms[[i]]$fstatistic[1], -->
<!--        lms[[i]]$fstatistic[2], -->
<!--        lms[[i]]$fstatistic[3], -->
<!--        lower.tail = FALSE) -->
<!--   reg_EOL_2 <- -->
<!--     rbind(reg_EOL_2, as.numeric( -->
<!--       c( -->
<!--         lms[[i]]$coefficients[2, 1], -->
<!--         lms[[i]]$coefficients[2, 2], -->
<!--         lms[[i]]$coefficients[2, 4], -->
<!--         lms[[i]]$coefficients[1, 1], -->
<!--         lms[[i]]$coefficients[1, 2], -->
<!--         lms[[i]]$coefficients[1, 4], -->
<!--         lms[[i]]$fstatistic[1], -->
<!--         lms[[i]]$fstatistic[3], -->
<!--         lms[[i]]$r.squared, -->
<!--         prob -->
<!--       ) -->
<!--     )) -->
<!-- } -->
<!-- colnames(reg_EOL_2) <- -->
<!--   c("Slope", -->
<!--     "SE Slope", -->
<!--     "P Slope", -->
<!--     "Intercept", -->
<!--     "SE int.", -->
<!--     "P int.", -->
<!--     "F", -->
<!--     "df", -->
<!--     "R2", -->
<!--     "P value") -->

<!-- row.names(reg_EOL_2) <- var_list -->
<!-- reg_anomalies_EOL_2 <- reg_EOL_2 -->

<!-- reg_anomalies_EOL_2 -->

<!-- ``` -->
<!-- ## Temperature -->

<!-- ```{r EOL 2 - plot anomalies + trends - temperature, echo=FALSE, warning=FALSE, message=FALSE} -->
<!-- plot_ano_temp_EOL_2 <- ggplot(data = anomalies_EOL_2, aes(x = datetime, y = temp_ano), na.rm=TRUE) + -->
<!--   scale_x_datetime(date_breaks="2 year", date_minor_breaks="1 years", labels = date_format("%Y")) + -->
<!--   geom_point(colour="blue", na.rm=TRUE, size=0.65) + -->
<!--   geom_smooth(method=lm, colour="black", fill="grey", linewidth=0.6, na.rm=TRUE) + -->
<!--   labs(title="EOL - Temperature anomalies + trend (2018-2022, daily)",x="", y="Temperature (°C)")+ annotate(geom="text", x=as.POSIXct("2019-12-03 01:00:00"), y=5, label="slope : -0.012 ± 0.020 (non-significative: P 5.6e-01", color="black") -->

<!-- plot_temp_EOL_2 <- EOL_daily %>% -->
<!--   ggplot() + -->
<!--   aes(x=datetime, y=temperature) + -->
<!--   geom_point(size=1) + -->
<!--   scale_x_datetime(name="") + -->
<!--   scale_y_continuous(name="Temperature (°C)") + -->
<!--   ggtitle("EOL time series observations of temperature (°C) 2018-2022, daily)") -->

<!-- plot_temp_EOL_2 -->
<!-- plot_ano_temp_EOL_2 -->
<!-- ``` -->

<!-- ## Salinity -->

<!-- ```{r EOL 2 - plot anomalies + trends - salinity, echo=FALSE, warning=FALSE, message=FALSE} -->

<!-- plot_ano_sal_EOL_2 <- ggplot(data = anomalies_EOL_2, aes(x = datetime, y = sal_ano), na.rm=TRUE) + -->
<!--   scale_x_datetime(date_breaks="2 year", date_minor_breaks="1 years", labels = date_format("%Y")) + -->
<!--   geom_point(colour="blue", na.rm=TRUE, size=0.65) + -->
<!--   geom_smooth(method=lm, colour="black", fill="grey", linewidth=0.6, na.rm=TRUE) + -->
<!--   labs(title="EOL - Salinity anomalies + trend (2018-2022, daily)",x="", y="Salinity") + -->
<!--   annotate(geom="text", x=as.POSIXct("2019-12-03 01:00:00"), y=1, label="slope : 0.017 ± 0.003 (significative: P 5.5e-07)", color="black") -->

<!-- plot_sal_EOL_2 <- EOL_daily %>% -->
<!--   ggplot() + -->
<!--   aes(x=datetime, y=salinity) + -->
<!--   geom_point(size=1) + -->
<!--   scale_x_datetime(name="") + -->
<!--   scale_y_continuous(name="Salinity") + -->
<!--   ggtitle("EOL time series observations of salinity (2018-2022, daily)") -->

<!-- plot_sal_EOL_2 -->
<!-- plot_ano_sal_EOL_2 -->

<!-- ``` -->

# Other methods:

**By autocorrelation + decompose function or Mann-Kendall : not possible because of gaps**

--> Filling gaps.

## Moving average

### Temperature

No gaps to fill.

```{r moving average EOL temp, echo=FALSE, warning=FALSE, message=FALSE}

# data_somlit_ts <- read_delim("./Data/data_PointB/DATA_SOMLIT_07-22_impute_0m.csv",
#     delim = ";", escape_double = FALSE, trim_ws = TRUE) %>%
#   dplyr::filter(datetime >= "2014-01-07") %>%
#   mutate(Month = format(datetime, format="%m"))

# #https://geomoer.github.io/moer-mpg-data-analysis/unit10/unit10-02_time_series_analysis.html
# spec <- spectrum(EOL_RAW$temperature)
# plot(1/spec$freq, spec$spec, type = "h")
# 1/spec$freq[spec$spe == max(spec$spec)]
# # result of 8000 hours = 333 jours
# 
# annual_trend <- zoo::rollapply(EOL_RAW$temperature, 8000, mean, align = "center", fill = NA)
# plot(EOL_RAW$datetime, EOL_RAW$temperature, type = "l") +
# lines(EOL_RAW$datetime, annual_trend, col = "red")
# 
# seasonal <- EOL_RAW$temperature - annual_trend
# seasonal_mean <- aggregate(seasonal, by = list(rep(seq(1,7988), 5)), FUN = mean, na.rm = TRUE)
# plot(EOL_RAW$datetime, seasonal, type = "l")+
# lines(EOL_RAW$datetime, rep(seasonal_mean$x, 5), col = "blue")
# 
# remainder <- EOL_RAW$temperature - annual_trend - seasonal_mean$x
# plot(EOL_RAW$datetime, remainder, type = "l")
# 
# acf(remainder, na.action = na.pass)

#Transfome data into TS, frequency : 365.25/7 = 52.17857
#use msts or ts
#ts_temp_eol <- msts(EOL_RAW$temperature, seasonal.periods= 365.25*24)

ts_temp_eol <- ts(EOL_RAW$temperature, freq = 8766)#9982  365,25*24=8766   8766/12 = 730,5

# look at the NA in the ts before filling
ggplot_na_distribution(x= ts_temp_eol, x_axis_labels= EOL_RAW$datetime )

# No gaps to be filled

 
#decompose on filled data : 
ts_temp_eol_decomp <- decompose(ts_temp_eol, type = "additive")

autoplot(ts_temp_eol_decomp) +
  xlab(' ')

#remove seasonality :
ts_temp_eol_less_season <- ts_temp_eol_decomp$x - ts_temp_eol_decomp$seasonal

plot_ts_temp_eol_less_season <- autoplot(ts_temp_eol_less_season) +
  xlab("") + ylab("Temp. (°C)") +
  ggtitle("Temperature time series at Pt B (without seasonality)") +
  geom_smooth(method=lm)
#regression values :
reg_val_ts_temp_eol_less_season <- as.data.frame(plot_ts_temp_eol_less_season$data)

# Pour faire la regresssion en secondes, mettre datetime en x et non pas les années decimales.
#reg_val_ts_temp_eol_less_season$datetime <- EOL_RAW$datetime
#names(reg_val_ts_temp_eol_less_season) <- c("value", "datetime")

model_ts_temp_less_season <- lm(formula=y~x, data=reg_val_ts_temp_eol_less_season)
summary(model_ts_temp_less_season)

#plot with slope value :
plot_ts_temp_eol_less_season <- autoplot(ts_temp_eol_less_season) +
  xlab("") + ylab("Temp. (°C)") +
  ggtitle("Temperature time series at Pt B (without seasonality)") +
  geom_smooth(method=lm) +
  annotate(geom="text", x=4, y=10.5, label="slope :  0.18 ± 0.014 (significative, P <2e-16)", color="black")

plot_ts_temp_eol_less_season
```

Slope is `r round(model_ts_temp_less_season$coefficients[[2]],3)`. Means that periode of 5 years is not suitable. Slope not reliable.

### Salinity

```{r moving average EOL sal, echo=FALSE, warning=FALSE, message=FALSE}
ts_sal_eol <- ts(EOL_RAW$salinity, freq = 8766)

# look at the NA in the ts before filling
ggplot_na_distribution(x= EOL_RAW$salinity, x_axis_labels= EOL_RAW$datetime )

# filling datat(imputation) with na_seadec() = Removes the seasonal component from the time series, performs imputation on the deseasonalized series and afterwards adds the seasonal component again.
ts_seadec_sal <-  na_seadec(EOL_RAW$salinity, algorithm = "ma",find_frequency = TRUE)  
# plot 
ggplot_na_imputations(x_with_na = EOL_RAW$salinity,
                          x_with_imputations = ts_seadec_sal,
                          color_imputations = "gold")

#decompose on filled data : ts_seadec_sal
ts_sal_eol_decomp <- decompose(ts_seadec_sal, type = "additive")

autoplot(ts_sal_eol_decomp) +
  xlab(' ')

#remove seasonality :
ts_sal_eol_less_season <- ts_sal_eol_decomp$x - ts_sal_eol_decomp$seasonal

plot_ts_sal_eol_less_season <- autoplot(ts_sal_eol_less_season) +
  xlab("") + ylab("Sal") +
  ggtitle("Salinity time series at EOL (without seasonality)") +
  geom_smooth(method=lm)

#regression values :
reg_val_ts_sal_eol_less_season <- as.data.frame(plot_ts_sal_eol_less_season$data)

model_ts_sal_less_season <- lm(formula=y~x, data=reg_val_ts_sal_eol_less_season)
summary(model_ts_sal_less_season)

#plot with slope value :
plot_ts_sal_eol_less_season <- autoplot(ts_sal_eol_less_season) +
  xlab("") + ylab("Sal (°C)") +
  ggtitle("Salinity time series at EOL (without seasonality)") +
  geom_smooth(method=lm) +
  annotate(geom="text", x=1200, y=38.5, label="slope :  0.0001 ± 0.000002 (significative, P <2e-16)", color="black")

plot_ts_sal_eol_less_season
```

Slope is `r round(model_ts_sal_less_season$coefficients[[2]],3)`. Means that periode of 5 years is not suitable. Slope not reliable.

### Oxygen

```{r moving average EOL oxy, echo=FALSE, warning=FALSE, message=FALSE}

ts_oxy_eol <- ts(EOL_RAW$oxygen, freq = 8766)

# look at the NA in the ts before filling
ggplot_na_distribution(x= EOL_RAW$oxygen, x_axis_labels= EOL_RAW$datetime )

# filling datat(imputation) with na_seadec() = Removes the seasonal component from the time series, performs imputation on the deseasonalized series and afterwards adds the seasonal component again.
ts_seadec_oxy <-  na_seadec(EOL_RAW$oxygen, algorithm = "ma",find_frequency = TRUE)  
# plot 
ggplot_na_imputations(x_with_na = EOL_RAW$oxygen,
                          x_with_imputations = ts_seadec_oxy,
                          color_imputations = "gold")

#decompose on filled data : ts_seadec_oxy
ts_oxy_eol_decomp <- decompose(ts_seadec_oxy, type = "additive")

autoplot(ts_oxy_eol_decomp) +
  xlab(' ')

#remove seasonality :
ts_oxy_eol_less_season <- ts_oxy_eol_decomp$x - ts_oxy_eol_decomp$seasonal

plot_ts_oxy_eol_less_season <- autoplot(ts_oxy_eol_less_season) +
  xlab("") + ylab("oxy") +
  ggtitle("Oxygen time series at EOL (without seasonality)") +
  geom_smooth(method=lm)

#regression values :
reg_val_ts_oxy_eol_less_season <- as.data.frame(plot_ts_oxy_eol_less_season$data)

model_ts_oxy_less_season <- lm(formula=y~x, data=reg_val_ts_oxy_eol_less_season)
summary(model_ts_oxy_less_season)


#plot with slope value :
plot_ts_oxy_eol_less_season <- autoplot(ts_oxy_eol_less_season) +
  xlab("") + ylab("oxy (°C)") +
  ggtitle("Oxygen time series at EOL (without seasonality)") +
  geom_smooth(method=lm) +
  annotate(geom="text", x=1200, y=150, label="slope :  -0.0079 ± 0.00018 (significative, P <2e-16)", color="black")

plot_ts_oxy_eol_less_season
```
Slope is `r round(model_ts_oxy_less_season$coefficients[[2]],3)`. Means that periode of 5 years is not suitable. Slope not reliable.

### AOU

```{r moving average EOL aou, echo=FALSE, warning=FALSE, message=FALSE}

ts_aou_eol <- ts(EOL_RAW$AOU, freq = 8766)

# look at the NA in the ts before filling
ggplot_na_distribution(x= EOL_RAW$AOU, x_axis_labels= EOL_RAW$datetime )

# filling datat(imputation) with na_seadec() = Removes the seasonal component from the time series, performs imputation on the deseasonalized series and afterwards adds the seasonal component again.
ts_seadec_aou <-  na_seadec(EOL_RAW$AOU, algorithm = "ma",find_frequency = TRUE)  
# plot 
ggplot_na_imputations(x_with_na = EOL_RAW$AOU,
                          x_with_imputations = ts_seadec_aou,
                          color_imputations = "gold")

#decompose on filled data : ts_seadec_aou
ts_aou_eol_decomp <- decompose(ts_seadec_aou, type = "additive")

autoplot(ts_aou_eol_decomp) +
  xlab(' ')

#remove seasonality :
ts_aou_eol_less_season <- ts_aou_eol_decomp$x - ts_aou_eol_decomp$seasonal

plot_ts_aou_eol_less_season <- autoplot(ts_aou_eol_less_season) +
  xlab("") + ylab("aou") +
  ggtitle("AOU time series at EOL (without seasonality)") +
  geom_smooth(method=lm)

#regression values :
reg_val_ts_aou_eol_less_season <- as.data.frame(plot_ts_aou_eol_less_season$data)

model_ts_aou_less_season <- lm(formula=y~x, data=reg_val_ts_aou_eol_less_season)
summary(model_ts_aou_less_season)


#plot with slope value :
plot_ts_aou_eol_less_season <- autoplot(ts_aou_eol_less_season) +
  xlab("") + ylab("aou (°C)") +
  ggtitle("AOU time series at EOL (without seasonality)") +
  geom_smooth(method=lm) +
  annotate(geom="text", x=1200, y=150, label="slope :  -0.0079 ± 0.00018 (significative, P <2e-16)", color="black")

plot_ts_aou_eol_less_season
```

Slope is `r round(model_ts_aou_less_season$coefficients[[2]],3)`. Means that periode of 5 years is not suitable. Slope not reliable.

## Mann-Kendall

- The function kendall.SeasonalTrendTest returns estimated values of Kendall's τ, the slope, and the intercept for each season, as well as a single estimate for each of these three quantities combined over all seasons. The overall estimate of τ is the weighted average of the p seasonal τ's

- The overall estimate of slope is the median of all two-point slopes computed within each season.

- The overall estimate of intercept is the median of the p seasonal estimates of intercept.

--------

- The Kendall Tau, or Kendall rank correlation coefficient, measures the monotony of the slope. Kendall's Tau varies between -1 and 1; it is positive when the trend increases and negative when the trend decreases.

- The Sen slope, which estimates the overall slope of the time series. This slope corresponds to the median of all the slopes calculated between each pair of points in the series.

- The significance, which represents the threshold for which the hypothesis that there is no trend is accepted. The trend is statistically significant when the p-value is less than 0.05.

Results:

### Temperature

```{r mann kendall EOL temp, echo=FALSE, warning=FALSE, message=FALSE}
FFtemp_eol <- kendallSeasonalTrendTest(temperature ~ month(datetime) + year(datetime), data = EOL_RAW)
FFtemp_eol$estimate
```

### Salinity

```{r mann kendall EOL sal, echo=FALSE, warning=FALSE, message=FALSE}

FFsal_eol <- kendallSeasonalTrendTest(salinity ~ month(datetime) + year(datetime), data = EOL_RAW)
FFsal_eol$estimate
```

### Oxygen

```{r mann kendall EOL oxy, echo=FALSE, warning=FALSE, message=FALSE}

FFoxy_eol <- kendallSeasonalTrendTest(oxygen ~ month(datetime) + year(datetime), data = EOL_RAW)
FFoxy_eol$estimate
```

### AOU

```{r mann kendall EOL aou, echo=FALSE, warning=FALSE, message=FALSE}

FFaou_eol <- kendallSeasonalTrendTest(AOU ~ month(datetime) + year(datetime), data = EOL_RAW)
FFaou_eol$estimate
```

Results from Mann-Kendall and anomalies (by day) are similars.
