---
title: Atmospheric pCO~2~
author: |
   | M&eacute;gane Ballin
   
date: '`r format(Sys.time(), "%d %B %Y %H:%M")`'

output:
    rmdformats::robobook:
     self_contained: true
     thumbnails: false
     lightbox: true
     gallery: true
     highlight: pygments
---

```{r libraries, echo=FALSE, warning=FALSE, message=FALSE}
library("readr")
library("viridis")
library("scales")
library("lubridate")
library("gsignal")
library("FactoMineR")
library("factoextra")
library("gplots")
library("heatwaveR")
library("dplyr")
library("imputeTS")
library("Metrics")
library("tidyr")
library("tseries")
library("forecast")
library("tsDyn")
library("tsoutliers")
library("seacarb")
library("lmodel2")
library("lmtest")
library("ggExtra")
library("corrplot")
library("ggcorrplot")
library("marelac")
library("kableExtra")
library("stringr")
library("berryFunctions")
```

# Description

→ 5 sites\
→ Frequency : hourly\
→ Converts xCO2 (mole fraction of CO2) into pCO2 (partial pressure of CO2)   
  
1 : Monte Cimone (8m) - 2011-2023\
2 : Plateau Rosa (10m) - 2007-2023\
3 : Corse Ersa (40m) - 2013-2023\
4 : Observatoire Haute Provence (10m) - 2014-2023\
5 : Lampedusa (10) - 2006-2023

```{r import data pco2 air, echo=FALSE, warning=FALSE, message=FALSE}

# unit (umol/mol, ppm)

# MONTE CIMONE
atmospheric_co2_MONTE_CIMONE <- read_delim("../../Data/atmospheric_pco2_icos/atmospheric_co2_MONTE_CIMONE_2011_2023.CO2", 
   delim = ";", escape_double = FALSE, trim_ws = TRUE)

#selection of 09:00 AM data :
atmospheric_co2_MONTE_CIMONE <- atmospheric_co2_MONTE_CIMONE %>% 
 unite(col='datetime', c('Year', 'Month', 'Day'), sep='-') %>%
 unite(col='time', c('Hour', 'Minute'), sep=':') %>% 
 dplyr::filter(time == "09:00")

#data sorting :
atmospheric_co2_MONTE_CIMONE <- atmospheric_co2_MONTE_CIMONE %>% 
 mutate(datetime = format(datetime, format="%Y-%m-%d")) %>% 
 mutate(datetime = as.POSIXct(datetime, format="%Y-%m-%d")) %>%
 rename(CO2_MC = co2) %>% 
 mutate(CO2_MC = case_when(CO2_MC == -999.990 ~ NA_real_ , TRUE ~ CO2_MC)) %>%
 select(datetime, CO2_MC)
 
####################

# PLATEAU ROSA
atmospheric_co2_PLATEAU_ROSA <- read_delim("../../Data/atmospheric_pco2_icos/atmospheric_co2_PLATEAU_ROSA_2007_2023.CO2", 
   delim = ";", escape_double = FALSE, trim_ws = TRUE)
 
#selection of 09:00 AM data :
atmospheric_co2_PLATEAU_ROSA <- atmospheric_co2_PLATEAU_ROSA %>% 
 unite(col='datetime', c('Year', 'Month', 'Day'), sep='-') %>% 
 unite(col='time', c('Hour', 'Minute'), sep=':') %>% 
 dplyr::filter(time == "09:00")
 
#data sorting :
atmospheric_co2_PLATEAU_ROSA <- atmospheric_co2_PLATEAU_ROSA %>% 
 mutate(datetime = format(datetime, format="%Y-%m-%d")) %>% 
 mutate(datetime = as.POSIXct(datetime, format="%Y-%m-%d")) %>%
 rename(CO2_PR = co2) %>% 
 mutate(CO2_PR = case_when(CO2_PR == -999.990 ~ NA_real_ , TRUE ~ CO2_PR)) %>% 
 select(datetime, CO2_PR)
 
 ####################
 
# CORSE ERSA
atmospheric_co2_CORSE <- read_delim("../../Data/atmospheric_pco2_icos/atmospheric_co2_CORSE_2013_2023.CO2",
    delim = ";", escape_double = FALSE, trim_ws = TRUE)

#selection of 09:00 AM data :
atmospheric_co2_CORSE <- atmospheric_co2_CORSE %>%
  unite(col='datetime', c('Year', 'Month', 'Day'), sep='-') %>%
  unite(col='time', c('Hour', 'Minute'), sep=':') %>%
  dplyr::filter(time == "09:00")

#data sorting :
atmospheric_co2_CORSE <- atmospheric_co2_CORSE %>%
  mutate(datetime = format(datetime, format="%Y-%m-%d")) %>%
  mutate(datetime = as.POSIXct(datetime, format="%Y-%m-%d")) %>%
  rename(CO2_CORSE = co2) %>%
  mutate(CO2_CORSE = case_when(CO2_CORSE == -999.990 ~ NA_real_ , TRUE ~ CO2_CORSE)) %>%
  select(datetime, CO2_CORSE)

####################

# OBSERVATOIRE HAUTE PROVENCE
atmospheric_co2_HAUTE_PROV <- read_delim("../../Data/atmospheric_pco2_icos/atmospheric_co2_HAUTE_PROVENCE_2014_2023.CO2",
   delim = ";", escape_double = FALSE, trim_ws = TRUE)

#selection of 09:00 AM data :
atmospheric_co2_HAUTE_PROV <- atmospheric_co2_HAUTE_PROV %>%
 unite(col='datetime', c('Year', 'Month', 'Day'), sep='-') %>%
 unite(col='time', c('Hour', 'Minute'), sep=':') %>%
 dplyr::filter(time == "09:00")

#data sorting :
atmospheric_co2_HAUTE_PROV <- atmospheric_co2_HAUTE_PROV %>%
 mutate(datetime = format(datetime, format="%Y-%m-%d")) %>%
 mutate(datetime = as.POSIXct(datetime, format="%Y-%m-%d")) %>%
 rename(CO2_HP = co2) %>%
 mutate(CO2_HP = case_when(CO2_HP == -999.990 ~ NA_real_ , TRUE ~ CO2_HP)) %>%
 select(datetime, CO2_HP)

####################

# LAMPEDUSA
atmospheric_co2_LAMPEDUSA <- read_delim("../../Data/atmospheric_pco2_icos/atmospheric_co2_LAMPEDUSA_2006_2023.CO2",
   delim = ";", escape_double = FALSE, trim_ws = TRUE)

#selection of 09:00 AM data :
atmospheric_co2_LAMPEDUSA <- atmospheric_co2_LAMPEDUSA %>%
 unite(col='datetime', c('Year', 'Month', 'Day'), sep='-') %>%
 unite(col='time', c('Hour', 'Minute'), sep=':') %>%
 dplyr::filter(time == "09:00")

#data sorting :
atmospheric_co2_LAMPEDUSA <- atmospheric_co2_LAMPEDUSA %>%
 mutate(datetime = format(datetime, format="%Y-%m-%d")) %>%
 mutate(datetime = as.POSIXct(datetime, format="%Y-%m-%d")) %>%
 rename(CO2_LAMP = co2) %>%
 mutate(CO2_LAMP = case_when(CO2_LAMP == -999.990 ~ NA_real_ , TRUE ~ CO2_LAMP)) %>%
 select(datetime, CO2_LAMP)
```

# Summary 5 sites (µatm)

```{r summary 5 sites atmospheric pCO2 NA values, echo=FALSE, warning=FALSE, message=FALSE}

#merge of 5 sites data : pCO2_air
 pCO2_atmos <- full_join(atmospheric_co2_LAMPEDUSA, atmospheric_co2_MONTE_CIMONE, by="datetime")
 pCO2_atmos <- full_join(pCO2_atmos, atmospheric_co2_PLATEAU_ROSA, by="datetime")
 pCO2_atmos <- full_join(pCO2_atmos, atmospheric_co2_CORSE, by="datetime")
 pCO2_atmos <- full_join(pCO2_atmos, atmospheric_co2_HAUTE_PROV, by="datetime")

#conversion dry mol fraction of CO2 into partial pressure of CO2(µatm) :
pCO2_atmos <- pCO2_atmos %>%
   mutate(pCO2_LAMP = x2pCO2(S=35, T=25, Patm=1.0, xCO2=CO2_LAMP),
          pCO2_MC = x2pCO2(S=35, T=25, Patm=1.0, xCO2=CO2_MC),
          pCO2_PR = x2pCO2(S=35, T=25, Patm=1.0, xCO2=CO2_PR),
          pCO2_CORSE = x2pCO2(S=35, T=25, Patm=1.0, xCO2=CO2_CORSE),
          pCO2_HP = x2pCO2(S=35, T=25, Patm=1.0, xCO2=CO2_HP))

#summary :
summary(pCO2_atmos[,c(7:11)])

#save :
write.table(pCO2_atmos, "../../Data/atmospheric_pco2_icos/DATA_pCO2_atmos_06-23.csv", sep=";", col.names = TRUE, row.names = FALSE)

```
  
  
# Visualisation of atmospheric pCO~2~ (µatm) in the 5 sites
  

```{r Visualisation 5 sites atmospheric pCO2, echo=FALSE, warning=FALSE, message=FALSE, fig.width=9,fig.height=4}

pCO2_atmos <- read_delim("../../Data/atmospheric_pco2_icos/DATA_pCO2_atmos_06-23.csv", 
    delim = ";", escape_double = FALSE, trim_ws = TRUE)


#creation column for legend
pCO2_atmos_long <- pCO2_atmos %>% 
  rename(Corsica = pCO2_CORSE, `Haute Provence` = pCO2_HP, Lampedusa = pCO2_LAMP, `Monte Cimone` = pCO2_MC, 
         `Plateau Rosa` = pCO2_PR) %>% 
  pivot_longer(cols=c(Corsica, `Haute Provence`, Lampedusa, `Monte Cimone`, `Plateau Rosa`), 
               names_to = "Legend", values_to = "value")

#plot
pCO2_atmos_long %>% 
  ggplot() +
  ggtitle("Atmospheric pCO2 of the 5 sites according to time") +
  geom_point(aes(x=datetime, y=value, color= Legend), size=0.5) +
  scale_y_continuous(name="pCO2 (µatm)") + 
  scale_x_date(name="", date_breaks="2 year", date_minor_breaks="1 years", date_labels = "%Y") +
  scale_colour_discrete(type=c("#0000FF", "#568203", "#048B9A", "#E9383F", "#C4698F"))


```
  
  
# Comparison of the 5 sites 

## Corsica vs Haute Provence

*Haute Provence = land site (variations explained by other variables like vegetation)*


```{r Corsica vs Haute Provence , echo=FALSE, warning=FALSE, message=FALSE, fig.width=8,fig.height=4,fig.align='center', fig.show='hold'}

#creation column for legend
corse_HP <- pCO2_atmos %>%
  select(datetime, pCO2_CORSE, pCO2_HP)

corse_HP_long <- corse_HP %>%
  rename(Corsica = pCO2_CORSE, `Haute Provence` = pCO2_HP) %>%
  pivot_longer(cols=Corsica :`Haute Provence`, names_to = "Legend", values_to = "value")

#plot
#corse_HP_long[4798: 11260,] %>%
# ggplot() +
#  ggtitle("Atmospheric pCO2 of Corsica vs Haute Provence according to time") +
#  geom_point(aes(x=datetime, y=value, color= Legend), size=0.5) +
#  scale_y_continuous(name="pCO2 (µatm)") +
#  scale_x_datetime(name="") +
#  scale_colour_discrete(type=c("#0000FF", "#568203"))
##

#scatter plot :
corse_HP[4798: 11260,] %>%
  ggplot() +
  aes(x=pCO2_HP, y=pCO2_CORSE) +
  geom_point(size=0.85) +
  geom_smooth(method="lm", formula = y ~ x) +
  scale_x_continuous(name="Haute Provence Observatory") +
  scale_y_continuous(name="Corsica") +
  ggtitle("Relationship between Corsica and HP observatory sites") +
  annotate(geom="text", x=420, y=430, label="slope : 0.79 ± 0.009 *  R-adj : 0.78", color="black")

#slope + intercept

reg_corse_HP <- lm(data = corse_HP, pCO2_CORSE ~ pCO2_HP)
summary(reg_corse_HP)

```



## Corsica vs Plateau Rosa

*Better fit*
*Necessity to take Plateau Rosa data to fill in the series before 2013 and after 2020*


```{r Corsica vs Plateau Rosa, echo=FALSE, warning=FALSE, message=FALSE, fig.width=8,fig.height=4,fig.align='center', fig.show='hold'}

#creation column for legend
corse_PR <- pCO2_atmos %>%
  select(datetime, pCO2_CORSE, pCO2_PR)

corse_PR_long <- corse_PR %>%
  rename(Corsica = pCO2_CORSE, `Plateau Rosa` = pCO2_PR) %>%
  pivot_longer(cols=Corsica :`Plateau Rosa`, names_to = "Legend", values_to = "value")

#plot
#corse_PR_long %>%
#  ggplot() +
#  ggtitle("Atmospheric pCO2 of Corsica vs Plateau Rosa according to time") +
#  geom_point(aes(x=datetime, y=value, color= Legend), size=0.5) +
#  scale_y_continuous(name="pCO2 (µatm)") +
#  scale_x_datetime(name="", date_breaks="2 year", date_minor_breaks="1 years", date_labels = "%Y") +
#  scale_colour_discrete(type=c("#0000FF", "#C4698F"))
##

#scatter plot :
corse_PR %>%
  ggplot() +
  aes(x=pCO2_PR, y=pCO2_CORSE) +
  geom_point(size=0.85) +
  stat_smooth(method="lm", formula = y ~ x) +
  scale_x_continuous(name="Plateau Rosa") +
  scale_y_continuous(name="Corsica") +
  ggtitle("Relationship between Corsica and Plateau Rosa sites") +
  annotate(geom="text", x=390, y=420, label="slope : 0.99 ± 0.009 *  R-adj : 0.87", color="black")

#slope + intercept

reg_corse_PR <- lm(data = corse_PR, pCO2_CORSE ~ pCO2_PR)
summary(reg_corse_PR)

```


## Corsica vs Lampedusa

*Good fit*
*Necessity to take Lampedusa data to fill in the series before 2013 and after 2020*


```{r Corsica vs Lampedusa, echo=FALSE, warning=FALSE, message=FALSE, fig.width=8,fig.height=4,fig.align='center', fig.show='hold'}

#creation column for legend
corse_LAMP <- pCO2_atmos %>%
  select(datetime, pCO2_CORSE, pCO2_LAMP)

corse_LAMP_long <- corse_LAMP %>%
  rename(Corsica = pCO2_CORSE, Lampedusa = pCO2_LAMP) %>%
  pivot_longer(cols=Corsica :Lampedusa, names_to = "Legend", values_to = "value")

#plot
#corse_LAMP_long %>%
#  ggplot() +
#  ggtitle("Atmospheric pCO2 of Corsica vs Lampedusa according to time") +
#  geom_point(aes(x=datetime, y=value, color= Legend), size=0.5) +
#  scale_y_continuous(name="pCO2 (µatm)") +
#  scale_x_datetime(name="", date_breaks="2 year", date_minor_breaks="1 years", date_labels = "%Y") +
#  scale_colour_discrete(type=c("#0000FF", "#048B9A"))
##

#scatter plot :
corse_LAMP %>%
  ggplot() +
  aes(x=pCO2_LAMP, y=pCO2_CORSE) +
  geom_point(size=0.85) +
  stat_smooth(method="lm", formula = y ~ x) +
  scale_x_continuous(name="Lampedusa") +
  scale_y_continuous(name="Corsica") +
  ggtitle("Relationship between Corsica and Lampedusa sites") +
  annotate(geom="text", x=390, y=420, label="slope : 0.99 ± 0.008 *  R-adj : 0.89", color="black")

#slope + intercept

reg_corse_LAMP <- lm(data = corse_LAMP, pCO2_CORSE ~ pCO2_LAMP)
summary(reg_corse_LAMP)

```


## Plateau Rosa vs Lampedusa

*Best fit*
*Necessity to take Lampedusa data to fill in the series before 2013 and after 2020*


```{r Plateau vs Lampedusa, echo=FALSE, warning=FALSE, message=FALSE, fig.width=8,fig.height=4,fig.align='center', fig.show='hold'}

#creation column for legend
PR_LAMP <- pCO2_atmos %>%
  select(datetime, pCO2_PR, pCO2_LAMP)

PR_LAMP_long <- PR_LAMP %>%
  rename(`Plateau Rosa` = pCO2_PR, Lampedusa = pCO2_LAMP) %>%
  pivot_longer(cols=c(`Plateau Rosa`, Lampedusa), names_to = "Legend", values_to = "value")

#plot
#PR_LAMP_long %>%
#  ggplot() +
#  ggtitle("Atmospheric pCO2 of Plateau Rosa vs Lampedusa according to time") +
#  geom_point(aes(x=datetime, y=value, color= Legend), size=0.5) +
#  scale_y_continuous(name="pCO2 (µatm)") +
#  scale_x_datetime(name="", date_breaks="2 year", date_minor_breaks="1 years", date_labels = "%Y") +
#  scale_colour_discrete(type=c("#C4698F", "#048B9A"))
##

#scatter plot :
PR_LAMP %>%
  ggplot() +
  aes(x=pCO2_LAMP, y=pCO2_PR) +
  geom_point(size=0.85) +
  stat_smooth(method="lm", formula = y ~ x) +
  scale_x_continuous(name="Lampedusa") +
  scale_y_continuous(name="Plateau Rosa") +
  ggtitle("Relationship between Plateau Rosa and Lampedusa sites") +
  annotate(geom="text", x=390, y=420, label="slope : 0.98 ± 0.004 *  R-adj : 0.94", color="black")

#slope + intercept

reg_PR_LAMP <- lm(data = PR_LAMP, pCO2_PR ~ pCO2_LAMP)
summary(reg_PR_LAMP)

```

  
  
# NA values  
  
## Lampedusa site (2006-2023)  
  
Number of NA : 1805  
  
```{r LAMPEDUSA NA values, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6,fig.height=4,fig.align='center', fig.show='hold'}

#LAMPEDUSA : 1805 NA
#summary(pCO2_atmos$pCO2_LAMP)

#plot nb NA par années
atmospheric_co2_LAMPEDUSA_NA_plot <- pCO2_atmos %>% 
  dplyr::mutate(Year = format(datetime, format="%Y"),
                Month = format(datetime, format="%m-%d")) %>% 
  dplyr::group_by(Year, Month) 

atmospheric_co2_LAMPEDUSA_NA_plot$pCO2_LAMP[is.na(atmospheric_co2_LAMPEDUSA_NA_plot$pCO2_LAMP)] <- "NA"

atmospheric_co2_LAMPEDUSA_NA_plot %>% 
  dplyr::filter(pCO2_LAMP == "NA") %>% 
  dplyr::group_by(Year) %>% 
  dplyr::summarise(number_of_NA = n()) %>% 
  ggplot(aes(x = Year, y = number_of_NA))  +
  ggtitle("NA occurences by year - Lampedusa site") +
  geom_segment(aes(x = Year, xend = Year, y = 0, yend = number_of_NA), 
               color = "grey", linewidth = 2) +
  geom_point(size = 4, fill = "#048B9A", shape = 21, color = "black") +
  coord_flip() +
  scale_x_discrete(name = "") +
  scale_y_continuous(name = "NA number", limits = c(0,300), breaks = seq(0, 300, 50))

```
  
  
## Plateau Rosa site (2007-2023)  
  
Number of NA : 1354
  
```{r PLATEAU ROSA NA values, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6, fig.height=4, fig.align='center', fig.show='hold'}

#PLATEAU ROSA : 1354 NA
#summary(pCO2_atmos$pCO2_PR)

#plot nb NA par années
atmospheric_co2_PLATEAU_ROSA_NA_plot <- pCO2_atmos %>% 
  dplyr::mutate(Year = format(datetime, format="%Y"),
                Month = format(datetime, format="%m-%d")) %>% 
  dplyr::group_by(Year, Month) 

atmospheric_co2_PLATEAU_ROSA_NA_plot$pCO2_PR[is.na(atmospheric_co2_PLATEAU_ROSA_NA_plot$pCO2_PR)] <- "NA"

atmospheric_co2_PLATEAU_ROSA_NA_plot %>% 
  dplyr::filter(pCO2_PR == "NA") %>% 
  dplyr::group_by(Year) %>% 
  dplyr::summarise(number_of_NA = n()) %>% 
  ggplot(aes(x = Year, y = number_of_NA))  +
  ggtitle("NA occurences by year - Plateau Rosa site") +
  geom_segment(aes(x = Year, xend = Year, y = 0, yend = number_of_NA), 
               color = "grey", linewidth = 2) +
  geom_point(size = 4, fill = "#C4698F", shape = 21, color = "black") +
  coord_flip() +
  scale_x_discrete(name = "") +
  scale_y_continuous(name = "NA number", limits = c(0,200), breaks = seq(0, 200, 50))

```
  
  
## Corsica site (2013-2023)

Number of NA : 3398

```{r visualisation CORSICA NA values, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6,fig.height=4,fig.align='center', fig.show='hold'}

#CORSICA : 3398 NA
#summary(pCO2_atmos$pCO2_CORSE)

#plot nb NA par années
atmospheric_co2_CORSICA_NA_plot <- pCO2_atmos %>% 
  dplyr::mutate(Year = format(datetime, format="%Y"),
                Month = format(datetime, format="%m-%d")) %>% 
  dplyr::group_by(Year, Month) 

atmospheric_co2_CORSICA_NA_plot$pCO2_CORSE[is.na(atmospheric_co2_CORSICA_NA_plot$pCO2_CORSE)] <- "NA"

atmospheric_co2_CORSICA_NA_plot %>% 
  dplyr::filter(pCO2_CORSE == "NA") %>% 
  dplyr::group_by(Year) %>% 
  dplyr::summarise(number_of_NA = n()) %>% 
  ggplot(aes(x = Year, y = number_of_NA))  +
  ggtitle("NA occurences by year - Corsica site") +
  geom_segment(aes(x = Year, xend = Year, y = 0, yend = number_of_NA), 
               color = "grey", linewidth = 2) +
  geom_point(size = 4, fill = "#0000FF", shape = 21, color = "black") +
  coord_flip() +
  scale_x_discrete(name = "") +
  scale_y_continuous(name = "NA number", limits = c(0,400), breaks = seq(0, 400, 50))

```
  
  

