---
title: Air-Sea CO~2~ fluxes
author: |
   | M&eacute;gane Ballin
   
date: '`r format(Sys.time(), "%d %B %Y %H:%M")`'

output:
    rmdformats::robobook:
     self_contained: true
     thumbnails: false
     lightbox: true
     gallery: true
     highlight: pygments
---

```{r libraries, echo=FALSE, warning=FALSE, message=FALSE}
library("readr")
library("viridis")
library("scales")
library("lubridate")
library("gsignal")
library("FactoMineR")
library("factoextra")
library("gplots")
library("heatwaveR")
library("dplyr")
library("imputeTS")
library("Metrics")
library("tidyr")
library("tseries")
library("forecast")
library("tsDyn")
library("tsoutliers")
library("seacarb")
library("lmodel2")
library("lmtest")
library("ggExtra")
library("corrplot")
library("ggcorrplot")
library("marelac")
library("kableExtra")
library("stringr")
library("berryFunctions")
```

# Description 

Data wind speed 10m (m/s):  
  
→ Azur buoy (model) - 2009-2022  
→ EOL buoy (model) - 2009-2022  
  
Frequency : Hourly  
Unit : m/s

```{r Importation data wind speed, echo=FALSE, warning=FALSE, message=FALSE}

#importation data wind speed (model) of Azur buoy : dec.2008 - dec.2022) :
Wind_Azur <- read_csv("../../Data/Wind_speed_10m_Azur_buoy/model_wind_speed_AZUR.csv") 

#delete december 2008 :
Wind_Azur <- Wind_Azur[319:121962,] #jan.2009 to dec.2022

##

#importation data wind speed (model) of EOL buoy (fin 2011-2022) :
Wind_EOL <- read_csv("../../Data/wind_cap_ferrat_model_EOL/model_wind_speed_EOL.csv") %>% arrange(date)

#delete december 2008 :
Wind_EOL <- Wind_EOL[319:121962,] #jan.2009 to dec.2022


```


# Histogram of wind speed distribution : pelagic (Azur buoy) vs coastal (EOL buoy)  

```{r Histogram of wind speed distribution, echo=FALSE, warning=FALSE, message=FALSE, fig.width=5, fig.height=3, fig.fullwidth=TRUE}

#plot 1
 Wind_Azur %>% 
   ggplot(aes(x=speed_ms)) +
   geom_histogram(fill="#69b3a2", color="#e9ecef", alpha=0.8) +
   ggtitle("Wind speed distribution (Azur buoy)") + 
   scale_y_continuous(name="Distribution") +
   scale_x_continuous(name="", breaks = c(0, 5, 10, 15, 20))
##
 
#plot 2
 Wind_EOL %>% 
   ggplot(aes(x=speed_ms)) +
   geom_histogram(fill="#ED7F10", color="#e9ecef", alpha=0.8) +
   ggtitle("Wind speed distribution (EOL buoy)") + 
   scale_y_continuous(name="Distribution") +
   scale_x_continuous(name="", breaks = c(0, 5, 10, 15, 20))


```
  
  
# Annual means of wind speed (2009-2022) : EOL buoy  
  

```{r Wind graph, echo=FALSE, warning=FALSE, message=FALSE}

Wind_EOL <- Wind_EOL %>% 
  mutate(Hour = format(date, format="%H:%M:%S")) %>% 
  dplyr::filter(Hour == "09:00:00") %>% 
  mutate(datetime = format(date, format="%Y-%m-%d"),
         datetime = as.POSIXct(datetime)) %>% 
dplyr::select(datetime, speed_ms)


graph_wind_eol <- Wind_EOL %>% 
  mutate(year = format(datetime, format="%Y")) %>% 
  group_by(year) %>% 
  mutate(mean = mean(speed_ms))

wind_mean <- graph_wind_eol %>% distinct(mean)

wind_mean %>%  
  ggplot() + 
  ggtitle("") +
  aes(x=year, y=mean) +
  geom_point(col="black", size=0.7) +
  geom_line(group=1) +
  scale_y_continuous(name=expression(paste("Moyennes annuelles (m ", s^-1,")"))) +
  scale_x_discrete(name="") 

# wind_mean %>%  
#   ggplot() + 
#   ggtitle("Annual means of wind speed (EOL)") +
#   aes(x=year, y=mean) +
#   geom_col(col="black", fill="#357AB7", size=0.7) +
#   geom_line() +
#   scale_y_continuous(name="Speed (m/s)") +
#   scale_x_discrete(name="") 

```
  
  
# Data *p*CO~2~ water Point B (µatm):  
  
period : jan.2007 - dec.2022 
  
Surface data,  
Frequency : Weekly (every tuesday at 09 AM)  
Unity : µatm  
  
  
```{r importation data pCO2 water, echo=FALSE, warning=FALSE, message=FALSE}

#period : jan. 2007 - dec. 2022
#measurements every tuesday at 09h AM

SOMLIT_pcO2 <- read_delim("../../Data/data_PointB/DATA_SOMLIT_07-22_impute_0m.csv", 
    delim = ";", escape_double = FALSE, trim_ws = TRUE)

#delta pCO2 = pCO2 seawater - pCO2 air 
```

  

# Annual cycle of seawater *p*CO~2~ at Point B (2007-2022) :  
  

```{r annual cycle of pCO2 (2007-2021), echo=FALSE, warning=FALSE, message=FALSE, fig.width=7,fig.height=5}

#Jan.2007 - Sept.2022 :
annual_cycle_pco2Plot <- SOMLIT_pcO2 %>% 
  ggplot() +
  geom_line(aes(x= as.Date(yday(datetime), "1970-01-01"), y=impute_pCO2_w, 
                group = factor(year(datetime)), 
                color = factor(year(datetime))), linewidth = 0.6) +
  scale_colour_viridis_d(option="mako", direction=-1) +
  ggtitle("") +
  scale_x_date(date_breaks="months", date_labels="%b", name = "") +
  labs(x="Months",colour="") +
  theme_bw() +
  scale_y_continuous(name = "pCO2 (µatm)") 


#save
 # ggsave(filename = "annual_cycle_pco2Plot.jpg", # Nommez le fichier dans lequel vous voulez enregistrer, ajoutez l'extension du format de fichier que vous voulez utiliser (ex. pdf).
 #        plot = annual_cycle_pco2Plot, # Fournir le nom de l'objet plot dans R
 #        height = 5, # Fournir les dimensions voulues
 #        width = 8, 
 #        units = "in") 



##polar plot :
ts_pco2w <- ts(SOMLIT_pcO2$impute_pCO2_w, start = c(2007,1), end = c(2022,39), freq = 52)

polar_plot_pco2w <- ggseasonplot(ts_pco2w, polar = TRUE, main="", xlab="", ylab = "",
             season.labels = c("Jan","Feb","Mar","Apr","May","Jun","Jul",
                               "Aug","Sep","Oct","Nov","Dec")) +
  scale_colour_discrete(name="Années", type=c("#80D0D0", "#83A697", "#97DFC6", "#B0F2B6",  "#F4FEFE", "#FFF0BC", 
                               "#F7E269", "#EFD807", "#DD985C", "#ED7F10", "#CC5500", "#FF866A", 
                               "#E9383F", "#C60800", "#850606", "#6E0B14"))

annual_cycle_pco2Plot
polar_plot_pco2w
```
  
  
# Calculation of N*p*CO~2~ & T*p*CO~2~ and deltas :  
  
NpCO~2~ = non-temperature effects on pCO~2~ water variations  
  
TpCO~2~ = only temperature effects on pCO~2~ water variations  
  
  
→ Calculation of deltas :  
  
-   delta pCO~2~(T) = temperature effects = pCO~2~ water - NpCO~2~  

-   delta pCO~2~(bio) : biological effects = pCO~2~ water - TpCO~2~  
  

```{r NpCO2 & TpCO2 & deltas calculations, echo=FALSE, warning=FALSE, message=FALSE, fig.width=8,fig.height=5}

#formula NpCO2 : pCO2w obs * e(0.0423*Tmean-Tobs)
#formula TpCO2 : pCO2mean * exp[0.0423(Tobs - Tmean)]

pco2_bio <- read_delim("../../Data/data_PointB/DATA_SOMLIT_07-22_impute_0m.csv", 
    delim = ";", escape_double = FALSE, trim_ws = TRUE)

pco2_bio <- pco2_bio %>% 
  mutate(Tmean = mean(impute_temp),
         pCO2_w_mean = mean(impute_pCO2_w, na.rm = TRUE),
         NpCO2 = impute_pCO2_w * exp(0.0423*(Tmean-impute_temp)),
         TpCO2 = pCO2_w_mean * exp(0.0423*(impute_temp-Tmean)))
##

#deltas calculation :
pco2_bio <- pco2_bio %>%
  mutate(delta_pCO2_T = (impute_pCO2_w - NpCO2),
         delta_pCO2_bio = (impute_pCO2_w - TpCO2))

##

#plot deltas :
pCO2bio_pCO2T_plot_long <- pco2_bio %>% 
  rename(`DpCO2 (T)` = delta_pCO2_T, `DpCO2 (bio)` = delta_pCO2_bio) %>%
  pivot_longer(cols = c(`DpCO2 (T)`, `DpCO2 (bio)`), names_to = "Legend", values_to = "value")

pCO2bio_pCO2T_plot <- pCO2bio_pCO2T_plot_long %>% 
  ggplot() +
  ggtitle("") +
  geom_line(aes(x=datetime, y=value, color= Legend), size=0.5) +
  labs(colour="Legend :") +
  theme(legend.position = c(0.025, 1), legend.justification = c("left", "top"),
        legend.box.just = "left", legend.margin = margin(6, 6, 6, 6)) +
  scale_y_continuous(name="ΔpCO2 (µatm)") + 
  scale_x_date(name="", date_breaks="2 year", date_minor_breaks="1 years", labels = date_format("%Y")) + 
  scale_colour_discrete(type=c("#096A09", "#E9383F"))

#pCO2bio_pCO2T_plot
##

##RAPPORT##
pco2_bio_plot_rapport <- pco2_bio %>% 
  ggplot() +
  geom_line(aes(x=datetime, y=delta_pCO2_T), col = "#BF3030", size = 0.5) +
  geom_line(aes(x=datetime, y=delta_pCO2_bio), col = "#096A09", size = 0.5) +
  scale_y_continuous(name=expression(paste(Delta, italic(p), CO[2], " (", mu, "atm)"))) + 
  scale_x_date(name="", date_breaks="2 year", date_minor_breaks="1 years", labels = date_format("%Y")) +
  geom_hline(col = "black", yintercept = 0, linetype="dashed")
#  theme(panel.background = element_rect(fill = "#CCCCCC", size = 0.3, linetype="solid"))

pco2_bio_plot_rapport

#save
 # ggsave(filename = "pCO2bio_pCO2T_plot.jpg", # Nommez le fichier dans lequel vous voulez enregistrer, ajoutez l'extension du format de fichier que vous voulez utiliser (ex. pdf).
 #         plot = pco2_bio_plot_rapport, # Fournir le nom de l'objet plot dans R
 #         height = 6, # Fournir les dimensions voulues
 #         width = 9, 
 #         units = "in") 
 

```
  
  

# Atmospheric and seawater *p*CO~2~ (µatm) according to time (2007-2022) :  
  

```{r atmospheric & water pCO2 plot, echo=FALSE, warning=FALSE, message=FALSE, fig.width=8,fig.height=6}

atmos_water_pCO2_plot_long <- pco2_bio %>% 
  rename(`pCO2 ocean` = impute_pCO2_w, `pCO2 atmosphere` = impute_pCO2_atmos, NpCO2 = NpCO2) %>% 
  pivot_longer(cols = c(`pCO2 ocean`, `pCO2 atmosphere`, "NpCO2"), names_to = "Legend", values_to = "value")

#plot
atmos_water_pCO2_plot <- atmos_water_pCO2_plot_long %>% 
  ggplot() +
  ggtitle("") +
  geom_line(aes(x=datetime, y=value, color= Legend), size=0.5) +
  scale_y_continuous(name=expression(paste(italic(p), CO[2], " (", mu, "atm)"))) + 
  scale_x_date(name="", date_breaks="2 year", date_minor_breaks="1 years", labels = date_format("%Y")) + 
  scale_colour_discrete(type=c("#708D23", "#CC5500", "#56739A"))

atmos_water_pCO2_plot
##

#save
 # ggsave(filename = "atmos_water_pCO2_plot.jpg", # Nommez le fichier dans lequel vous voulez enregistrer, ajoutez l'extension du format de fichier que vous voulez utiliser (ex. pdf).
 #        plot = atmos_water_pCO2_plot, # Fournir le nom de l'objet plot dans R
 #        height = 6, # Fournir les dimensions voulues
 #        width = 10, 
 #        units = "in") 

#save merge3 :
#write.table(pco2_bio, "DATA_SOMLIT_07-22_impute_0m_deltas_pCO2.csv", sep=";", col.names = TRUE, row.names = FALSE)

```
  
  

# Air-sea CO~2~ fluxes  
  
  
## Flux calculation  
  
→ **negative values = flux into the ocean**  
  
  
Method used : Wanninkhof (2014)  
  

```{r calcul flux, echo=FALSE, warning=FALSE, message=FALSE, fig.width=8,fig.height=5}

tab_flux <- read_delim("../../Data/data_PointB/DATA_SOMLIT_07-22_impute_0m_deltas_pCO2.csv", 
    delim = ";", escape_double = FALSE, trim_ws = TRUE)


#Calcul of F
tab_flux <- tab_flux %>% dplyr::filter(datetime >= "2009-01-06") %>% 
    mutate(Flux_CO2_2014 = (0.00077 * speed_ms^2 * delta_pCO2)/ 365, #per day
           Flux_CO2_2014 = Flux_CO2_2014/24) #to have a flux per hour


#plot Flux Wanninkhof 2014 :
tab_flux %>% 
  ggplot() + 
  aes(x=datetime, y=Flux_CO2_2014) + 
  geom_line(col="#967C5C") +
  scale_x_date(name="", date_breaks = "2 year", date_minor_breaks = "1 year", date_labels = "%Y") +
  scale_y_continuous(name="Fluxes (mol of CO2/m²/y)", labels = label_scientific(digits=1), n.breaks=8) +
  ggtitle("CO2 fluxes ocean-atmosphere (Wanninkhof 2014)") 


# tab_flux %>% 
#   ggplot() +
#   aes(x=datetime, y=impute_pCO2_atmos) +
#   geom_line() +
#   scale_x_date(name="", date_breaks = "2 years", date_minor_breaks = "1 year", date_label="%Y") +
#   scale_y_continuous(name=expression(paste(italic(p), CO[2], " atmosphérique (μatm)"))) +
#   geom_smooth(formula = y~x, col="steelblue")


```
  
  
Method used : Wanninkhof (1992)  
  
   
Formulas :  
  
-   K600 = 0.266 \* U²  
  
-   SC = 2073.1 - (125.62 \* T) + (3.6276 \* T²) - (0.043219 \* T\^3)  
  
-   k = K600 \* (SC/600)\^0.5 (cm/h)  
  
-   K0 = Henry's constant (seacarb)  
  
  
Final formula : F = k \* K0 \* ΔpCO2  
  
unity : mol of C / m² / year  
  
```{r calcul Flux Wanninkhof (1992), echo=FALSE, warning=FALSE, message=FALSE, fig.width=10,fig.height=6}

tab_flux <- tab_flux %>% 
  mutate(SC = 2073.1 - (125.62 * impute_temp) + (3.6276 * impute_temp^2) - (0.043219 * impute_temp^3),
         K600 = 0.266 * speed_ms^2,
         k = K600 * (SC/600)^0.5,
         rho = rho(S=impute_sal, T=impute_temp, P=0)/1000,
         K0 = (K0(S=impute_sal, T=impute_temp, P=0, Patm=1, warn="n")),
         K0 = (K0 * rho)/1000000, 
         Flux_wan_92 = (k * K0 * delta_pCO2)*10) # *24*365 to have in mol/m²/y
##


##unity steps :

#K0 en mol/kg/atm
#rho (density of water) en kg/m3
#donc rho/1000 = kg/L
#donc K0*(rho/1000) = mol/l/atm

#K0/10^6 pour passer de mol/L/atm a mol/L/µatm

#Flux*10 pour passer de 10 mol/m²/h a mol/m²/h
#Flux*24*365 pour passer de mol/m²/h a mol/m²/y


#plot Flux wanninkhof 1992 :
tab_flux %>% 
  ggplot() + 
  aes(x=datetime, y=Flux_wan_92) + 
  geom_line(col="#677179") +
  scale_x_date(name="", date_breaks = "2 year", date_minor_breaks = "1 year", date_labels = "%Y") +
  scale_y_continuous(name="Fluxes (mol CO2/m²/h)", labels = label_scientific(digits=1), n.breaks=8) +
  ggtitle("CO2 fluxes ocean-atmosphere (Wanninkhof 1992)")


```
  
  

## Plot comparing the 2 CO~2~ fluxes :  
  

```{r 2 methods comparison - plot, echo=FALSE, warning=FALSE, message=FALSE, fig.width=10,fig.height=6}

#creation column for legend
tab_flux_long <- tab_flux %>% 
  rename(`Flux W (1992)` = Flux_wan_92, `Flux W (2014)` = Flux_CO2_2014) %>% 
  pivot_longer(cols=c(`Flux W (1992)`, `Flux W (2014)`), names_to = "Legend", values_to = "value")

#plot
fluxes_co2_plot <- tab_flux_long %>% 
  ggplot() +
  ggtitle("") +
  geom_line(aes(x=datetime, y=value, color= Legend), size=0.5) +
  scale_y_continuous(name=expression(paste("Flux (mol ", m^-2, " ", h^-1, ")")), labels = label_scientific(digits=1), n.breaks=8) + 
  scale_x_date(name="", date_breaks = "2 year", date_minor_breaks = "1 year", date_labels = "%Y") +
  scale_colour_discrete(type=c("#BF3030", "#677179")) + 
  geom_hline(col = "black", yintercept = 0, linetype="dashed")

fluxes_co2_plot


#save
 # ggsave(filename = "flux_CO2_plot.jpg", # Nommez le fichier dans lequel vous voulez enregistrer, ajoutez l'extension du format de fichier que vous voulez utiliser (ex. pdf).
 #        plot = fluxes_co2_plot, # Fournir le nom de l'objet plot dans R
 #        height = 6, # Fournir les dimensions voulues
 #        width = 10, 
 #        units = "in")


```
  
  
## Annual fluxes (Wanninkhof, 2014)  
  

```{r flux annuels, echo=FALSE, warning=FALSE, message=FALSE, fig.width=8,fig.height=5}

#calculation of annual means of flux in hour :
# annual_fluxes <- tab_flux %>% 
#   group_by(Year) %>% 
#   mutate(moy_flux = mean(Flux_wan_92, na.rm=T)) %>% 
#   mutate(yyy = format(datetime, format = "%Y"))


# annual fluxes (W 2014) : change of unity (per year and not per hour)
tab_flux <- tab_flux %>% 
  mutate(flux_annual_2014 = (0.00077 * speed_ms^2 * delta_pCO2)) %>% 
  group_by(Year) %>% 
  mutate(mean_flux_annual = mean(flux_annual_2014, na.rm=TRUE))

#selection of only years and mean of fluxes 2014 :
y <- tab_flux %>% distinct(mean_flux_annual)

#only year 2009 :
y_2009 <- tab_flux %>% dplyr::filter(Year == "2009")
mean(y_2009$flux_annual_2014, na.rm=TRUE) # 0.14

y %>% 
  ggplot() + 
  aes(x=Year, y=mean_flux_annual) + 
  geom_col(fill="#303030") + 
  scale_x_discrete(name="", limits=c("2009" : "2022")) + 
  scale_y_continuous(name=expression(paste("Flux (mol ", m^-2, " ", an^-1, ")")), n.breaks = 10) +
  geom_hline(col = "black", yintercept = 0)


## save flux
#write.table(tab_flux, "DATA_SOMLIT_07-22_impute_0m_deltas_pCO2_and_fluxes_CO2.csv", sep=";", col.names = TRUE, row.names = FALSE)

```
  
  
## Comparison with DE CARLO results (jan.2007 - oct.2011)  
  
  
Point B : 2009-2011, so comparison for 2009-2011 period  
  

```{r Comparison with DE CARLO results, echo=FALSE, warning=FALSE, message=FALSE}

#select 2007-2011 on tab-flux :
de_carlo <- tab_flux %>% 
  dplyr::filter(datetime >= "2009-01-06" & datetime <= "2011-10-04")

de_carlo %>% 
   ggplot() + 
   aes(x=datetime, y=Flux_CO2_2014) + 
   geom_line(col="#967C5C") +
   scale_x_date(name="") +
   scale_y_continuous(name="Flux (mol of C/m²/y)") +
   ggtitle("2009-2011 : CO2 fluxes ocean-atmosphere (Wanninkhof 2014)")

de_carlo <- de_carlo %>% 
  mutate(average_flux_h = mean(Flux_CO2_2014, na.rm=T),
         max_flux_h = max(Flux_CO2_2014, na.rm=T),
         min_flux_h = min(Flux_CO2_2014, na.rm=T))

mean(de_carlo$Flux_CO2_2014, na.rm=TRUE)

```
  
  
  
De Carlo (2007-2011) :  
- average hourly flux : -2.19e-05  
- annualized flux (by year) : -1.91e-01  
- maximum hourly flux (winter) : -4.57e-04  
- minimum hourly flux (summer): 4.29e-04  
  
  
De Carlo (01/06/2009 - 10/04/2011) :  
- average hourly flux : -2.77e-05  
- annualized flux (per year) : 2.42e-01  
- maximum hourly flux (winter) : -4.57e-04  
- minimum hourly flux (summer): 4.29e-04  
   
Point B (01/06/2009 - 10/04/2011) :  
- average hourly flux : -1.65e-06  
- annualized flux (by year) : 1.45e-02
- maximum hourly flux (winter) : -3.23e-04  
- minimum hourly flux (summer) : 5.70e-04  
  


