---
title: Interpolations
author: |
   | M&eacute;gane Ballin
   
date: '`r format(Sys.time(), "%d %B %Y %H:%M")`'

output:
    rmdformats::robobook:
     self_contained: true
     thumbnails: false
     lightbox: true
     gallery: true
     highlight: pygments
---


```{r libraries, echo=FALSE, warning=FALSE, message=FALSE}
library("readr")
library("viridis")
library("scales")
library("lubridate")
library("gsignal")
library("FactoMineR")
library("factoextra")
library("gplots")
library("heatwaveR")
library("dplyr")
library("imputeTS")
library("Metrics")
library("tidyr")
library("tseries")
library("forecast")
library("tsDyn")
library("tsoutliers")
library("seacarb")
library("lmodel2")
library("lmtest")
library("stringr")
```
  

# Point B : surface (2007-2022) 
  
Utilisation of : na_seadec() function  
= Removes the seasonal component from the time series, performs imputation on the deseasonalized series and afterwards adds the seasonal component again.  
  
  
## Temperature & salinity
  

```{r interpolation temp & salinity, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7,fig.height=5}

SOMLIT_pcO2 <- read_delim("../../Data/Data_PointB/DATA_SOMLIT_07-22_impute_0m.csv", 
    delim = ";", escape_double = FALSE, trim_ws = TRUE) %>% 
  dplyr::filter(datetime <= "2021-11-09")

impute_temp <- na_seadec(SOMLIT_pcO2$temperature, algorithm = "interpolation", find_frequency = F)
impute_sal <- na_seadec(SOMLIT_pcO2$salinity, algorithm = "interpolation", find_frequency = F)


SOMLIT_pcO2$temperature[is.na(SOMLIT_pcO2$pCO2_atmos)] <- "NA"

SOMLIT_pcO2 %>% 
  dplyr::filter(temperature == "NA") %>% 
  dplyr::group_by(Year) %>% 
  dplyr::summarise(number_of_NA = n()) %>% 
  ggplot(aes(x = Year, y = number_of_NA))  +
  geom_segment(aes(x = Year, xend = Year, y = 0, yend = number_of_NA), 
               color = "grey", linewidth = 2) +
  geom_point(size = 5, fill = "#689D71", shape = 21, color = "black") +
  coord_flip() +
  scale_y_continuous(name = "Nombre de données manquantes", limits = c(0,15), breaks = seq(0, 15, 1)) + 
  scale_x_discrete(name="", limits=c("2007":"2021"))

SOMLIT_pcO2 <- SOMLIT_pcO2 %>% 
  mutate(pCO2_atmos = as.numeric(pCO2_atmos))


#ggplot_na_gapsize(SOMLIT_pcO2$pCO2_atmos)


# ggplot_na_imputations(SOMLIT_pcO2$pCO2_atmos, SOMLIT_pcO2$impute_pCO2_atmos, x_axis_labels = SOMLIT_pcO2$datetime, xlab = "", 
#                       ylab=expression(paste(italic(p), CO[2], " atmosphérique ", "(atm)")), 
#                       title="", color_lines = "lightslategray",
#                       color_points = "steelblue", size_points = 1.2, size_imputations = 1.7) +
#   scale_x_date(date_breaks = "2 years", date_labels = "%Y") 
#   



```

## *p*CO~2~ water
  

```{r interpolation pCO2_w, echo=FALSE, warning=FALSE, message=FALSE}

#interpolation pCO2_w : - NAs at the end

impute_pCO2_w <- na_seadec(SOMLIT_pcO2$pCO2_w, algorithm = "interpolation", find_frequency = F)


ggplot_na_imputations(SOMLIT_pcO2$pCO2_w, SOMLIT_pcO2$impute_pCO2_w, x_axis_labels = SOMLIT_pcO2$datetime,
                      xlab = "", ylab="pCO2 (µatm)", 
                      title="Point B - Ocean pCO2 time series (µatm) - period 2007-2022")



```

<!-- ## *A*~T~ & *C*~T~ -->
  
```{r interpolation AT & CT, echo=FALSE, warning=FALSE, message=FALSE}
# 
# #interpolation alcalinity : - NAs at the end
# 
# impute_ta <- na_seadec(SOMLIT_pcO2$ta, algorithm = "interpolation", find_frequency = T)
# 
# ggplot_na_imputations(SOMLIT_pcO2$ta, SOMLIT_pcO2$impute_ta, x_axis_labels = SOMLIT_pcO2$datetime, xlab = "", 
#                       ylab="AT (µmol/kg)", 
#                       title="Point B - Total alkalinity time series (µmol/kg) - period 2007-2022")
# 
# 
# #interpolation CT (dic) : - NAs at the end
# 
# impute_dic <- na_seadec(Somlit_interpol$dic, algorithm = "interpolation", find_frequency = T)
# 
# ggplot_na_imputations(SOMLIT_pcO2$dic, SOMLIT_pcO2$impute_dic, x_axis_labels = SOMLIT_pcO2$datetime, xlab = "", 
#                       ylab="AT (µmol/kg)", 
#                       title="Point B - Total carbon time series (µmol/kg) - period 2007-2022")
```
  
## Atmospheric *p*CO~2~ (µatm) 
  
Period: Jan.2007 - Dec.2023 
Frequency: daily (09:00 AM) - To be changed
Unit: µatm  
  
→ Corse Ersa (40m) - 2013-2020  
→ Plateau Rosa (10m) - 2007-2022  
→ Lampedusa (10m) - 2006-2022  
  
```{r importation pCO2 air, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7,fig.height=4}

 pCO2_air <- read_delim("../../Data/atmospheric_pco2_icos/DATA_pCO2_atmos_06-23.csv", 
     delim = ";", escape_double = FALSE, trim_ws = TRUE) %>% 
   dplyr::filter(datetime >= "2007-01-01" & datetime <= "2023-12-31") %>% 
   dplyr::select(-c(CO2_LAMP, CO2_CORSE, CO2_PR, CO2_HP, CO2_MC, pCO2_MC, pCO2_HP))
 
 
merge1 <- left_join(SOMLIT_pcO2, pCO2_air, by="datetime")

##RAPPORT

reg1 <- summary(lm(pCO2_air$pCO2_LAMP ~ pCO2_air$pCO2_CORSE)) #0.86
reg2 <- summary(lm(pCO2_air$pCO2_PR ~ pCO2_air$pCO2_CORSE)) #0.85


 pCO2_air %>% 
   ggplot() + 
   geom_point(aes(x=datetime, y=pCO2_CORSE), col = "#960018", size=0.6) +
   geom_point(aes(x=datetime, y=pCO2_LAMP), col = "#1FA055", size=0.5) +
   scale_x_date(name="", date_breaks = "2 years", date_minor_breaks = "1 year", date_labels = "%Y") +
   scale_y_continuous(name=expression(paste(italic(p), CO[2], " (",mu, "atm)"))) +
   annotate(geom="text", x=as.Date("2015-02-03"), y=420, label="R² = 0.85 *", color="black")

pCO2_air %>% 
  ggplot() + 
  geom_point(aes(x=datetime, y=pCO2_CORSE), col = "#960018", size=0.6) +
  geom_point(aes(x=datetime, y=pCO2_PR), col = "#357AB7", size=0.5) +
  scale_x_date(name="", date_breaks = "2 years", date_minor_breaks = "1 year", date_labels = "%Y") +
  scale_y_continuous(name=expression(paste(italic(p), CO[2], " (",mu, "atm)"))) +
  annotate(geom="text", x=as.Date("2015-02-03"), y=420, label="R² = 0.83 *", color="black")


```
  

→ Cross interpolation between 3 sites : Corsica x Lampedusa x Plateau Rosa  
  
→ Then, linear interpolation (110 NAs) :   
  

```{r cross interpolation atmospheric pCO2, echo=FALSE, warning=FALSE, message=FALSE}

# pCO2_atmos = interpolation between 3 sites (Corsica x Lampedusa x Plateau Rosa)

merge1 <- merge1 %>% 
 mutate(pCO2_CORSE = str_replace_na(pCO2_CORSE.x, replacement = "NA"),
        pCO2_PR = str_replace_na(pCO2_PR.x, replacement = "NA"),
        pCO2_LAMP = str_replace_na(pCO2_LAMP.x, replacement = "NA"),
        pCO2_atmos = case_when(pCO2_CORSE == "NA" ~ pCO2_LAMP, TRUE ~ pCO2_CORSE),
        pCO2_atmos = case_when(pCO2_atmos == "NA" ~ pCO2_PR, TRUE ~ pCO2_atmos))
 
 
merge1 <- merge1 %>% 
 mutate(pCO2_CORSE = as.numeric(pCO2_CORSE),
        pCO2_PR = as.numeric(pCO2_PR),
        pCO2_LAMP = as.numeric(pCO2_LAMP),
        pCO2_atmos = as.numeric(pCO2_atmos)) #110 NAs
 
# creation merge1_interpol = merge1 without last rows ( - pCO2_atmos NAs)

merge1_interpol <- merge1 %>% dplyr::filter(datetime <= "2021-12-28")

#linear interpolation pCO2 atmos :

impute_pCO2_atmos <- na_seadec(merge1_interpol$pCO2_atmos, algorithm="interpolation")
 
ggplot_na_imputations(SOMLIT_pcO2$pCO2_atmos, SOMLIT_pcO2$impute_pCO2_atmos, x_axis_labels = SOMLIT_pcO2$datetime, 
                       xlab = "", ylab="pCO2 (µatm)", 
                       title="Atmospheric pCO2 (µatm) time series interpolation - period 2007-2021")
##

# #merge of impute pCO2_atmos at merge3 : 
# merge1_interpol <- cbind(merge1_interpol, impute_pCO2_atmos)
#  
# #join merge1_interpol to SOMLIT_pcO2 :
# SOMLIT_pcO2 <- left_join(SOMLIT_pcO2, merge1_interpol)


```
  
  
Idem for 50m  
  

```{r Point B - 50m - interpolations, echo=FALSE, warning=FALSE, message=FALSE}

# #50m data :
# SOMLIT_50 <- SOMLIT_pcO2_raw %>% 
#   dplyr::filter(depth==50) %>% 
#   rename(pCO2_w = pCO2, datetime = sampling_date) %>%
#   mutate(salinity = case_when(salinity < 35 ~ NA_real_ , TRUE ~ salinity))
# ##
# SOMLIT_50 <- distinct(SOMLIT_50)
# 
# #creation tab somlit without last rows to do interpolations correctly for pCO2, pH, ta & dic (CT) :
# Somlit_interpol_50 <- SOMLIT_50[c(1:821),]
# ##
# 
# 
# SOMLIT_50 <- SOMLIT_50 %>% 
#   mutate(impute_temp = na_seadec(temperature, algorithm = "interpolation", find_frequency = T),
#          impute_sal = na_seadec(salinity, algorithm = "interpolation", find_frequency = T))
#   
# Somlit_interpol_50 <- Somlit_interpol_50 %>% 
#   mutate(impute_pco2_w = na_seadec(pCO2_w, algorithm = "interpolation", find_frequency = T),
#          impute_ta = na_seadec(ta, algorithm = "interpolation", find_frequency = T),
#          impute_dic = na_seadec(dic, algorithm = "interpolation", find_frequency = T))
# 
# 
# SOMLIT_50 <- left_join(SOMLIT_50, Somlit_interpol_50)

```





