---
title: Comparison with Kapsenberg et al. (2017)
author: |
   | M&eacute;gane Ballin
   
date: '`r format(Sys.time(), "%d %B %Y %H:%M")`'

output:
    rmdformats::robobook:
     self_contained: true
     thumbnails: false
     lightbox: true
     gallery: true
     highlight: pygments
---

# Description

Time series trends analysis by residuals (substracting climatological monthly means)

Kapsenberg et al. (2017)

```{r libraries, echo=FALSE, warning=FALSE, message=FALSE}
library("readr")
library("viridis")
library("scales")
library("lubridate")
library("gsignal")
library("FactoMineR")
library("factoextra")
library("gplots")
library("heatwaveR")
library("dplyr")
library("imputeTS")
library("Metrics")
library("tidyr")
library("tseries")
library("forecast")
library("tsDyn")
library("tsoutliers")
library("seacarb")
library("lmodel2")
library("lmtest")
library("cowplot")
library("kableExtra")
library("knitr")
```

```{r import data (0m et 50m) + periode 2007-2015, echo=FALSE, warning=FALSE, message=FALSE}
data_raw <- read_csv("../../Data/data_PointB/data_pH_steeve.csv")
data_raw_surf <- data_raw %>% dplyr::filter(depth==0)
data_raw_prof <- data_raw %>% dplyr::filter(depth==50)

data_0715 <- data_raw_surf %>% dplyr::filter(sampling_date < "2016-01-05") #470 values
data_0715_50m <- data_raw_prof %>% dplyr::filter(sampling_date < "2016-01-05") #462 values


#theme for plots :
Mytheme <- function(size_labs = 6, face_font="plain", ...) {
  theme_bw() +
    theme(axis.text.x = element_text(face=face_font, size=size_labs, color="black"),
          axis.title.x = element_text(face=face_font, size=size_labs, margin=margin(0,0,0,0,"pt")),
          axis.text.y = element_text(face=face_font, color="black", size=size_labs),
          axis.title.y = element_text(face=face_font, size=size_labs),
          axis.ticks.x = element_line(size=0.1),
          axis.ticks.y = element_line(size=0.1),
          axis.ticks.length = unit(1.1, "mm"),
          panel.grid.major = element_line(size = 0.25, color="black", linetype="dotted"),
          aspect.ratio = 1 / 3,
          plot.margin = margin(t = 0, r = 1, b = 0, l = 0, unit = "lines")
    )
}
##
```

# 2007-2015 : Comparison with Kapsenberg results

## SURFACE

### [Table of monthly means (2007-2015) - surface data]{style="color:blue"}

➝ Comparaison supplement, Tab S1 (Kapsenberg) same period : results ok ✓

```{r Calcul monthly means 0715 - surface, echo=TRUE, warning=FALSE, message=FALSE}

monthly_means_0m_0715 <- ungroup(data_0715) %>% 
  group_by(monthd) %>%
  summarise(
    Salinity_month = mean(salinity, na.rm = TRUE),
    Temperature_month = mean(temperature, na.rm = TRUE),
    dic_month = mean(dic, na.rm = TRUE),
    ta_month = mean(ta, na.rm = TRUE),
    pH_calc_month = mean(pH_calc, na.rm = TRUE),
    pH18_month = mean(pH18, na.rm = TRUE),
    pCO2_month = mean(pCO2, na.rm = TRUE),
    OmegaCalcite_month = mean(OmegaCalcite, na.rm = TRUE),
    OmegaAragonite_month = mean(OmegaAragonite, na.rm = TRUE),
    Nta_month = mean(Nta, na.rm = TRUE),
    Ndic_month = mean(dic, na.rm = TRUE))
monthly_means_0m_0715
```

### [Table of time series anomalies regression analyses (2007-2015) - surface data]{style="color:green"}

➝ Comparison Table 2 (Kapsenberg) same period : results ok ✓

```{r anomalies + regressions 2007-2015 (Kapsenberg) - surface, echo=TRUE, warning=FALSE, message=FALSE}

ano_0m_0715 <- left_join(ungroup(data_0715), monthly_means_0m_0715, by = "monthd") %>%
  mutate(
    Salinity_ano = salinity - Salinity_month,
    Temperature_ano = temperature - Temperature_month,
    dic_ano = dic - dic_month,
    ta_ano = ta - ta_month,
    pH_calc_ano = pH_calc - pH_calc_month,
    pH18_ano = pH18 - pH18_month,
    pCO2_ano = pCO2 - pCO2_month,
    OmegaCalcite_ano = OmegaCalcite - OmegaCalcite_month,
    OmegaAragonite_ano = OmegaAragonite - OmegaAragonite_month,
    Nta_ano = Nta - Nta_month,
    Ndic_ano = dic - Ndic_month)

# Regressions and tables of key parameters
var_list <-
  c("salinity",
    "temperature",
    "dic",
    "ta",
    "pH_calc",
    "pH18",
    "pCO2",
    "OmegaCalcite",
    "OmegaAragonite",
    "Nta",
    "Ndic")

lm.test <- vector("list", length(var_list))
 
 for(i in seq_along(var_list)){
     lm.test[[i]] <- lm(reformulate(var_list[i], "sampling_date"), data = ano_0m_0715)
 }


lms <- lapply(var_list, function(x) {
  summary(lm(substitute(i ~ decimal_date(sampling_date), list(i = as.name(x))), 
             data = ano_0m_0715))
})

reg <- NULL
for (i in 1:length(var_list)) {
  #one loops through all anomalies
  # calculate probability of fstatistic because it cannot be extracted from lms above
  # see http://stats.stackexchange.com/questions/92824/extracting-the-model-p-value-for-a-multiple-regression-in-r
  # slope returned by lm is per second
  prob <-
    pf(lms[[i]]$fstatistic[1],
       lms[[i]]$fstatistic[2],
       lms[[i]]$fstatistic[3],
       lower.tail = FALSE)
  reg <-
    rbind(reg, as.numeric(
      c(
        lms[[i]]$coefficients[2, 1],
        lms[[i]]$coefficients[2, 2],
        lms[[i]]$coefficients[2, 4],
        lms[[i]]$coefficients[1, 1],
        lms[[i]]$coefficients[1, 2],
        lms[[i]]$coefficients[1, 4],
        lms[[i]]$fstatistic[1],
        lms[[i]]$fstatistic[3],
        lms[[i]]$r.squared,
        prob
      )
    ))
}

colnames(reg) <-
  c("Slope",
    "SE Slope",
    "P Slope",
    "Intercept",
    "SE int.",
    "P int.",
    "F",
    "df",
    "R2",
    "P value")
row.names(reg) <- var_list
reg_ano_surf <- reg
slope_temp <- reg_ano_surf[2, 1]

# Regression and table anomalies
var_list <-
  c(
    "Salinity_ano",
    "Temperature_ano",
    "dic_ano",
    "ta_ano",
    "pH_calc_ano",
    "pH18_ano",
    "pCO2_ano",
    "OmegaCalcite_ano",
    "OmegaAragonite_ano",
    "Nta_ano",
    "Ndic_ano"
  )
lms <- lapply(var_list, function(x) {
  summary(lm(substitute(i ~ decimal_date(sampling_date), list(i = as.name(x))), 
             data = ano_0m_0715))
})
reg <- NULL
for (i in 1:length(var_list)) {
  #one loops through all anomalies
  # calculate probability of fstatistic because it cannot be extracted from lms above
  # see http://stats.stackexchange.com/questions/92824/extracting-the-model-p-value-for-a-multiple-regression-in-r
  # slope returned by lm is per second
  prob <-
    pf(lms[[i]]$fstatistic[1],
       lms[[i]]$fstatistic[2],
       lms[[i]]$fstatistic[3],
       lower.tail = FALSE)
  reg <-
    rbind(reg, as.numeric(
      c(
        lms[[i]]$coefficients[2, 1],
        lms[[i]]$coefficients[2, 2],
        lms[[i]]$coefficients[2, 4],
        lms[[i]]$coefficients[1, 1],
        lms[[i]]$coefficients[1, 2],
        lms[[i]]$coefficients[1, 4],
        lms[[i]]$fstatistic[1],
        lms[[i]]$fstatistic[3],
        lms[[i]]$r.squared,
        prob
      )
    ))
}
colnames(reg) <-
  c("Slope",
    "SE Slope",
    "P Slope",
    "Intercept",
    "SE int.",
    "P int.",
    "F",
    "df",
    "R2",
    "P value")
row.names(reg) <- var_list
reg_ano_surf <- reg
slope_temp <- reg_ano_surf[2, 1]
slope_pH <- reg_ano_surf[7, 1]

reg_ano_surf

```

### Plot of anomalies : temperature (2007-2015) - surface

```{r, plot anomalies temperature 2007-2015 - surface, echo=FALSE, warning=FALSE, message=FALSE, fig.width=8,fig.height=4,fig.align='center'}

#Anomalies 0 m
plot_ano_temp_0715 <- ggplot(data = ano_0m_0715, aes(x = sampling_date, y = Temperature_ano), na.rm=TRUE) +
  scale_x_datetime(date_breaks="2 year", date_minor_breaks="1 years", labels = date_format("%Y")) +
  geom_point(colour="blue", na.rm=TRUE, size=1) + 
  geom_smooth(method=lm, colour="black", fill="grey", linewidth=0.6, na.rm=TRUE) +
  labs(title=NULL,x="", y="Temp. (°C)") +
  annotate("text", x =as.POSIXct("2016-01-01"), y= min(ano_0m_0715$Temperature_ano, na.rm=TRUE), 
           label="", colour="black", size=9, fontface="plain") +
  coord_fixed() +
  Mytheme(size_labs = 8) +
  theme(axis.text.x=element_blank()) + 
  ggtitle("Temperature anomalies + trend")

plot_temp_0715 <- data_0715 %>% 
  ggplot() +
  aes(x=sampling_date, y=temperature) + 
  geom_point(size=1) +
  scale_x_datetime(name="") +
  scale_y_continuous(name="Temp. (°C)") +
  ggtitle("Time series observations at 0m of temperature (°C) 2007-2015")

plot_temp_0715
plot_ano_temp_0715


```

### Plot of anomalies : salinity (2007-2015) - surface

```{r, plot anomalies salinity 2007-2015 - surface, echo=FALSE, warning=FALSE, message=FALSE, fig.width=8,fig.height=4,fig.align='center'}

plot_ano_sal_0715 <- ggplot(data = ano_0m_0715, aes(x = sampling_date, y = Salinity_ano), na.rm=TRUE) +
  scale_x_datetime(date_breaks="2 year", date_minor_breaks="1 years", labels = date_format("%Y")) +
  scale_y_continuous(breaks=c(-2,-1,0,0.5)) +
  geom_point(colour="blue", na.rm=TRUE, size=1) + 
  geom_smooth(method=lm, colour="black", fill="grey", linewidth=0.6, na.rm=TRUE) +
  labs(title=NULL,x="", y="Salinity") +
  coord_fixed() +
  Mytheme(size_labs = 8) +
  theme(axis.text.x=element_blank())

plot_sal_0715 <- data_0715 %>% 
  ggplot() +
  aes(x=sampling_date, y=salinity) + 
  geom_point(size=1) +
  scale_x_datetime(name="") +
  scale_y_continuous(name="Salinity") +
  ggtitle("Time series observations at 0m of salinity 2007-2015")

plot_sal_0715
plot_ano_sal_0715
```

### Plot of anomalies : pH (2007-2015) - surface

```{r, plot anomalies pH 2007-2015 - surface, echo=FALSE, warning=FALSE, message=FALSE, fig.width=8,fig.height=4,fig.align='center'}

plot_ano_ph_0715 <- ggplot(data = ano_0m_0715, aes(x = sampling_date, y = pH_calc_ano), na.rm=TRUE) +
  scale_x_datetime(date_breaks="2 year", date_minor_breaks="1 years", labels = date_format("%Y")) +
  geom_point(colour="blue", na.rm=TRUE, size=1) + 
  geom_smooth(method=lm, colour="black", fill="grey", linewidth=0.6, na.rm=TRUE) +
  labs(title=NULL, x="", y=expression(paste(pH[T]," calculated"))) +
  annotate("text", x =as.POSIXct("2016-01-01"), y= min(ano_0m_0715$pH_calc_ano, na.rm=TRUE), 
           label="", colour="black", size=9, fontface="plain") +
  coord_fixed() +
  Mytheme(size_labs = 8) +
  theme(axis.text.x=element_blank())

plot_ph_0715 <- data_0715 %>% 
  ggplot() +
  aes(x=sampling_date, y=pH_calc) + 
  geom_point(size=1) +
  scale_x_datetime(name="") +
  scale_y_continuous(name="pH") +
  ggtitle("Time series observations at 0m of pH (2007-2015)")

plot_ph_0715
plot_ano_ph_0715
```

### Plot of anomalies : alkalinity (2007-2015) - surface

```{r, plot anomalies alcalinity 2007-2015 - surface, echo=FALSE, warning=FALSE, message=FALSE, fig.width=8,fig.height=4,fig.align='center'}

plot_ano_ta_0715 <- ggplot(data = ano_0m_0715, aes(x = sampling_date, y = ta_ano), na.rm=TRUE) +
  scale_x_datetime(date_breaks="2 year", date_minor_breaks="1 years", labels = date_format("%Y")) +
  geom_point(colour="blue", na.rm=TRUE, size=1) + 
  geom_smooth(method=lm, colour="black", fill="grey", linewidth=0.6, na.rm=TRUE) +
  labs(title=NULL,x="",y=expression(paste(pCO[2], " (",mu, "atm)"))) +
  annotate("text", x =as.POSIXct("2016-01-01"), y= min(ano_0m_0715$ta_ano, na.rm=TRUE), 
           label="", colour="black", size=9, fontface="plain") +
  coord_fixed() +
  Mytheme(size_labs = 8) +
  theme(axis.text.x=element_blank())

plot_ta_0715 <- data_0715 %>% 
  ggplot() +
  aes(x=sampling_date, y=ta) + 
  geom_point(size=1) +
  scale_x_datetime(name="") +
  scale_y_continuous(name="Alcalinity (umol/kg") +
  ggtitle("Time series observations at 0m of alcalinity (2007-2015)")

plot_ta_0715
plot_ano_ta_0715
```

### Plot of anomalies : pCO2 (2007-2015) - surface

```{r, plot anomalies CO2 1 - surface, echo=FALSE, warning=FALSE, message=FALSE, fig.width=8,fig.height=4,fig.align='center'}
plot_ano_pCO2_0715 <- ggplot(data = ano_0m_0715, aes(x = sampling_date, y = pCO2_ano), na.rm=TRUE) +
  scale_x_datetime(date_breaks="2 year", date_minor_breaks="1 years", labels = date_format("%Y")) +
  geom_point(colour="blue", na.rm=TRUE, size=1) + 
  geom_smooth(method=lm, colour="black", fill="grey", linewidth=0.6, na.rm=TRUE) +
  labs(title=NULL,x="",y=expression(paste(pCO[2], " (",mu, "atm)"))) +
  annotate("text", x =as.POSIXct("2016-01-01"), y= min(ano_0m_0715$pCO2_ano, na.rm=TRUE), 
           label="", colour="black", size=9, fontface="plain") +
  coord_fixed() +
  Mytheme(size_labs = 8) +
  theme(axis.text.x=element_blank())

plot_pCO2_0715 <- data_0715 %>% 
  ggplot() +
  aes(x=sampling_date, y=pCO2) + 
  geom_point(size=1) +
  scale_x_datetime(name="") +
  scale_y_continuous(name="pCO2 (uatm)") +
  ggtitle("Time series observations of pCO2 (2007-2015)")

plot_pCO2_0715
plot_ano_pCO2_0715
```

## DEPTH 50 m

### [Table of monthly means (2007-2015) - 50m data]{style="color:blue"}

➝ Comparaison supplement, Tab S1 (Kapsenberg) same period : results ok ✓

```{r Calcul monthly means 0715 - 50m, echo=FALSE, warning=FALSE, message=FALSE}

monthly_means_50m_0715 <- ungroup(data_0715_50m) %>% 
  group_by(monthd) %>%
  summarise(
    Salinity_month = mean(salinity, na.rm = TRUE),
    Temperature_month = mean(temperature, na.rm = TRUE),
    dic_month = mean(dic, na.rm = TRUE),
    ta_month = mean(ta, na.rm = TRUE),
    pH_calc_month = mean(pH_calc, na.rm = TRUE),
    pH18_month = mean(pH18, na.rm = TRUE),
    pCO2_month = mean(pCO2, na.rm = TRUE),
    OmegaCalcite_month = mean(OmegaCalcite, na.rm = TRUE),
    OmegaAragonite_month = mean(OmegaAragonite, na.rm = TRUE),
    Nta_month = mean(Nta, na.rm = TRUE),
    Ndic_month = mean(dic, na.rm = TRUE))
monthly_means_50m_0715
```

### [Table of time series anomalies regression analyses (2007-2015) - 50m data]{style="color:green"}

➝ Comparison Table 2 (Kapsenberg) same period: results ok ✓

```{r anomalies + regressions 2007-2015 (Kapsenberg) - 50m, echo=FALSE, warning=FALSE, message=FALSE}

ano_50m_0715 <- left_join(ungroup(data_0715_50m), monthly_means_50m_0715, by = "monthd") %>%
  mutate(
    Salinity_ano = salinity - Salinity_month,
    Temperature_ano = temperature - Temperature_month,
    dic_ano = dic - dic_month,
    ta_ano = ta - ta_month,
    pH_calc_ano = pH_calc - pH_calc_month,
    pH18_ano = pH18 - pH18_month,
    pCO2_ano = pCO2 - pCO2_month,
    OmegaCalcite_ano = OmegaCalcite - OmegaCalcite_month,
    OmegaAragonite_ano = OmegaAragonite - OmegaAragonite_month,
    Nta_ano = Nta - Nta_month,
    Ndic_ano = dic - Ndic_month)

# Regressions and tables of key parameters
var_list_50 <-
  c("salinity",
    "temperature",
    "dic",
    "ta",
    "pH_calc",
    "pH18",
    "pCO2",
    "OmegaCalcite",
    "OmegaAragonite",
    "Nta",
    "Ndic")

lm.test_50 <- vector("list", length(var_list_50))
 
 for(i in seq_along(var_list_50)){
     lm.test_50[[i]] <- lm(reformulate(var_list_50[i], "sampling_date"), data = ano_50m_0715)
 }


lms_50 <- lapply(var_list_50, function(x) {
  summary(lm(substitute(i ~ decimal_date(sampling_date), list(i = as.name(x))), 
             data = ano_50m_0715))
})

reg_50 <- NULL
for (i in 1:length(var_list_50)) {
  #one loops through all anomalies
  # calculate probability of fstatistic because it cannot be extracted from lms above
  # see http://stats.stackexchange.com/questions/92824/extracting-the-model-p-value-for-a-multiple-regression-in-r
  # slope returned by lm is per second
  prob <-
    pf(lms_50[[i]]$fstatistic[1],
       lms_50[[i]]$fstatistic[2],
       lms_50[[i]]$fstatistic[3],
       lower.tail = FALSE)
  reg_50 <-
    rbind(reg_50, as.numeric(
      c(
        lms_50[[i]]$coefficients[2, 1],
        lms_50[[i]]$coefficients[2, 2],
        lms_50[[i]]$coefficients[2, 4],
        lms_50[[i]]$coefficients[1, 1],
        lms_50[[i]]$coefficients[1, 2],
        lms_50[[i]]$coefficients[1, 4],
        lms_50[[i]]$fstatistic[1],
        lms_50[[i]]$fstatistic[3],
        lms_50[[i]]$r.squared,
        prob
      )
    ))
}

colnames(reg_50) <-
  c("Slope",
    "SE Slope",
    "P Slope",
    "Intercept",
    "SE int.",
    "P int.",
    "F",
    "df",
    "R2",
    "P value")
row.names(reg_50) <- var_list_50
reg_ano_50m <- reg_50

# Regression and table anomalies
var_list_50 <-
  c(
    "Salinity_ano",
    "Temperature_ano",
    "dic_ano",
    "ta_ano",
    "pH_calc_ano",
    "pH18_ano",
    "pCO2_ano",
    "OmegaCalcite_ano",
    "OmegaAragonite_ano",
    "Nta_ano",
    "Ndic_ano"
  )
lms_50 <- lapply(var_list_50, function(x) {
  summary(lm(substitute(i ~ decimal_date(sampling_date), list(i = as.name(x))), 
             data = ano_50m_0715))
})
reg_50 <- NULL
for (i in 1:length(var_list_50)) {
  #one loops through all anomalies
  # calculate probability of fstatistic because it cannot be extracted from lms above
  # see http://stats.stackexchange.com/questions/92824/extracting-the-model-p-value-for-a-multiple-regression-in-r
  # slope returned by lm is per second
  prob <-
    pf(lms_50[[i]]$fstatistic[1],
       lms_50[[i]]$fstatistic[2],
       lms_50[[i]]$fstatistic[3],
       lower.tail = FALSE)
  reg_50 <-
    rbind(reg_50, as.numeric(
      c(
        lms_50[[i]]$coefficients[2, 1],
        lms_50[[i]]$coefficients[2, 2],
        lms_50[[i]]$coefficients[2, 4],
        lms_50[[i]]$coefficients[1, 1],
        lms_50[[i]]$coefficients[1, 2],
        lms_50[[i]]$coefficients[1, 4],
        lms_50[[i]]$fstatistic[1],
        lms_50[[i]]$fstatistic[3],
        lms_50[[i]]$r.squared,
        prob
      )
    ))
}
colnames(reg_50) <-
  c("Slope",
    "SE Slope",
    "P Slope",
    "Intercept",
    "SE int.",
    "P int.",
    "F",
    "df",
    "R2",
    "P value")
row.names(reg_50) <- var_list_50
reg_ano_50 <- reg_50

reg_ano_50

```

### Plot of anomalies : temperature (2007-2015) - 50m

```{r, plot anomalies temperature 2007-2015 - 50m, echo=FALSE, warning=FALSE, message=FALSE, fig.width=8,fig.height=4,fig.align='center'}

#Anomalies 50 m
plot_ano_temp_0715_50m <- ggplot(data = ano_50m_0715, aes(x = sampling_date, y = Temperature_ano), na.rm=TRUE) +
  scale_x_datetime(date_breaks="2 year", date_minor_breaks="1 years", labels = date_format("%Y")) +
  geom_point(colour="blue", na.rm=TRUE, size=1) + 
  geom_smooth(method=lm, colour="black", fill="grey", linewidth=0.6, na.rm=TRUE) +
  labs(title=NULL,x="", y="Temp. (°C)") +
  annotate("text", x =as.POSIXct("2016-01-01"), y= min(ano_50m_0715$Temperature_ano, na.rm=TRUE), 
           label="", colour="black", size=9, fontface="plain") +
  coord_fixed() +
  Mytheme(size_labs = 8) +
  theme(axis.text.x=element_blank()) + 
  ggtitle("Temperature anomalies + trend")

plot_temp_0715_50m <- data_0715_50m %>% 
  ggplot() +
  aes(x=sampling_date, y=temperature) + 
  geom_point(size=1) +
  scale_x_datetime(name="") +
  scale_y_continuous(name="Temp. (°C)") +
  ggtitle("Time series observations at 50m of temperature (°C) 2007-2015")

plot_temp_0715_50m
plot_ano_temp_0715_50m

```

### Plot of anomalies : salinity (2007-2015) - 50m

```{r, plot anomalies salinity 2007-2015 - 50m, echo=FALSE, warning=FALSE, message=FALSE, fig.width=8,fig.height=4,fig.align='center'}

plot_ano_sal_0715_50m <- ggplot(data = ano_50m_0715, aes(x = sampling_date, y = Salinity_ano), na.rm=TRUE) +
  scale_x_datetime(date_breaks="2 years", date_minor_breaks="1 years", labels = date_format("%Y")) +
  scale_y_continuous(breaks=c(-0.50,-0.25,0,0.25)) +
  geom_point(colour="blue", na.rm=TRUE, size=1) + 
  geom_smooth(method=lm, colour="black", fill="grey", linewidth=0.6, na.rm=TRUE) +
  labs(title=NULL,x="", y="Salinity") +
  coord_fixed() +
  Mytheme(size_labs = 8) +
  theme(axis.text.x=element_blank()) +
  ggtitle("Salinity anomalies + trend")

plot_sal_0715_50m <- data_0715_50m %>% 
  ggplot() +
  aes(x=sampling_date, y=salinity) + 
  geom_point(size=1) +
  scale_x_datetime(name="") +
  scale_y_continuous(name="Salinity") +
  ggtitle("Time series observations at 50m of salinity 2007-2015")

plot_sal_0715_50m
plot_ano_sal_0715_50m
```

### Plot of anomalies : pH (2007-2015) - 50m

```{r, plot anomalies pH 2007-2015 - 50m, echo=FALSE, warning=FALSE, message=FALSE, fig.width=8,fig.height=4,fig.align='center'}

plot_ano_ph_0715_50m <- ggplot(data = ano_50m_0715, aes(x = sampling_date, y = pH_calc_ano), na.rm=TRUE) +
  scale_x_datetime(date_breaks="2 year", date_minor_breaks="1 years", labels = date_format("%Y")) +
  geom_point(colour="blue", na.rm=TRUE, size=1) + 
  geom_smooth(method=lm, colour="black", fill="grey", linewidth=0.6, na.rm=TRUE) +
  labs(title=NULL, x="", y=expression(paste(pH[T]," calculated"))) +
  annotate("text", x =as.POSIXct("2016-01-01"), y= min(ano_50m_0715$pH_calc_ano, na.rm=TRUE), 
           label="", colour="black", size=9, fontface="plain") +
  coord_fixed() +
  Mytheme(size_labs = 8) +
  theme(axis.text.x=element_blank())

plot_ph_0715_50m <- data_0715_50m %>% 
  ggplot() +
  aes(x=sampling_date, y=pH_calc) + 
  geom_point(size=1) +
  scale_x_datetime(name="") +
  scale_y_continuous(name="pH") +
  ggtitle("Time series observations at 50m of pH (2007-2015)")

plot_ph_0715_50m
plot_ano_ph_0715_50m
```

### Plot of anomalies : alcalinity (2007-2015) - 50m

```{r, plot anomalies alcalinity 2007-2015 - 50m, echo=FALSE, warning=FALSE, message=FALSE, fig.width=8,fig.height=4,fig.align='center'}

plot_ano_ta_0715_50m <- ggplot(data = ano_50m_0715, aes(x = sampling_date, y = ta_ano), na.rm=TRUE) +
  scale_x_datetime(date_breaks="2 year", date_minor_breaks="1 years", labels = date_format("%Y")) +
  geom_point(colour="blue", na.rm=TRUE, size=1) + 
  geom_smooth(method=lm, colour="black", fill="grey", linewidth=0.6, na.rm=TRUE) +
  labs(title=NULL,x="",y=expression(paste(pCO[2], " (",mu, "atm)"))) +
  annotate("text", x =as.POSIXct("2016-01-01"), y= min(ano_50m_0715$ta_ano, na.rm=TRUE), 
           label="", colour="black", size=9, fontface="plain") +
  coord_fixed() +
  Mytheme(size_labs = 8) +
  theme(axis.text.x=element_blank())

plot_ta_0715_50m <- data_0715_50m %>% 
  ggplot() +
  aes(x=sampling_date, y=ta) + 
  geom_point(size=1) +
  scale_x_datetime(name="") +
  scale_y_continuous(name="Alcalinity (umol/kg") +
  ggtitle("Time series observations at 50m of alcalinity (2007-2015)")

plot_ta_0715_50m
plot_ano_ta_0715_50m
```

### Plot of anomalies : pCO2 (2007-2015) - 50m

```{r, plot anomalies CO2 2007-2015 - 50m, echo=FALSE, warning=FALSE, message=FALSE, fig.width=8,fig.height=4,fig.align='center'}

plot_ano_pCO2_0715_50m <- ggplot(data = ano_50m_0715, aes(x = sampling_date, y = pCO2_ano), na.rm=TRUE) +
  scale_x_datetime(date_breaks="2 year", date_minor_breaks="1 years", labels = date_format("%Y")) +
  geom_point(colour="blue", na.rm=TRUE, size=1) + 
  geom_smooth(method=lm, colour="black", fill="grey", linewidth=0.6, na.rm=TRUE) +
  labs(title=NULL,x="",y=expression(paste(pCO[2], " (",mu, "atm)"))) +
  annotate("text", x =as.POSIXct("2016-01-01"), y= min(ano_50m_0715$pCO2_ano, na.rm=TRUE), 
           label="", colour="black", size=9, fontface="plain") +
  coord_fixed() +
  Mytheme(size_labs = 8) +
  theme(axis.text.x=element_blank())

plot_pCO2_0715_50m <- data_0715_50m %>% 
  ggplot() +
  aes(x=sampling_date, y=pCO2) + 
  geom_point(size=1) +
  scale_x_datetime(name="") +
  scale_y_continuous(name="pCO2 (uatm)") +
  ggtitle("Time series observations at 50m of pCO2 (2007-2015)")

plot_pCO2_0715_50m
plot_ano_pCO2_0715_50m
```

# 2007-2022

*→ SOMLIT clean + interpolations ok ☑*

## SURFACE

### [Table of monthly means (2007-2022) - surface data]{style="color:blue"}

```{r import data 2007-2022 at 0m and 50m, echo=FALSE, warning=FALSE, message=FALSE}

#data surface
data_0722 <- read_delim("../../Data/data_PointB/DATA_SOMLIT_07-22_impute_0m.csv", 
    delim = ";", escape_double = FALSE, trim_ws = TRUE) %>% 
    mutate(Month = as.numeric(Month))

data_0722_50m <- read_delim("../../Data/data_PointB/DATA_SOMLIT_07-22_impute_50m.csv", 
    delim = ";", escape_double = FALSE, trim_ws = TRUE) %>% 
    mutate(Month = as.numeric(Month))


```

```{r calculate monthly means 2007-2022 - surface, echo=FALSE, warning=FALSE, message=FALSE, fig.width=8,fig.height=10}

#choix pH 18 car moyenne temperature sur toute la periode 2007-2022 : 18.83

monthly_means_0m_0722 <- ungroup(data_0722) %>% 
  group_by(Month) %>%
  summarise(
    Salinity_month = mean(impute_sal, na.rm = TRUE),
    SD_sal = sd(impute_sal, na.rm = TRUE),
    Temperature_month = mean(impute_temp, na.rm = TRUE),
    SD_temp = sd(impute_temp, na.rm = TRUE),
    dic_month = mean(impute_dic, na.rm = TRUE),
    SD_dic = sd(impute_dic, na.rm = TRUE),
    ta_month = mean(impute_ta, na.rm = TRUE),
    SD_ta = sd(impute_ta, na.rm = TRUE),
    pH_calc_month = mean(pH_calc, na.rm = TRUE),
    SD_PH = sd(pH_calc, na.rm = TRUE),
    pH18_month = mean(pH_18, na.rm = TRUE),
    SD_PH18 = sd(pH_18, na.rm = TRUE),
    pCO2_month = mean(impute_pCO2_w, na.rm = TRUE),
    SD_pCO2 = sd(impute_pCO2_w, na.rm = TRUE))


# ##RAPPORT##
# 
# monthly_means_0m_0722_rapport <- monthly_means_0m_0722 %>% 
#   mutate(`T (°C)` = paste(round(Temperature_month, digits=2), "±", round(SD_temp, digits=2)),
#          S = paste(round(Salinity_month, digits=2), "±", round(SD_sal, digits=2)),
#         `pH (18)` = paste(round(pH18_month, digits=2), "±", round(SD_PH18, digits=2)),
#         `*A*~T~ (μmol.kg^-1^)` = paste(round(ta_month, digits=2), "±", round(SD_ta, digits=2)),
#         `*C*~T~ (μmol.kg^-1^)` = paste(round(dic_month, digits=2), "±", round(SD_dic, digits=2)),
#         `*p*CO~2~ (μatm)` = paste(round(pCO2_month, digits=2), "±", round(SD_pCO2, digits=2)))
# 
# monthly_means_0m_0722_rapport <- monthly_means_0m_0722_rapport %>% 
#   dplyr::select(Month, `T (°C)`, S, `pH (18)`, `*A*~T~ (μmol.kg^-1^)`, 
#                 `*C*~T~ (μmol.kg^-1^)`, `*p*CO~2~ (μatm)`) %>% 
#   rename(Mois = Month)
#         
#         
# kable(monthly_means_0m_0722_rapport, caption="Moyennes mensuelles au Point B (surface)",
#       format = "pipe", align='l')


```

### [Table of time series anomaly regression analyses (2007-2022) - surface data]{style="color:green"}

```{r calculate anomalies + regression 2007-2022 - surface, echo=FALSE, warning=FALSE, message=FALSE}

ano_0m_0722 <- left_join(ungroup(data_0722), monthly_means_0m_0722, by = "Month") %>%
  mutate(
    Salinity_ano = impute_sal - Salinity_month,
    Temperature_ano = impute_temp - Temperature_month,
    dic_ano = impute_dic - dic_month,
    ta_ano = impute_ta - ta_month,
    pH_calc_ano = pH_calc - pH_calc_month,
    pH18_ano = pH_18 - pH18_month,
    pCO2_ano = pCO2_w - pCO2_month)


# Regression and table anomalies
var_list2 <-
  c("Temperature_ano",
    "Salinity_ano",
    "pH_calc_ano",
    "pH18_ano",
    "ta_ano",
    "dic_ano",
    "pCO2_ano"
  )
lms2 <- lapply(var_list2, function(x) {
  summary(lm(substitute(i ~ decimal_date(datetime), list(i = as.name(x))), data = ano_0m_0722))
})
reg2 <- NULL
for (i in 1:length(var_list2)) {
  #one loops through all anomalies
  # calculate probability of fstatistic because it cannot be extracted from lms above
  # see http://stats.stackexchange.com/questions/92824/extracting-the-model-p-value-for-a-multiple-regression-in-r
  # slope returned by lm is per second
  prob2 <-
    pf(lms2[[i]]$fstatistic[1],
       lms2[[i]]$fstatistic[2],
       lms2[[i]]$fstatistic[3],
       lower.tail = FALSE)
  reg2 <-
    rbind(reg2, as.numeric(
      c(
        lms2[[i]]$coefficients[2, 1],
        lms2[[i]]$coefficients[2, 2],
        lms2[[i]]$coefficients[2, 4],
        lms2[[i]]$coefficients[1, 1],
        lms2[[i]]$coefficients[1, 2],
        lms2[[i]]$coefficients[1, 4],
        lms2[[i]]$fstatistic[1],
        lms2[[i]]$fstatistic[3],
        lms2[[i]]$r.squared,
        prob2
      )
    ))
}
colnames(reg2) <-
  c("Slope",
    "SE Slope",
    "P Slope",
    "Intercept",
    "SE int.",
    "P int.",
    "F",
    "df",
    "R2",
    "P value")
row.names(reg2) <- var_list2
reg_ano_surf2 <- reg2

reg_ano_surf2

# reg_ano_surf2_rapport <- as.data.frame(reg_ano_surf2)
# reg_ano_surf2_rapport <- reg_ano_surf2_rapport %>% 
#   mutate(Variables = c("T (°C)", "S", "pH~calc~", "pH~(18)~", "*A*~T~ (μmol.kg^-1^)",
#                        "*C*~T~ (μmol.kg^-1^)", "*p*CO~2~ (μatm)"))
# 
# reg_ano_surf2_rapport <- reg_ano_surf2_rapport %>% 
#   mutate( 
#   `Pente ± sd` = paste(round(reg_ano_surf2_rapport$Slope, digits=3), "±", round(reg_ano_surf2_rapport$`SE Slope`, digits=4)), 
#   `P pente` = format(reg_ano_surf2_rapport$`P Slope`, scientific=TRUE),
#   `Int. ± sd` = paste(round(reg_ano_surf2_rapport$Intercept, digits=2), "±", round(reg_ano_surf2_rapport$`SE int.`, digits=2)),
#   `P Int.` = format(reg_ano_surf2_rapport$`P int.`,scientific=TRUE), 
#   `R^2^` = round(reg_ano_surf2_rapport$R2, digits=4))
# 
# 
# reg_ano_surf2_rapport <- reg_ano_surf2_rapport[,c(11:16)]
# 
# kable(reg_ano_surf2_rapport, caption="Time series anomaly regression analyses (Point B surface)", 
#       format = "pipe", align='l')

```

### Plot of anomalies : Temperature (2007-2022) - surface

```{r, plot anomalies temperature 2007-2022 - surface, echo=FALSE, warning=FALSE, message=FALSE, fig.width=8,fig.height=4,fig.align='center'}

plot_ano_temp2 <- ggplot(data = ano_0m_0722, aes(x = datetime, y = Temperature_ano), na.rm=TRUE) +
  scale_x_date(date_breaks="4 year", date_minor_breaks="1 years", labels = NULL) +
  geom_point(colour="#22427C", na.rm=TRUE, size=0.7) + 
  geom_smooth(method=lm, colour="black", fill="grey", linewidth=0.6, na.rm=TRUE) +
  labs(title="",x="", y="Temp. (°C)")


plot_temp_0722 <- data_0722 %>% 
  ggplot() +
  aes(x=datetime, y=impute_temp) + 
  geom_line(size=0.7, col="#303030") +
  scale_x_date(name="", date_breaks="4 year", date_minor_breaks="1 years", labels = NULL) +
  scale_y_continuous(name="Temp. (°C)") 
#  ggtitle("Surface time series at Point B (2007-2022)")

plot_temp_0722
plot_ano_temp2
```

### Plot of anomalies : salinity (2007-2022) - surface

```{r, plot anomalies salinity 2007-2022 - surface, echo=FALSE, warning=FALSE, message=FALSE, fig.width=8,fig.height=4,fig.align='center'}
plot_ano_sal2 <- ggplot(data = ano_0m_0722, aes(x = datetime, y = Salinity_ano), na.rm=TRUE) +
  scale_x_date(date_breaks="4 year", date_minor_breaks="1 years", labels = NULL) +
  scale_y_continuous(breaks=c(-2,-1,0,0.5)) +
  geom_point(colour="#22427C", na.rm=TRUE, size=0.7) + 
  geom_smooth(method=lm, colour="black", fill="grey", linewidth=0.6, na.rm=TRUE) +
  labs(title="",x="", y="Salinité")

plot_sal_0722 <- data_0722 %>% 
  ggplot() +
  aes(x=datetime, y=impute_sal) + 
  geom_line(size=0.7, col="#303030") +
  scale_x_date(name="", date_breaks="4 year", date_minor_breaks="1 years", labels = NULL) +
  scale_y_continuous(name="Salinité") 
#  ggtitle("Surface salinity time series at Point B (2007-2022)")

plot_sal_0722
plot_ano_sal2
```

### Plot of anomalies : pH_calc (2007-2022) - surface

```{r, plot anomalies pH 2007-2022 - surface, echo=FALSE, warning=FALSE, message=FALSE, fig.width=8,fig.height=4,fig.align='center'}

plot_ano_ph2 <- ggplot(data = ano_0m_0722, aes(x = datetime, y = pH_calc_ano), na.rm=TRUE) +
  scale_x_date(date_breaks="2 year", date_minor_breaks="1 years", labels = NULL) +
  geom_point(colour="#22427C", na.rm=TRUE, size=0.7) + 
  geom_smooth(method=lm, colour="black", fill="grey", linewidth=0.6, na.rm=TRUE) +
  labs(title="", x="", y="pH_calc")

plot_ph_0722 <- data_0722 %>% 
  ggplot() +
  aes(x=datetime, y=pH_calc) + 
  geom_line(size=0.5, col="#303030") +
  scale_x_date(name="", date_breaks="2 year", date_minor_breaks="1 years", labels = NULL) +
  scale_y_continuous(name="pH_calc") 
#  ggtitle("Time series observations at 0m of pH (2007-2022)")

plot_ph_0722
plot_ano_ph2
```

### Plot of anomalies : pH(18) (2007-2022) - surface

```{r, plot anomalies pH_18 2007-2022 - surface, echo=FALSE, warning=FALSE, message=FALSE, fig.width=8,fig.height=4,fig.align='center'}

plot_ano_ph_18_2 <- ggplot(data = ano_0m_0722, aes(x = datetime, y = pH18_ano), na.rm=TRUE) +
  scale_x_date(date_breaks="2 year", date_minor_breaks="1 years", labels = NULL) +
  geom_point(colour="#22427C", na.rm=TRUE, size=0.7) + 
  geom_smooth(method=lm, colour="black", fill="grey", linewidth=0.6, na.rm=TRUE) +
  labs(title="", x="", y="pH (18)")

plot_ph_18_0722 <- data_0722 %>% 
  ggplot() +
  aes(x=datetime, y=pH_18) + 
  geom_line(size=0.5, col="#303030") +
  scale_x_date(name="", date_breaks="2 year", date_minor_breaks="1 years", labels = NULL) +
  scale_y_continuous(name="pH (18)") 
#  ggtitle("Time series observations at 0m of pH (2007-2022)")

plot_ph_18_0722
plot_ano_ph_18_2
```

### Plot of anomalies : alkalinity (2007-2022) - surface

```{r, plot anomalies alkalinity 2007-2022 - surface, echo=FALSE, warning=FALSE, message=FALSE, fig.width=8,fig.height=4,fig.align='center'}

plot_ano_ta_0722 <- ggplot(data = ano_0m_0722, aes(x = datetime, y = ta_ano), na.rm=TRUE) +
  scale_x_date(date_breaks="2 year", date_minor_breaks="1 years", labels = NULL) +
  geom_point(colour="#22427C", na.rm=TRUE, size=0.7) + 
  geom_smooth(method=lm, colour="black", fill="grey", linewidth=0.6, na.rm=TRUE) +
  labs(title="",x="",y=expression(paste(italic(A)[T], " (",mu, "mol ", kg^-1,")")))

plot_ta_0722 <- data_0722 %>% 
  ggplot() +
  aes(x=datetime, y=impute_ta) + 
  geom_line(size=0.5, col="#303030") +
  scale_x_date(name="", date_breaks="2 year", date_minor_breaks="1 years", labels = NULL) +
  labs(title="",x="",y=expression(paste(italic(A)[T], " (",mu, "mol ", kg^-1,")")))
  


plot_ta_0722
plot_ano_ta_0722
```

### Plot of anomalies : DIC (2007-2022) - surface

```{r, plot anomalies DIC 2007-2022 - surface, echo=FALSE, warning=FALSE, message=FALSE, fig.width=8,fig.height=4,fig.align='center'}

plot_ano_dic_0722 <- ggplot(data = ano_0m_0722, aes(x = datetime, y = dic_ano), na.rm=TRUE) +
  scale_x_date(date_breaks="2 year", date_minor_breaks="1 years", labels = NULL) +
  geom_point(colour="#22427C", na.rm=TRUE, size=0.7) + 
  geom_smooth(method=lm, colour="black", fill="grey", linewidth=0.6, na.rm=TRUE) +
  labs(title="",x="",y=expression(paste(italic(C)[T], " (",mu, "mol ", kg^-1,")")))

plot_dic_0722 <- data_0722 %>% 
  ggplot() +
  aes(x=datetime, y=impute_dic) + 
  geom_line(size=0.5, col="#303030") +
  scale_x_date(name="", date_breaks="2 year", date_minor_breaks="1 years", labels = NULL) +
  labs(title="",x="",y=expression(paste(italic(C)[T], " (",mu, "mol ", kg^-1,")")))


plot_dic_0722
plot_ano_dic_0722
```

### Plot of anomalies : pCO2 (2007-2022) - surface

```{r, plot anomalies CO2 2007-2015 - surface, echo=FALSE, warning=FALSE, message=FALSE, fig.width=8,fig.height=4,fig.align='center'}

plot_ano_pCO22 <- ggplot(data = ano_0m_0722, aes(x = datetime, y = pCO2_ano), na.rm=TRUE) +
  scale_x_date(date_breaks="4 year", date_minor_breaks="1 years", labels = date_format("%Y")) +
  geom_point(colour="#22427C", na.rm=TRUE, size=0.7) + 
  geom_smooth(method=lm, colour="black", fill="grey", linewidth=0.6, na.rm=TRUE) +
  labs(title="",x="",y=expression(paste(italic(p), CO[2], " (",mu, "atm)")))

plot_pCO2_0722 <- data_0722 %>% 
  ggplot() +
  aes(x=datetime, y=impute_pCO2_w) + 
  geom_line(size=0.7, col="#303030") +
  scale_x_date(date_breaks="4 year", date_minor_breaks="1 years", labels = date_format("%Y")) +
  labs(title="",x="",y=expression(paste(italic(p), CO[2], " (",mu, "atm)")))
#  ggtitle("Time series observations at 0m of pCO2 (2007-2022)")

plot_pCO2_0722
plot_ano_pCO22
```

```{r plot pour rapport, echo=FALSE, warning=FALSE, message=FALSE, fig.width=9,fig.height=5}

plot_grid_1 <- plot_grid(plot_ph_18_0722, plot_ano_ph_18_2, plot_ta_0722, plot_ano_ta_0722, 
                         plot_dic_0722, plot_ano_dic_0722, align='vh', ncol=2)

plot_grid_1


```

## DEPTH 50 m

### [Table of monthly means (2007-2022) - 50m data]{style="color:blue"}

*(rajouter les SD)*

```{r calculate monthly means 2007-2022 - 50m, echo=FALSE, warning=FALSE, message=FALSE, fig.width=8,fig.height=10}

monthly_means_50m_0722 <- ungroup(data_0722_50m) %>% 
  group_by(Month) %>%
  summarise(
    Salinity_month = mean(impute_sal, na.rm = TRUE),
    SD_sal = sd(impute_sal, na.rm = TRUE),
    Temperature_month = mean(impute_temp, na.rm = TRUE),
    SD_temp = sd(impute_temp, na.rm = TRUE),
    dic_month = mean(impute_dic, na.rm = TRUE),
    SD_dic = sd(impute_dic, na.rm = TRUE),
    ta_month = mean(impute_ta, na.rm = TRUE),
    SD_ta = sd(impute_ta, na.rm = TRUE),
    pH_calc_month = mean(pH_calc, na.rm = TRUE),
    SD_PH = sd(pH_calc, na.rm = TRUE),
    pH18_month = mean(pH_18, na.rm = TRUE),
    SD_PH18 = sd(pH_18, na.rm = TRUE),
    pCO2_month = mean(impute_pco2_w, na.rm = TRUE),
    SD_pCO2 = sd(impute_pco2_w, na.rm = TRUE))

monthly_means_50m_0722


##RAPPORT##

# monthly_means_50m_0722_rapport <- monthly_means_50m_0722 %>% 
#   mutate(`T (°C)` = paste(round(Temperature_month, digits=2), "±", round(SD_temp, digits=2)),
#          S = paste(round(Salinity_month, digits=2), "±", round(SD_sal, digits=2)),
#         `pH (18)` = paste(round(pH18_month, digits=2), "±", round(SD_PH18, digits=2)),
#         `*A*~T~ (μmol.kg^-1^)` = paste(round(ta_month, digits=2), "±", round(SD_ta, digits=2)),
#         `*C*~T~ (μmol.kg^-1^)` = paste(round(dic_month, digits=2), "±", round(SD_dic, digits=2)),
#         `*p*CO~2~ (μatm)` = paste(round(pCO2_month, digits=2), "±", round(SD_pCO2, digits=2)))
# 
# monthly_means_50m_0722_rapport <- monthly_means_50m_0722_rapport %>% 
#   dplyr::select(Month, `T (°C)`, S, `pH (18)`, `*A*~T~ (μmol.kg^-1^)`, 
#                 `*C*~T~ (μmol.kg^-1^)`, `*p*CO~2~ (μatm)`) %>% 
#   rename(Mois = Month)
#         
#         
# kable(monthly_means_50m_0722_rapport, caption="Moyennes mensuelles au Point B (50 m)",
#       format = "pipe", align='l')

```

### [Table of time series anomaly regression analyses (2007-2022) - 50m data]{style="color:green"}

```{r calculate anomalies + regression 2007-2022 - 50m, echo=FALSE, warning=FALSE, message=FALSE}

ano_50m_0722 <- left_join(ungroup(data_0722_50m), monthly_means_50m_0722, by = "Month") %>%
  mutate(
    Salinity_ano = impute_sal - Salinity_month,
    Temperature_ano = impute_temp - Temperature_month,
    dic_ano = impute_dic - dic_month,
    ta_ano = impute_ta - ta_month,
    pH_calc_ano = pH_calc - pH_calc_month,
    pH18_ano = pH_18 - pH18_month,
    pCO2_ano = impute_pco2_w - pCO2_month)


# Regression and table anomalies
var_list <-
  c("Temperature_ano",
    "Salinity_ano",
    "pH_calc_ano",
    "pH18_ano",
    "ta_ano",
    "dic_ano",
    "pCO2_ano"
  )
lms <- lapply(var_list, function(x) {
  summary(lm(substitute(i ~ decimal_date(datetime), list(i = as.name(x))), data = ano_50m_0722))
})
reg <- NULL
for (i in 1:length(var_list)) {
  #one loops through all anomalies
  # calculate probability of fstatistic because it cannot be extracted from lms above
  # see http://stats.stackexchange.com/questions/92824/extracting-the-model-p-value-for-a-multiple-regression-in-r
  # slope returned by lm is per second
  prob <-
    pf(lms[[i]]$fstatistic[1],
       lms[[i]]$fstatistic[2],
       lms[[i]]$fstatistic[3],
       lower.tail = FALSE)
  reg <-
    rbind(reg, as.numeric(
      c(
        lms[[i]]$coefficients[2, 1],
        lms[[i]]$coefficients[2, 2],
        lms[[i]]$coefficients[2, 4],
        lms[[i]]$coefficients[1, 1],
        lms[[i]]$coefficients[1, 2],
        lms[[i]]$coefficients[1, 4],
        lms[[i]]$fstatistic[1],
        lms[[i]]$fstatistic[3],
        lms[[i]]$r.squared,
        prob
      )
    ))
}
colnames(reg) <-
  c("Slope",
    "SE Slope",
    "P Slope",
    "Intercept",
    "SE int.",
    "P int.",
    "F",
    "df",
    "R2",
    "P value")
row.names(reg) <- var_list
reg_ano_50m <- reg

reg_ano_50m

##RAPPORT##

# reg_ano_50m_rapport <- as.data.frame(reg_ano_50m)
# reg_ano_50m_rapport <- reg_ano_50m_rapport %>% 
#   mutate(Variables = c("T (°C)", "S", "pH~calc~", "pH~(18)~", "*A*~T~ (μmol.kg^-1^)",
#                        "*C*~T~ (μmol.kg^-1^)", "*p*CO~2~ (μatm)"))
# 
# reg_ano_50m_rapport <- reg_ano_50m_rapport %>% 
#   mutate( 
#   `Pente ± sd` = paste(round(reg_ano_50m_rapport$Slope, digits=3), "±", round(reg_ano_50m_rapport$`SE Slope`, digits=4)), 
#   `P pente` = format(reg_ano_50m_rapport$`P Slope`, scientific=TRUE),
#   `Int. ± sd` = paste(round(reg_ano_50m_rapport$Intercept, digits=2), "±", round(reg_ano_50m_rapport$`SE int.`, digits=2)),
#   `P Int.` = format(reg_ano_50m_rapport$`P int.`,scientific=TRUE), 
#   `R^2^` = round(reg_ano_50m_rapport$R2, digits=4))
# 
# 
# reg_ano_50m_rapport <- reg_ano_50m_rapport[,c(11:16)]
# 
# kable(reg_ano_50m_rapport, caption="Time series anomaly regression analyses (Point B 50 m)", 
#       format = "pipe", align='l')

```

### Plot of anomalies : Temperature (2007-2022) - 50m

```{r, plot anomalies temperature 2007-2022 - 50m, echo=FALSE, warning=FALSE, message=FALSE, fig.width=8,fig.height=4,fig.align='center'}

plot_ano_temp2_50m <- ggplot(data = ano_50m_0722, aes(x = datetime, y = Temperature_ano), na.rm=TRUE) +
  scale_x_date(date_breaks="4 year", date_minor_breaks="1 years", labels = NULL) +
  geom_point(colour="#22427C", na.rm=TRUE, size=0.7) + 
  geom_smooth(method=lm, colour="black", fill="grey", linewidth=0.6, na.rm=TRUE) +
  labs(title="", x="", y="Temp. (°C)")

plot_temp_0722_50m <- data_0722_50m %>% 
  ggplot() +
  aes(x=datetime, y=impute_temp) + 
  geom_line(size=0.7, col="#303030") +
  scale_x_date(name="", date_breaks="4 year", date_minor_breaks="1 years", labels = NULL) +
  scale_y_continuous(name="Temp. (°C)") 
#  ggtitle("50m temperature time series (2007-2022)")

plot_temp_0722_50m
plot_ano_temp2_50m
```

### Plot of anomalies : salinity (2007-2022) - 50m

```{r, plot anomalies salinity 2007-2022 - 50m, echo=FALSE, warning=FALSE, message=FALSE, fig.width=8,fig.height=4,fig.align='center'}

plot_ano_sal2_50m <- ggplot(data = ano_50m_0722, aes(x = datetime, y = Salinity_ano), na.rm=TRUE) +
  scale_x_date(date_breaks="4 year", date_minor_breaks="1 years", labels = NULL) +
  scale_y_continuous(breaks=c(-0.50,-0.25,0,0.25, 0.50)) +
  geom_point(colour="#22427C", na.rm=TRUE, size=0.7) + 
  geom_smooth(method=lm, colour="black", fill="grey", linewidth=0.6, na.rm=TRUE) +
  labs(title="",x="", y="Salinity") 

plot_sal_0722_50m <- data_0722_50m %>% 
  ggplot() +
  aes(x=datetime, y=impute_sal) + 
  geom_line(size=0.7, col="#303030") +
  scale_x_date(name="", date_breaks="4 year", date_minor_breaks="1 years", labels = NULL) +
  scale_y_continuous(name="Salinity") 
#  ggtitle("50m salinity time series (2007-2022)")

plot_sal_0722_50m
plot_ano_sal2_50m
```

### Plot of anomalies : pH_calc (2007-2022) - 50m

```{r, plot anomalies pH 2007-2022 - 50m, echo=FALSE, warning=FALSE, message=FALSE, fig.width=8,fig.height=4,fig.align='center'}

plot_ano_ph2_50m <- ggplot(data = ano_50m_0722, aes(x = datetime, y = pH_calc_ano), na.rm=TRUE) +
  scale_x_date(date_breaks="2 year", date_minor_breaks="1 years", labels = NULL) +
  geom_point(colour="#22427C", na.rm=TRUE, size=0.7) + 
  geom_smooth(method=lm, colour="black", fill="grey", linewidth=0.6, na.rm=TRUE) +
  labs(title="", x="", y="pH calc")

plot_ph_0722_50m <- data_0722_50m %>% 
  ggplot() +
  aes(x=datetime, y=pH_calc) + 
  geom_line(size=0.5, col="#303030") +
  scale_x_date(name="", date_breaks="2 year", date_minor_breaks="1 years", labels = NULL) +
  scale_y_continuous(name="pH calc") 
#  ggtitle("Time series observations at 50m of pH (2007-2022)")

plot_ph_0722_50m
plot_ano_ph2_50m
```

### Plot of anomalies : pH(18) (2007-2022) - 50m

```{r, plot anomalies pH_18 2007-2022 - 50m, echo=FALSE, warning=FALSE, message=FALSE, fig.width=8,fig.height=4,fig.align='center'}

plot_ano_ph_18_2_50m <- ggplot(data = ano_50m_0722, aes(x = datetime, y = pH18_ano), na.rm=TRUE) +
  scale_x_date(date_breaks="2 year", date_minor_breaks="1 years", labels = NULL) +
  geom_point(colour="#22427C", na.rm=TRUE, size=0.7) + 
  geom_smooth(method=lm, colour="black", fill="grey", linewidth=0.6, na.rm=TRUE) +
  labs(title="", x="", y="pH (18)")

plot_ph_18_0722_50m <- data_0722_50m %>% 
  ggplot() +
  aes(x=datetime, y=pH_18) + 
  geom_line(size=0.5, col="#303030") +
  scale_x_date(name="", date_breaks="2 year", date_minor_breaks="1 years", labels = NULL) +
  scale_y_continuous(name="pH (18)") 
#  ggtitle("Time series observations at 50m of pH (2007-2022)")

plot_ph_18_0722_50m
plot_ano_ph_18_2_50m
```

### Plot of anomalies : alkalinity (2007-2022) - 50m

```{r, plot anomalies alcalinity 2007-2022 - 50m, echo=FALSE, warning=FALSE, message=FALSE, fig.width=8,fig.height=4,fig.align='center'}

plot_ano_ta_0722_50m <- ggplot(data = ano_50m_0722, aes(x = datetime, y = ta_ano), na.rm=TRUE) +
  scale_x_date(date_breaks="2 year", date_minor_breaks="1 years", labels = NULL) +
  geom_point(colour="#22427C", na.rm=TRUE, size=0.7) + 
  geom_smooth(method=lm, colour="black", fill="grey", linewidth=0.6, na.rm=TRUE) +
  labs(title="",x="",y=expression(paste(italic(A)[T], " (",mu, "mol ", kg^-1,")")))

plot_ta_0722_50m <- data_0722_50m %>% 
  ggplot() +
  aes(x=datetime, y=impute_ta) + 
  geom_line(size=0.5, col="#303030") +
  scale_x_date(name="", date_breaks="2 year", date_minor_breaks="1 years", labels = NULL) +
  labs(title="",x="",y=expression(paste(italic(A)[T], " (",mu, "mol ", kg^-1,")")))


plot_ta_0722_50m
plot_ano_ta_0722_50m

```

### Plot of anomalies : DIC (2007-2022) - 50m

```{r, plot anomalies DIC 2007-2022 - 50m, echo=FALSE, warning=FALSE, message=FALSE, fig.width=8,fig.height=4,fig.align='center'}

plot_ano_dic_0722_50m <- ggplot(data = ano_50m_0722, aes(x = datetime, y = dic_ano), na.rm=TRUE) +
  scale_x_date(date_breaks="2 year", date_minor_breaks="1 years", labels = NULL) +
  geom_point(colour="#22427C", na.rm=TRUE, size=0.7) + 
  geom_smooth(method=lm, colour="black", fill="grey", linewidth=0.6, na.rm=TRUE) +
  labs(title="",x="",y=expression(paste(italic(C)[T], " (",mu, "mol ", kg^-1,")")))

plot_dic_0722_50m <- data_0722_50m %>% 
  ggplot() +
  aes(x=datetime, y=impute_dic) + 
  geom_line(size=0.5, col="#303030") +
  scale_x_date(name="", date_breaks="2 year", date_minor_breaks="1 years", labels = NULL) +
  labs(title="",x="",y=expression(paste(italic(C)[T], " (",mu, "mol ", kg^-1,")")))


plot_dic_0722_50m
plot_ano_dic_0722_50m
```

### Plot of anomalies : pCO2 (2007-2022) - 50m

```{r, plot anomalies CO2 2007-2022 - 50m, echo=FALSE, warning=FALSE, message=FALSE, fig.width=8,fig.height=4,fig.align='center'}

plot_ano_pCO22_50m <- ggplot(data = ano_50m_0722, aes(x = datetime, y = pCO2_ano), na.rm=TRUE) +
  scale_x_date(date_breaks="4 year", date_minor_breaks="1 years", labels = date_format("%Y")) +
  geom_point(colour="#22427C", na.rm=TRUE, size=0.7) + 
  geom_smooth(method=lm, colour="black", fill="grey", linewidth=0.6, na.rm=TRUE) +
  labs(title="",x="",y=expression(paste(pCO[2], " (",mu, "atm)")))

plot_pCO2_0722_50m <- data_0722_50m %>% 
  ggplot() +
  aes(x=datetime, y=impute_pco2_w) + 
  geom_line(size=0.7, col="#303030") +
  scale_x_date(name="", date_breaks="4 year", date_minor_breaks="1 years", labels = date_format("%Y")) +
  scale_y_continuous(name="pCO2 (μatm)") 
#  ggtitle("Time series observations at 50m of pCO2 (2007-2022)")

plot_pCO2_0722_50m
plot_ano_pCO22_50m
```

```{r rapport_50m, echo=FALSE, warning=FALSE, message=FALSE, fig.width=12,fig.height=11}

plot_grid_2 <- plot_grid(plot_temp_0722_50m, plot_ano_temp2_50m, plot_sal_0722_50m, plot_ano_sal2_50m, 
                         plot_ph_18_0722_50m, plot_ano_ph_18_2_50m, plot_ta_0722_50m, plot_ano_ta_0722_50m, 
                         plot_dic_0722_50m, plot_ano_dic_0722_50m, plot_pCO2_0722_50m, plot_ano_pCO22_50m,
                         align='vh', ncol=2)

plot_grid_2

 # ggsave(filename = "kapsenberg_07-22_50m.jpg", # Nommez le fichier dans lequel vous voulez enregistrer, ajoutez l'extension du format de fichier que vous voulez utiliser (ex. pdf).
 #          plot = plot_grid_2, # Fournir le nom de l'objet plot dans R
 #          height = 11, # Fournir les dimensions voulues
 #          width = 10, 
 #          units = "in") 

```

# 1995-2022 : temperature & salinity

05-92 to 12-22 but not enough data before Jan. 1995

## SURFACE

```{r import data temperature + salinity Point B - 0m et 50m, echo=FALSE, warning=FALSE, message=FALSE}

TS_9522_surf <- read_delim("../../Data/data_PointB/PointB_T_S_1995-2022_CLEAN_impute_0m.csv", 
    delim = ";", escape_double = FALSE, trim_ws = TRUE) %>% 
  mutate(month = format(datetime, format = "%m"))


```

### [Table of monthly means for salinity and temperature (1995-2022) - 0m data]{style="color:blue"}

```{r calculate monthly means 1992-2022 - 0m, echo=FALSE, warning=FALSE, message=FALSE}

monthly_means_0m_9522 <- TS_9522_surf %>% 
  group_by(month) %>%
  summarise(Salinity_month = mean(impute_sal, na.rm = TRUE),
            SD_sal = sd(impute_sal, na.rm = TRUE),
            Temperature_month = mean(impute_temp, na.rm = TRUE),
            SD_temp = sd(impute_temp, na.rm = TRUE))

monthly_means_0m_9522

```

### [Table of salinity and temperature time series anomaly regression analyses (1995-2022) - 0m data]{style="color:green"}

```{r calculate anomalies + regression on salinity and temperature (1992-2022) - 0m, echo=FALSE, warning=FALSE, message=FALSE}

ano_ST_0m_9522 <- left_join(ungroup(TS_9522_surf), monthly_means_0m_9522, by = "month") %>%
  mutate(
    Salinity_ano = impute_sal - Salinity_month,
    Temperature_ano = impute_temp - Temperature_month)


# Regression and table anomalies
var_list_ST2 <-
  c(
    "Salinity_ano",
    "Temperature_ano")

lms_ST2 <- lapply(var_list_ST2, function(x) {
  summary(lm(substitute(i ~ decimal_date(datetime), list(i = as.name(x))), data = ano_ST_0m_9522))
})

reg_ST2 <- NULL
for (i in 1:length(var_list_ST2)) {
  #one loops through all anomalies
  # calculate probability of fstatistic because it cannot be extracted from lms above
  # see http://stats.stackexchange.com/questions/92824/extracting-the-model-p-value-for-a-multiple-regression-in-r
  # slope returned by lm is per second
  prob_ST2 <-
    pf(lms_ST2[[i]]$fstatistic[1],
       lms_ST2[[i]]$fstatistic[2],
       lms_ST2[[i]]$fstatistic[3],
       lower.tail = FALSE)
  reg_ST2 <-
    rbind(reg_ST2, as.numeric(
      c(
        lms_ST2[[i]]$coefficients[2, 1],
        lms_ST2[[i]]$coefficients[2, 2],
        lms_ST2[[i]]$coefficients[2, 4],
        lms_ST2[[i]]$coefficients[1, 1],
        lms_ST2[[i]]$coefficients[1, 2],
        lms_ST2[[i]]$coefficients[1, 4],
        lms_ST2[[i]]$fstatistic[1],
        lms_ST2[[i]]$fstatistic[3],
        lms_ST2[[i]]$r.squared,
        prob_ST2
      )
    ))
}
colnames(reg_ST2) <-
  c("Slope",
    "SE Slope",
    "P Slope",
    "Intercept",
    "SE int.",
    "P int.",
    "F",
    "df",
    "R2",
    "P value")
row.names(reg_ST2) <- var_list_ST2
reg_ano_ST_0m_9522 <- reg_ST2

reg_ano_ST_0m_9522

```

### Plot of anomalies : Temperature (1995-2022) - 0m

```{r plot anomalies temperature 1995-2022 - 0m, echo=FALSE, warning=FALSE, message=FALSE, fig.width=8,fig.height=4,fig.align='center'}

plot_ano_temp_9522 <- ggplot(data = ano_ST_0m_9522, aes(x = datetime, y = Temperature_ano), na.rm=TRUE) +
  #scale_x_date(date_breaks="4 year", date_minor_breaks="1 years", labels = NULL) +
  geom_point(colour="#22427C", na.rm=TRUE, size=0.7) + 
  geom_smooth(method=lm, colour="black", fill="grey", linewidth=0.6, na.rm=TRUE) +
  labs(title="",x="Year", y="Temp. (°C)")


plot_temp_9522 <- TS_9522_surf %>% 
  ggplot() +
  aes(x=datetime, y=impute_temp) + 
  geom_line(size=0.5, col="#303030") +
  #scale_x_date(name="", date_breaks="4 year", date_minor_breaks="1 years", labels = NULL) +
  labs(title="",x="Year", y="Temp. (°C)")

plot_temp_9522
plot_ano_temp_9522

```

### Plot of anomalies : salinity (1995-2022) - 0m

```{r plot anomalies salinity 1992-2022 - 0m, echo=FALSE, warning=FALSE, message=FALSE, fig.width=9,fig.height=16,fig.align='center'}

plot_ano_sal_9522 <- ggplot(data = ano_ST_0m_9522, aes(x = datetime, y = Salinity_ano), na.rm=TRUE) +
  #scale_x_date(date_breaks="4 year", date_minor_breaks="1 years", date_labels = "%Y") +
  scale_y_continuous(breaks=c(-2,-1,0,0.5)) +
  geom_point(colour="#22427C", na.rm=TRUE, size=0.7) + 
  geom_smooth(method=lm, colour="black", fill="grey", linewidth=0.6, na.rm=TRUE) +
  labs(title="",x="", y="Salinity")

plot_sal_9522 <- TS_9522_surf %>% 
  ggplot() +
  aes(x=datetime, y=impute_sal) + 
  geom_line(size=0.5, col="#303030") +
  #scale_x_date(name="", date_breaks="4 year", date_minor_breaks="1 years", date_labels = "%Y") +
  labs(title="",x="Year", y="Salinity")

plot_sal_9522
plot_ano_sal_9522

 # plot_grid(plot_temp_9522, plot_ano_temp_9522, 
 #           plot_sal_9522, plot_ano_sal_9522, plot_AOU_surf, plot_ano_AOU_surf,
 #           plot_ph_0722, plot_ano_ph2, plot_ph_18_0722, plot_ano_ph_18_2, 
 #           plot_ta_0722, plot_ano_ta_0722, plot_dic_0722, plot_ano_dic_0722,
 #           plot_pco2_therm, plot_ano_pCO2therm, plot_pco2_nontherm, plot_ano_pCO2nontherm,
 #           align='vh', ncol=2)

```

## DEPTH 50

### [Table of monthly means for salinity and temperature (1995-2022) - 50m data]{style="color:blue"}

```{r calculate monthly means 1995-2022 - 50m, echo=FALSE, warning=FALSE, message=FALSE}

TS_9522_prof <- read_delim("../../Data/data_PointB/PointB_T_S_1995-2022_CLEAN_impute_50m.csv", 
    delim = ";", escape_double = FALSE, trim_ws = TRUE) %>% 
  mutate(month = format(datetime, format = "%m"),
         impute_sal = case_when(impute_sal < 37 ~ NA_real_, TRUE ~ salinity))

monthly_means_50m_9522 <- TS_9522_prof %>% 
  group_by(month) %>%
  summarise(Salinity_month = mean(impute_sal, na.rm = TRUE),
            SD_sal = sd(impute_sal, na.rm = TRUE),
            Temperature_month = mean(impute_temp, na.rm = TRUE),
            SD_temp = sd(impute_temp, na.rm = TRUE))

monthly_means_50m_9522

```

### [Table of salinity and temperature time series anomaly regression analyses (1995-2022) - 50m data]{style="color:green"}

```{r calculate anomalies + regression on salinity and temperature (1995-2022) - 50m, echo=FALSE, warning=FALSE, message=FALSE}

ano_ST_50m_9522 <- left_join(ungroup(TS_9522_prof), monthly_means_50m_9522, by = "month") %>%
  mutate(
    Salinity_ano = impute_sal - Salinity_month,
    Temperature_ano = impute_temp - Temperature_month)


# Regression and table anomalies
var_list_ST2 <-
  c(
    "Salinity_ano",
    "Temperature_ano")

lms_ST2 <- lapply(var_list_ST2, function(x) {
  summary(lm(substitute(i ~ decimal_date(datetime), list(i = as.name(x))), data = ano_ST_50m_9522))
})

reg_ST2 <- NULL
for (i in 1:length(var_list_ST2)) {
  #one loops through all anomalies
  # calculate probability of fstatistic because it cannot be extracted from lms above
  # see http://stats.stackexchange.com/questions/92824/extracting-the-model-p-value-for-a-multiple-regression-in-r
  # slope returned by lm is per second
  prob_ST2 <-
    pf(lms_ST2[[i]]$fstatistic[1],
       lms_ST2[[i]]$fstatistic[2],
       lms_ST2[[i]]$fstatistic[3],
       lower.tail = FALSE)
  reg_ST2 <-
    rbind(reg_ST2, as.numeric(
      c(
        lms_ST2[[i]]$coefficients[2, 1],
        lms_ST2[[i]]$coefficients[2, 2],
        lms_ST2[[i]]$coefficients[2, 4],
        lms_ST2[[i]]$coefficients[1, 1],
        lms_ST2[[i]]$coefficients[1, 2],
        lms_ST2[[i]]$coefficients[1, 4],
        lms_ST2[[i]]$fstatistic[1],
        lms_ST2[[i]]$fstatistic[3],
        lms_ST2[[i]]$r.squared,
        prob_ST2
      )
    ))
}
colnames(reg_ST2) <-
  c("Slope",
    "SE Slope",
    "P Slope",
    "Intercept",
    "SE int.",
    "P int.",
    "F",
    "df",
    "R2",
    "P value")
row.names(reg_ST2) <- var_list_ST2
reg_ano_ST_50m_9522 <- reg_ST2

reg_ano_ST_50m_9522

```

### Plot of anomalies : Temperature (1995-2022) - 50m

```{r plot anomalies temperature 1995-2022 - 50m, echo=FALSE, warning=FALSE, message=FALSE, fig.width=8,fig.height=4,fig.align='center'}

plot_ano_temp_9522_50m <- ggplot(data = ano_ST_50m_9522, aes(x = datetime, y = Temperature_ano), na.rm=TRUE) +
  #scale_x_date(date_breaks="4 year", date_minor_breaks="1 years", labels = NULL) +
  geom_point(colour="#22427C", na.rm=TRUE, size=0.7) + 
  geom_smooth(method=lm, colour="black", fill="grey", linewidth=0.6, na.rm=TRUE) +
  labs(title="", x="Year", y="Temperature (°C)")

plot_temp_9522_50m <- TS_9522_prof %>% 
  ggplot() +
  aes(x=datetime, y=impute_temp) + 
  geom_line(size=0.5) +
  #scale_x_date(name="", date_breaks="4 year", date_minor_breaks="1 years", labels = NULL) +
  labs(title="",x="Year", y="Temperature (°C)")

plot_temp_9522_50m
plot_ano_temp_9522_50m

```

### Plot of anomalies : salinity (1995-2022) - 50m

```{r plot anomalies salinity 1995-2022 - 50m, echo=FALSE, warning=FALSE, message=FALSE, fig.width=9,fig.height=16,fig.align='center'}

plot_ano_sal_9522_prof <- ggplot(data = ano_ST_50m_9522, aes(x = datetime, y = Salinity_ano), na.rm=TRUE) +
  #scale_x_date(date_breaks="4 year", date_minor_breaks="1 years", date_labels = "%Y") +
  scale_y_continuous(breaks=c(-2,-1,0,0.5), limits=c(-2,0.5)) +
  geom_point(colour="#22427C", na.rm=TRUE, size=0.7) + 
  geom_smooth(method=lm, colour="black", fill="grey", linewidth=0.6, na.rm=TRUE) +
  labs(title="",x="Year", y="Salinity")

plot_sal_9522_prof <- TS_9522_prof %>% 
  ggplot() +
  aes(x=datetime, y=impute_sal) + 
  geom_line(size=0.5, col="#303030") +
  #scale_x_date(name="", date_breaks="4 year", date_minor_breaks="1 years", date_labels = "%Y") +
  scale_y_continuous(name="Salinity", limits=c(35, 39)) 

plot_sal_9522_prof
plot_ano_sal_9522_prof

 # plot_grid(plot_temp_9522_50m, plot_ano_temp_9522_50m, plot_sal_9522_prof, plot_ano_sal_9522_prof,
 #           plot_AOU_prof, plot_ano_AOU_prof, plot_ph_0722_50m, plot_ano_ph2_50m, 
 #           plot_ph_18_0722_50m, plot_ano_ph_18_2_50m, plot_ta_0722_50m, plot_ano_ta_0722_50m,
 #           plot_dic_0722_50m, plot_ano_dic_0722_50m, plot_pco2_therm_prof, plot_ano_pCO2therm_prof, 
 #           plot_pco2_nontherm_prof, plot_ano_pCO2nontherm_prof,
 #           align='vh', ncol=2)

```

# 2000-2022 : Analysis every 10 years

## SURFACE

### Jan.2000 - Dec.2009

#### [Table of monthly means (2000-2009) - 0m data]{style="color:blue"}

```{r calculate monthly means 2000-2009 - 0m, echo=FALSE, warning=FALSE, message=FALSE}

Data_SOMLIT_2000_surf <- TS_9522_surf %>% 
  dplyr::filter(datetime >= "2000-01-05" & datetime <= "2009-12-22")%>% 
  mutate(Month = format(datetime, format="%m"))
##

monthly_means_0m_2000 <- Data_SOMLIT_2000_surf %>% 
  group_by(month) %>%
  summarise(Salinity_month = mean(impute_sal),
            SD_sal = sd(impute_sal), 
            Temperature_month = mean(impute_temp),
            SD_temp = sd(impute_temp))

monthly_means_0m_2000

```

#### [Table of anomaly regression analyses (2000-2009) - 0m data]{style="color:green"}

```{r calculate anomalies + regression (2000-2009) - 0m, echo=FALSE, warning=FALSE, message=FALSE}

ano_0m_2000 <- left_join(ungroup(Data_SOMLIT_2000_surf), monthly_means_0m_2000, by = "month") %>%
  mutate(
    Salinity_ano = impute_sal - Salinity_month,
    Temperature_ano = impute_temp - Temperature_month)

# Regressions and tables of key parameters
var_list <-
  c("salinity",
    "temperature")

lm.test <- vector("list", length(var_list))

for(i in seq_along(var_list)){
  lm.test[[i]] <- lm(reformulate(var_list[i], "datetime"), data = ano_0m_2000)
}

#cutoff_date <- "2016-01-01" #to limit regressions on complete years

lms <- lapply(var_list, function(x) {
  summary(lm(substitute(i ~ decimal_date(datetime), list(i = as.name(x))), 
             data = ano_0m_2000))
})

reg <- NULL
for (i in 1:length(var_list)) {
  #one loops through all anomalies
  # calculate probability of fstatistic because it cannot be extracted from lms above
  # see http://stats.stackexchange.com/questions/92824/extracting-the-model-p-value-for-a-multiple-regression-in-r
  # slope returned by lm is per second
  prob <-
    pf(lms[[i]]$fstatistic[1],
       lms[[i]]$fstatistic[2],
       lms[[i]]$fstatistic[3],
       lower.tail = FALSE)
  reg <-
    rbind(reg, as.numeric(
      c(
        lms[[i]]$coefficients[2, 1],
        lms[[i]]$coefficients[2, 2],
        lms[[i]]$coefficients[2, 4],
        lms[[i]]$coefficients[1, 1],
        lms[[i]]$coefficients[1, 2],
        lms[[i]]$coefficients[1, 4],
        lms[[i]]$fstatistic[1],
        lms[[i]]$fstatistic[3],
        lms[[i]]$r.squared,
        prob
      )
    ))
}

colnames(reg) <-
  c("Slope",
    "SE Slope",
    "P Slope",
    "Intercept",
    "SE int.",
    "P int.",
    "F",
    "df",
    "R2",
    "P value")
row.names(reg) <- var_list
reg_ano_0m_2000 <- reg

# Regression and table anomalies
var_list <-
  c(
    "Salinity_ano",
    "Temperature_ano")

lms <- lapply(var_list, function(x) {
  summary(lm(substitute(i ~ decimal_date(datetime), list(i = as.name(x))), data = ano_0m_2000))
})

reg <- NULL
for (i in 1:length(var_list)) {
  #one loops through all anomalies
  # calculate probability of fstatistic because it cannot be extracted from lms above
  # see http://stats.stackexchange.com/questions/92824/extracting-the-model-p-value-for-a-multiple-regression-in-r
  # slope returned by lm is per second
  prob <-
    pf(lms[[i]]$fstatistic[1],
       lms[[i]]$fstatistic[2],
       lms[[i]]$fstatistic[3],
       lower.tail = FALSE)
  reg <-
    rbind(reg, as.numeric(
      c(
        lms[[i]]$coefficients[2, 1],
        lms[[i]]$coefficients[2, 2],
        lms[[i]]$coefficients[2, 4],
        lms[[i]]$coefficients[1, 1],
        lms[[i]]$coefficients[1, 2],
        lms[[i]]$coefficients[1, 4],
        lms[[i]]$fstatistic[1],
        lms[[i]]$fstatistic[3],
        lms[[i]]$r.squared,
        prob
      )
    ))
}
colnames(reg) <-
  c("Slope",
    "SE Slope",
    "P Slope",
    "Intercept",
    "SE int.",
    "P int.",
    "F",
    "df",
    "R2",
    "P value")
row.names(reg) <- var_list
reg_ano_0m_2000 <- reg

reg_ano_0m_2000

```

#### Plot of anomalies : Temperature (2000-2009) - 0m

```{r plot anomalies temperature 2000-2009 - 0m, echo=FALSE, warning=FALSE, message=FALSE, fig.width=8,fig.height=4,fig.align='center'}

plot_ano_temp_2000_0m <- ggplot(data = ano_0m_2000, aes(x = datetime, y = Temperature_ano), na.rm=TRUE) +
  #scale_x_datetime(date_breaks="2 year", labels = date_format("%Y")) +
  geom_point(colour="blue", na.rm=TRUE, size=1) + 
  geom_smooth(method=lm, colour="black", fill="grey", linewidth=0.6, na.rm=TRUE) +
  labs(title="Temperature anomalies + trend (2000-2009)",x="", y="Temp. (°C)")

plot_temp_2000_0m <- Data_SOMLIT_2000_surf %>% 
  ggplot() +
  aes(x=datetime, y=impute_temp) + 
  geom_line(size=0.7, col = "#24445C") +
  #scale_x_datetime(name="") +
  scale_y_continuous(name="Temp. (°C)") +
  ggtitle("Surface temperature time series (2000-2009)")

plot_temp_2000_0m
plot_ano_temp_2000_0m

```

#### Plot of anomalies : Salinity (2000-2009) - 0m

```{r plot anomalies salinity 2000-2009 - 0m, echo=FALSE, warning=FALSE, message=FALSE, fig.width=8,fig.height=4,fig.align='center'}

plot_ano_sal_2000_0m <- ggplot(data = ano_0m_2000, aes(x = datetime, y = Salinity_ano), na.rm=TRUE) +
  #scale_x_datetime(date_breaks="2 year", labels = date_format("%Y")) +
  geom_point(colour="blue", na.rm=TRUE, size=1) + 
  geom_smooth(method=lm, colour="black", fill="grey", linewidth=0.6, na.rm=TRUE) +
  labs(title="Salinity anomalies + trend (2000-2009)",x="Year", y="Salinity")

plot_sal_2000_0m <- Data_SOMLIT_2000_surf %>% 
  ggplot() +
  aes(x=datetime, y=impute_sal) + 
  geom_line(size=0.7, col = "#24445C") +
  #scale_x_datetime(name="") +
  scale_y_continuous(name="Salinity") +
  labs(title="Salinity anomalies + trend (2000-2009)",x="Year", y="Salinity")

plot_sal_2000_0m
plot_ano_sal_2000_0m

```

### Jan.2010 - Dec.2022

#### [Table of monthly means (2010-2022) - 0m data]{style="color:blue"}

```{r calculate monthly means 2010-2022 - 0m, echo=FALSE, warning=FALSE, message=FALSE}

Data_SOMLIT_2010_surf <- TS_9522_surf %>% 
  dplyr::filter(datetime >= "2010-01-04")
##

monthly_means_0m_2010 <- Data_SOMLIT_2010_surf %>% 
  group_by(month) %>%
  summarise(Salinity_month = mean(impute_sal),
            SD_sal = sd(impute_sal),
            Temperature_month = mean(impute_temp),
            SD_temp = sd(impute_temp))

monthly_means_0m_2010

```

#### [Table of anomaly regression analyses (2010-2022) - 0m data]{style="color:green"}

```{r calculate anomalies + regression (2010-2022) - 0m, echo=FALSE, warning=FALSE, message=FALSE}

ano_0m_2010 <- left_join(ungroup(Data_SOMLIT_2010_surf), monthly_means_0m_2010, by = "month") %>%
  mutate(
    Salinity_ano = impute_sal - Salinity_month,
    Temperature_ano = impute_temp - Temperature_month)

# Regressions and tables of key parameters
var_list <-
  c("salinity",
    "temperature")

lm.test <- vector("list", length(var_list))

for(i in seq_along(var_list)){
  lm.test[[i]] <- lm(reformulate(var_list[i], "datetime"), data = ano_0m_2010)
}

#cutoff_date <- "2016-01-01" #to limit regressions on complete years

lms <- lapply(var_list, function(x) {
  summary(lm(substitute(i ~ decimal_date(datetime), list(i = as.name(x))), 
             data = ano_0m_2010))
})

reg <- NULL
for (i in 1:length(var_list)) {
  #one loops through all anomalies
  # calculate probability of fstatistic because it cannot be extracted from lms above
  # see http://stats.stackexchange.com/questions/92824/extracting-the-model-p-value-for-a-multiple-regression-in-r
  # slope returned by lm is per second
  prob <-
    pf(lms[[i]]$fstatistic[1],
       lms[[i]]$fstatistic[2],
       lms[[i]]$fstatistic[3],
       lower.tail = FALSE)
  reg <-
    rbind(reg, as.numeric(
      c(
        lms[[i]]$coefficients[2, 1],
        lms[[i]]$coefficients[2, 2],
        lms[[i]]$coefficients[2, 4],
        lms[[i]]$coefficients[1, 1],
        lms[[i]]$coefficients[1, 2],
        lms[[i]]$coefficients[1, 4],
        lms[[i]]$fstatistic[1],
        lms[[i]]$fstatistic[3],
        lms[[i]]$r.squared,
        prob
      )
    ))
}

colnames(reg) <-
  c("Slope",
    "SE Slope",
    "P Slope",
    "Intercept",
    "SE int.",
    "P int.",
    "F",
    "df",
    "R2",
    "P value")
row.names(reg) <- var_list
reg_ano_0m_2010 <- reg

# Regression and table anomalies
var_list <-
  c(
    "Salinity_ano",
    "Temperature_ano")

lms <- lapply(var_list, function(x) {
  summary(lm(substitute(i ~ decimal_date(datetime), list(i = as.name(x))), data = ano_0m_2010))
})

reg <- NULL
for (i in 1:length(var_list)) {
  #one loops through all anomalies
  # calculate probability of fstatistic because it cannot be extracted from lms above
  # see http://stats.stackexchange.com/questions/92824/extracting-the-model-p-value-for-a-multiple-regression-in-r
  # slope returned by lm is per second
  prob <-
    pf(lms[[i]]$fstatistic[1],
       lms[[i]]$fstatistic[2],
       lms[[i]]$fstatistic[3],
       lower.tail = FALSE)
  reg <-
    rbind(reg, as.numeric(
      c(
        lms[[i]]$coefficients[2, 1],
        lms[[i]]$coefficients[2, 2],
        lms[[i]]$coefficients[2, 4],
        lms[[i]]$coefficients[1, 1],
        lms[[i]]$coefficients[1, 2],
        lms[[i]]$coefficients[1, 4],
        lms[[i]]$fstatistic[1],
        lms[[i]]$fstatistic[3],
        lms[[i]]$r.squared,
        prob
      )
    ))
}
colnames(reg) <-
  c("Slope",
    "SE Slope",
    "P Slope",
    "Intercept",
    "SE int.",
    "P int.",
    "F",
    "df",
    "R2",
    "P value")
row.names(reg) <- var_list
reg_ano_0m_2010 <- reg

reg_ano_0m_2010

```

#### Plot of anomalies : Temperature (2010-2022) - 0m

```{r plot anomalies temperature 2010-2022 - 0m, echo=FALSE, warning=FALSE, message=FALSE, fig.width=8,fig.height=4,fig.align='center'}

plot_ano_temp_2010_0m <- ggplot(data = ano_0m_2010, aes(x = datetime, y = Temperature_ano), na.rm=TRUE) +
  #scale_x_datetime(date_breaks="2 year", labels = date_format("%Y")) +
  geom_point(colour="blue", na.rm=TRUE, size=1) + 
  geom_smooth(method=lm, colour="black", fill="grey", linewidth=0.6, na.rm=TRUE) +
  labs(title="Temperature anomalies + trend (2010-2022)",x="", y="Temp. (°C)")

plot_temp_2010_0m <- Data_SOMLIT_2010_surf %>% 
  ggplot() +
  aes(x=datetime, y=impute_temp) + 
  geom_line(size=0.7, col = "#0F056B") +
  #scale_x_datetime(name="") +
  scale_y_continuous(name="Temp. (°C)") +
  ggtitle("Surface temperature time series (2010-2022)")

plot_temp_2010_0m
plot_ano_temp_2010_0m

```

#### Plot of anomalies : Salinity (2010-2022) - 0m

```{r plot anomalies salinity 2010-2022 - 0m, echo=FALSE, warning=FALSE, message=FALSE, fig.width=8,fig.height=4,fig.align='center'}

plot_ano_sal_2010_0m <- ggplot(data = ano_0m_2010, aes(x = datetime, y = Salinity_ano), na.rm=TRUE) +
  #scale_x_datetime(date_breaks="2 year", labels = date_format("%Y")) +
  geom_point(colour="blue", na.rm=TRUE, size=1) + 
  geom_smooth(method=lm, colour="black", fill="grey", linewidth=0.6, na.rm=TRUE) +
  labs(title="Salinity anomalies + trend (2010-2022)",x="", y="Salinity")

plot_sal_2010_0m <- Data_SOMLIT_2010_surf %>% 
  ggplot() +
  aes(x=datetime, y=impute_sal) + 
  geom_line(size=0.7, col = "#0F056B") +
  #scale_x_datetime(name="") +
  scale_y_continuous(name="Salinity") +
  ggtitle("Surface salinity time series (2010-2022)")

plot_sal_2010_0m
plot_ano_sal_2010_0m

```

## DEPTH 50

### Jan.2000 - Dec.2009

#### [Table of monthly means (2000-2009) - 50m data]{style="color:blue"}

```{r calculate monthly means 2000-2009 - 50m, echo=FALSE, warning=FALSE, message=FALSE}

Data_SOMLIT_2000_prof <- TS_9522_prof %>% 
  dplyr::filter(datetime >= "2000-01-05" & datetime <= "2009-12-22")
##

monthly_means_50m_2000 <- Data_SOMLIT_2000_prof %>% 
  group_by(month) %>%
  summarise(Salinity_month = mean(impute_sal),
            SD_sal = sd(impute_sal), 
            Temperature_month = mean(impute_temp),
            SD_temp = sd(impute_temp))

monthly_means_50m_2000

```

#### [Table of anomaly regression analyses (2000-2009) - 50m data]{style="color:green"}

```{r calculate anomalies + regression (2000-2009) - 50m, echo=FALSE, warning=FALSE, message=FALSE}

ano_50m_2000 <- left_join(ungroup(Data_SOMLIT_2000_prof), monthly_means_50m_2000, by = "month") %>%
  mutate(
    Salinity_ano = impute_sal - Salinity_month,
    Temperature_ano = impute_temp - Temperature_month)

# Regressions and tables of key parameters
var_list <-
  c("salinity",
    "temperature")

lm.test <- vector("list", length(var_list))

for(i in seq_along(var_list)){
  lm.test[[i]] <- lm(reformulate(var_list[i], "datetime"), data = ano_50m_2000)
}

#cutoff_date <- "2016-01-01" #to limit regressions on complete years

lms <- lapply(var_list, function(x) {
  summary(lm(substitute(i ~ decimal_date(datetime), list(i = as.name(x))), 
             data = ano_50m_2000))
})

reg <- NULL
for (i in 1:length(var_list)) {
  #one loops through all anomalies
  # calculate probability of fstatistic because it cannot be extracted from lms above
  # see http://stats.stackexchange.com/questions/92824/extracting-the-model-p-value-for-a-multiple-regression-in-r
  # slope returned by lm is per second
  prob <-
    pf(lms[[i]]$fstatistic[1],
       lms[[i]]$fstatistic[2],
       lms[[i]]$fstatistic[3],
       lower.tail = FALSE)
  reg <-
    rbind(reg, as.numeric(
      c(
        lms[[i]]$coefficients[2, 1],
        lms[[i]]$coefficients[2, 2],
        lms[[i]]$coefficients[2, 4],
        lms[[i]]$coefficients[1, 1],
        lms[[i]]$coefficients[1, 2],
        lms[[i]]$coefficients[1, 4],
        lms[[i]]$fstatistic[1],
        lms[[i]]$fstatistic[3],
        lms[[i]]$r.squared,
        prob
      )
    ))
}

colnames(reg) <-
  c("Slope",
    "SE Slope",
    "P Slope",
    "Intercept",
    "SE int.",
    "P int.",
    "F",
    "df",
    "R2",
    "P value")
row.names(reg) <- var_list
reg_ano_50m_2000 <- reg

# Regression and table anomalies
var_list <-
  c(
    "Salinity_ano",
    "Temperature_ano")

lms <- lapply(var_list, function(x) {
  summary(lm(substitute(i ~ decimal_date(datetime), list(i = as.name(x))), data = ano_50m_2000))
})

reg <- NULL
for (i in 1:length(var_list)) {
  #one loops through all anomalies
  # calculate probability of fstatistic because it cannot be extracted from lms above
  # see http://stats.stackexchange.com/questions/92824/extracting-the-model-p-value-for-a-multiple-regression-in-r
  # slope returned by lm is per second
  prob <-
    pf(lms[[i]]$fstatistic[1],
       lms[[i]]$fstatistic[2],
       lms[[i]]$fstatistic[3],
       lower.tail = FALSE)
  reg <-
    rbind(reg, as.numeric(
      c(
        lms[[i]]$coefficients[2, 1],
        lms[[i]]$coefficients[2, 2],
        lms[[i]]$coefficients[2, 4],
        lms[[i]]$coefficients[1, 1],
        lms[[i]]$coefficients[1, 2],
        lms[[i]]$coefficients[1, 4],
        lms[[i]]$fstatistic[1],
        lms[[i]]$fstatistic[3],
        lms[[i]]$r.squared,
        prob
      )
    ))
}
colnames(reg) <-
  c("Slope",
    "SE Slope",
    "P Slope",
    "Intercept",
    "SE int.",
    "P int.",
    "F",
    "df",
    "R2",
    "P value")
row.names(reg) <- var_list
reg_ano_50m_2000 <- reg

reg_ano_50m_2000

```

#### Plot of anomalies : Temperature (2000-2009) - 50m

```{r plot anomalies temperature 2000-2009 - 50m, echo=FALSE, warning=FALSE, message=FALSE, fig.width=8,fig.height=4,fig.align='center'}

plot_ano_temp_2000_50m <- ggplot(data = ano_50m_2000, aes(x = datetime, y = Temperature_ano), na.rm=TRUE) +
  #scale_x_datetime(date_breaks="2 year", labels = date_format("%Y")) +
  geom_point(colour="blue", na.rm=TRUE, size=1) + 
  geom_smooth(method=lm, colour="black", fill="grey", linewidth=0.6, na.rm=TRUE) +
  labs(title="Temperature anomalies + trend (2000-2009)",x="", y="Temp. (°C)")

plot_temp_2000_50m <- Data_SOMLIT_2000_prof %>% 
  ggplot() +
  aes(x=datetime, y=impute_temp) + 
  geom_line(size=0.7, col = "#24445C") +
  #scale_x_datetime(name="") +
  scale_y_continuous(name="Temp. (°C)") +
  ggtitle("Surface temperature time series (2000-2009)")

plot_temp_2000_50m
plot_ano_temp_2000_50m

```

#### Plot of anomalies : Salinity (2000-2009) - 50m

```{r plot anomalies salinity 2000-2009 - 50m, echo=FALSE, warning=FALSE, message=FALSE, fig.width=8,fig.height=4,fig.align='center'}

plot_ano_sal_2000_50m <- ggplot(data = ano_50m_2000, aes(x = datetime, y = Salinity_ano), na.rm=TRUE) +
  #scale_x_datetime(date_breaks="2 year", labels = date_format("%Y")) +
  geom_point(colour="blue", na.rm=TRUE, size=1) + 
  geom_smooth(method=lm, colour="black", fill="grey", linewidth=0.6, na.rm=TRUE) +
  labs(title="Salinity anomalies + trend (2000-2009)",x="", y="Salinity")

plot_sal_2000_50m <- Data_SOMLIT_2000_prof %>% 
  ggplot() +
  aes(x=datetime, y=impute_sal) + 
  geom_line(size=0.7, col = "#24445C") +
  #scale_x_datetime(name="") +
  scale_y_continuous(name="Salinity") +
  ggtitle("Surface salinity time series (2000-2009)")

plot_sal_2000_50m
plot_ano_sal_2000_50m

```

### Jan.2010 - Dec.2022

#### [Table of monthly means (2010-2022) - 50m data]{style="color:blue"}

```{r calculate monthly means 2010-2022 - 50m, echo=FALSE, warning=FALSE, message=FALSE}

Data_SOMLIT_2010_prof <- TS_9522_prof %>% 
  dplyr::filter(datetime >= "2010-01-04")
##

monthly_means_50m_2010 <- Data_SOMLIT_2010_prof %>% 
  group_by(month) %>%
  summarise(Salinity_month = mean(impute_sal),
            SD_sal = sd(impute_sal),
            Temperature_month = mean(impute_temp),
            SD_temp = sd(impute_temp))

monthly_means_50m_2010

```

#### [Table of anomaly regression analyses (2010-2022) - 50m data]{style="color:green"}

```{r calculate anomalies + regression (2010-2022) - 50m, echo=FALSE, warning=FALSE, message=FALSE}

ano_50m_2010 <- left_join(ungroup(Data_SOMLIT_2010_prof), monthly_means_50m_2010, by = "month") %>%
  mutate(
    Salinity_ano = impute_sal - Salinity_month,
    Temperature_ano = impute_temp - Temperature_month)

# Regressions and tables of key parameters
var_list <-
  c("salinity",
    "temperature")

lm.test <- vector("list", length(var_list))

for(i in seq_along(var_list)){
  lm.test[[i]] <- lm(reformulate(var_list[i], "datetime"), data = ano_50m_2010)
}

#cutoff_date <- "2016-01-01" #to limit regressions on complete years

lms <- lapply(var_list, function(x) {
  summary(lm(substitute(i ~ decimal_date(datetime), list(i = as.name(x))), 
             data = ano_50m_2010))
})

reg <- NULL
for (i in 1:length(var_list)) {
  #one loops through all anomalies
  # calculate probability of fstatistic because it cannot be extracted from lms above
  # see http://stats.stackexchange.com/questions/92824/extracting-the-model-p-value-for-a-multiple-regression-in-r
  # slope returned by lm is per second
  prob <-
    pf(lms[[i]]$fstatistic[1],
       lms[[i]]$fstatistic[2],
       lms[[i]]$fstatistic[3],
       lower.tail = FALSE)
  reg <-
    rbind(reg, as.numeric(
      c(
        lms[[i]]$coefficients[2, 1],
        lms[[i]]$coefficients[2, 2],
        lms[[i]]$coefficients[2, 4],
        lms[[i]]$coefficients[1, 1],
        lms[[i]]$coefficients[1, 2],
        lms[[i]]$coefficients[1, 4],
        lms[[i]]$fstatistic[1],
        lms[[i]]$fstatistic[3],
        lms[[i]]$r.squared,
        prob
      )
    ))
}

colnames(reg) <-
  c("Slope",
    "SE Slope",
    "P Slope",
    "Intercept",
    "SE int.",
    "P int.",
    "F",
    "df",
    "R2",
    "P value")
row.names(reg) <- var_list
reg_ano_50m_2010 <- reg

# Regression and table anomalies
var_list <-
  c(
    "Salinity_ano",
    "Temperature_ano")

lms <- lapply(var_list, function(x) {
  summary(lm(substitute(i ~ decimal_date(datetime), list(i = as.name(x))), data = ano_50m_2010))
})

reg <- NULL
for (i in 1:length(var_list)) {
  #one loops through all anomalies
  # calculate probability of fstatistic because it cannot be extracted from lms above
  # see http://stats.stackexchange.com/questions/92824/extracting-the-model-p-value-for-a-multiple-regression-in-r
  # slope returned by lm is per second
  prob <-
    pf(lms[[i]]$fstatistic[1],
       lms[[i]]$fstatistic[2],
       lms[[i]]$fstatistic[3],
       lower.tail = FALSE)
  reg <-
    rbind(reg, as.numeric(
      c(
        lms[[i]]$coefficients[2, 1],
        lms[[i]]$coefficients[2, 2],
        lms[[i]]$coefficients[2, 4],
        lms[[i]]$coefficients[1, 1],
        lms[[i]]$coefficients[1, 2],
        lms[[i]]$coefficients[1, 4],
        lms[[i]]$fstatistic[1],
        lms[[i]]$fstatistic[3],
        lms[[i]]$r.squared,
        prob
      )
    ))
}
colnames(reg) <-
  c("Slope",
    "SE Slope",
    "P Slope",
    "Intercept",
    "SE int.",
    "P int.",
    "F",
    "df",
    "R2",
    "P value")
row.names(reg) <- var_list
reg_ano_50m_2010 <- reg

reg_ano_50m_2010

```

#### Plot of anomalies : Temperature (2010-2022) - 50m

```{r plot anomalies temperature 2010-2022 - 50m, echo=FALSE, warning=FALSE, message=FALSE, fig.width=8,fig.height=4,fig.align='center'}

plot_ano_temp_2010_50m <- ggplot(data = ano_50m_2010, aes(x = datetime, y = Temperature_ano), na.rm=TRUE) +
  #scale_x_datetime(date_breaks="2 year", labels = date_format("%Y")) +
  geom_point(colour="blue", na.rm=TRUE, size=1) + 
  geom_smooth(method=lm, colour="black", fill="grey", linewidth=0.6, na.rm=TRUE) +
  labs(title="Temperature anomalies + trend (2010-2022)",x="", y="Temp. (°C)")

plot_temp_2010_50m <- Data_SOMLIT_2010_prof %>% 
  ggplot() +
  aes(x=datetime, y=impute_temp) + 
  geom_line(size=0.7, col = "#0F056B") +
  #scale_x_datetime(name="") +
  scale_y_continuous(name="Temp. (°C)") +
  ggtitle("Surface temperature time series (2010-2022)")

plot_temp_2010_50m
plot_ano_temp_2010_50m

```

#### Plot of anomalies : Salinity (2010-2022) - 50m

```{r plot anomalies salinity 2010-2022 - 50m, echo=FALSE, warning=FALSE, message=FALSE, fig.width=8,fig.height=4,fig.align='center'}

plot_ano_sal_2010_50m <- ggplot(data = ano_50m_2010, aes(x = datetime, y = Salinity_ano), na.rm=TRUE) +
  #scale_x_datetime(date_breaks="2 year", labels = date_format("%Y")) +
  geom_point(colour="blue", na.rm=TRUE, size=1) + 
  geom_smooth(method=lm, colour="black", fill="grey", linewidth=0.6, na.rm=TRUE) +
  labs(title="Salinity anomalies + trend (2010-2022)",x="", y="Salinity")

plot_sal_2010_50m <- Data_SOMLIT_2010_prof %>% 
  ggplot() +
  aes(x=datetime, y=impute_sal) + 
  geom_line(size=0.7, col = "#0F056B") +
  #scale_x_datetime(name="") +
  scale_y_continuous(name="Salinity") +
  ggtitle("Surface salinity time series (2010-2022)")

plot_sal_2010_50m
plot_ano_sal_2010_50m

```

------------------------------------------------------------------------

# Comparison SOMLIT - EOL results : 2018-2022

## SURFACE

temperature (°C) & salinity

### [Table and plot of monthly means (2017-2022) - EOL temperature & salinity]{style="color:blue"}

```{r importation Data EOL hourly + creation SOMLIT 18-22, echo=FALSE, warning=FALSE, message=FALSE}

#period : Jan.2018 - Dec.2022

#EOL :
EOL_hourly_1822 <- read_delim("../../Data/data_EOL/EOL_18-23_hourly.csv", 
    delim = ";", escape_double = FALSE, trim_ws = TRUE) %>% 
  rename(temperature=temp_eol_c, salinity=sal_eol_psu) %>% 
  mutate(datetime = as.POSIXct(datetime, format = "%Y-%m-%d")) %>% 
  dplyr::filter(datetime <= "2022-12-31 01:00:00")

#SOMLIT :
SOMLIT_1822 <- data_0722 %>% 
  dplyr::filter(datetime >= "2018-01-03") %>% 
  mutate(Month = format(datetime, format="%m"))

```

```{r calcul monthly means - temperature & salinity - EOL - 2018-2022, echo=TRUE, warning=FALSE, message=FALSE}

monthly_means_EOL_1822 <- ungroup(EOL_hourly_1822) %>% 
  group_by(Month) %>%
  summarise(
    Salinity_month = mean(salinity, na.rm = TRUE),
    SD_sal = sd(salinity, na.rm = TRUE),
    Temperature_month = mean(temperature, na.rm = TRUE),
    SD_temp = sd(temperature, na.rm = TRUE))

monthly_means_EOL_1822

```

```{r calcul monthly means - temperature & salinity - SOMLIT - 2018-2022, echo=TRUE, warning=FALSE, message=FALSE}

monthly_means_somlit_1822 <- ungroup(SOMLIT_1822) %>%
  group_by(Month) %>%
  summarise(
    Salinity_month = mean(impute_sal, na.rm = TRUE),
    SD_sal = sd(impute_sal, na.rm = TRUE),
    Temperature_month = mean(impute_temp, na.rm = TRUE),
    SD_temp = sd(impute_temp, na.rm = TRUE))

monthly_means_somlit_1822

```

### [Table of anomalies regression analyses (2018-2022) - EOL temperature & salinity]{style="color:green"}

```{r regression analysis - temperature & salinity - EOL - 2018-2022, echo=FALSE, warning=FALSE, message=FALSE}

anomalies_EOL_1822 <- left_join(ungroup(EOL_hourly_1822), monthly_means_EOL_1822, by = "Month") %>%
  mutate(Salinity_ano = salinity - Salinity_month,
         Temperature_ano = temperature - Temperature_month) %>% 
  unite(col=datetime_fusion, "datetime", "Hour", sep=" ") %>% 
  mutate(datetime_fusion = as.POSIXct(datetime_fusion, format = "%Y-%m-%d %H:%M:%S"))

# Regressions and tables of key parameters
var_list <-
  c("salinity",
    "temperature")

lm.test <- vector("list", length(var_list))
 
 for(i in seq_along(var_list)){
     lm.test[[i]] <- lm(reformulate(var_list[i], "datetime_fusion"), data = anomalies_EOL_1822)
 }


lms <- lapply(var_list, function(x) {
  summary(lm(substitute(i ~ decimal_date(datetime_fusion), list(i = as.name(x))), 
             data = anomalies_EOL_1822))
})

reg <- NULL
for (i in 1:length(var_list)) {
  #one loops through all anomalies
  # calculate probability of fstatistic because it cannot be extracted from lms above
  # see http://stats.stackexchange.com/questions/92824/extracting-the-model-p-value-for-a-multiple-regression-in-r
  # slope returned by lm is per second
  prob <-
    pf(lms[[i]]$fstatistic[1],
       lms[[i]]$fstatistic[2],
       lms[[i]]$fstatistic[3],
       lower.tail = FALSE)
  reg <-
    rbind(reg, as.numeric(
      c(
        lms[[i]]$coefficients[2, 1],
        lms[[i]]$coefficients[2, 2],
        lms[[i]]$coefficients[2, 4],
        lms[[i]]$coefficients[1, 1],
        lms[[i]]$coefficients[1, 2],
        lms[[i]]$coefficients[1, 4],
        lms[[i]]$fstatistic[1],
        lms[[i]]$fstatistic[3],
        lms[[i]]$r.squared,
        prob
      )
    ))
}

colnames(reg) <-
  c("Slope",
    "SE Slope",
    "P Slope",
    "Intercept",
    "SE int.",
    "P int.",
    "F",
    "df",
    "R2",
    "P value")
row.names(reg) <- var_list
reg_ano_EOL_1822_surf <- reg

# Regression and table anomalies
var_list <-
  c(
    "Salinity_ano",
    "Temperature_ano")

lms <- lapply(var_list, function(x) {
  summary(lm(substitute(i ~ decimal_date(datetime_fusion), list(i = as.name(x))), 
             data = anomalies_EOL_1822))
})
reg <- NULL
for (i in 1:length(var_list)) {
  #one loops through all anomalies
  # calculate probability of fstatistic because it cannot be extracted from lms above
  # see http://stats.stackexchange.com/questions/92824/extracting-the-model-p-value-for-a-multiple-regression-in-r
  # slope returned by lm is per second
  prob <-
    pf(lms[[i]]$fstatistic[1],
       lms[[i]]$fstatistic[2],
       lms[[i]]$fstatistic[3],
       lower.tail = FALSE)
  reg <-
    rbind(reg, as.numeric(
      c(
        lms[[i]]$coefficients[2, 1],
        lms[[i]]$coefficients[2, 2],
        lms[[i]]$coefficients[2, 4],
        lms[[i]]$coefficients[1, 1],
        lms[[i]]$coefficients[1, 2],
        lms[[i]]$coefficients[1, 4],
        lms[[i]]$fstatistic[1],
        lms[[i]]$fstatistic[3],
        lms[[i]]$r.squared,
        prob
      )
    ))
}
colnames(reg) <-
  c("Slope",
    "SE Slope",
    "P Slope",
    "Intercept",
    "SE int.",
    "P int.",
    "F",
    "df",
    "R2",
    "P value")

row.names(reg) <- var_list
reg_ano_EOL_1822_surf <- reg

reg_ano_EOL_1822_surf
```

### [Table of anomalies regression analyses (2018-2022) - SOMLIT temperature & salinity]{style="color:green"}

```{r regression analysis - temperature & salinity - SOMLIT - 2018-2022, echo=FALSE, warning=FALSE, message=FALSE}

anomalies_SOM_1822 <- left_join(ungroup(SOMLIT_1822), monthly_means_somlit_1822, by = "Month") %>% 
  mutate(Salinity_ano = impute_sal - Salinity_month,
         Temperature_ano = impute_temp - Temperature_month)

# Regressions and tables of key parameters
var_list <-
  c("salinity",
    "temperature")

lm.test <- vector("list", length(var_list))
 
 for(i in seq_along(var_list)){
     lm.test[[i]] <- lm(reformulate(var_list[i], "datetime"), data = anomalies_SOM_1822)
 }


lms <- lapply(var_list, function(x) {
  summary(lm(substitute(i ~ decimal_date(datetime), list(i = as.name(x))), 
             data = anomalies_SOM_1822))
})

reg <- NULL
for (i in 1:length(var_list)) {
  #one loops through all anomalies
  # calculate probability of fstatistic because it cannot be extracted from lms above
  # see http://stats.stackexchange.com/questions/92824/extracting-the-model-p-value-for-a-multiple-regression-in-r
  # slope returned by lm is per second
  prob <-
    pf(lms[[i]]$fstatistic[1],
       lms[[i]]$fstatistic[2],
       lms[[i]]$fstatistic[3],
       lower.tail = FALSE)
  reg <-
    rbind(reg, as.numeric(
      c(
        lms[[i]]$coefficients[2, 1],
        lms[[i]]$coefficients[2, 2],
        lms[[i]]$coefficients[2, 4],
        lms[[i]]$coefficients[1, 1],
        lms[[i]]$coefficients[1, 2],
        lms[[i]]$coefficients[1, 4],
        lms[[i]]$fstatistic[1],
        lms[[i]]$fstatistic[3],
        lms[[i]]$r.squared,
        prob
      )
    ))
}

colnames(reg) <-
  c("Slope",
    "SE Slope",
    "P Slope",
    "Intercept",
    "SE int.",
    "P int.",
    "F",
    "df",
    "R2",
    "P value")
row.names(reg) <- var_list
reg_ano_SOM_1822 <- reg

# Regression and table anomalies
var_list <-
  c(
    "Salinity_ano",
    "Temperature_ano")

lms <- lapply(var_list, function(x) {
  summary(lm(substitute(i ~ decimal_date(datetime), list(i = as.name(x))), 
             data = anomalies_SOM_1822))
})
reg <- NULL
for (i in 1:length(var_list)) {
  #one loops through all anomalies
  # calculate probability of fstatistic because it cannot be extracted from lms above
  # see http://stats.stackexchange.com/questions/92824/extracting-the-model-p-value-for-a-multiple-regression-in-r
  # slope returned by lm is per second
  prob <-
    pf(lms[[i]]$fstatistic[1],
       lms[[i]]$fstatistic[2],
       lms[[i]]$fstatistic[3],
       lower.tail = FALSE)
  reg <-
    rbind(reg, as.numeric(
      c(
        lms[[i]]$coefficients[2, 1],
        lms[[i]]$coefficients[2, 2],
        lms[[i]]$coefficients[2, 4],
        lms[[i]]$coefficients[1, 1],
        lms[[i]]$coefficients[1, 2],
        lms[[i]]$coefficients[1, 4],
        lms[[i]]$fstatistic[1],
        lms[[i]]$fstatistic[3],
        lms[[i]]$r.squared,
        prob
      )
    ))
}
colnames(reg) <-
  c("Slope",
    "SE Slope",
    "P Slope",
    "Intercept",
    "SE int.",
    "P int.",
    "F",
    "df",
    "R2",
    "P value")

row.names(reg) <- var_list
reg_ano_SOM_1822 <- reg

reg_ano_SOM_1822
```

### EOL : Plot of anomalies : Temperature (°C) (2018-2022)

```{r plot anomalies + trends - temperature - EOL 1822, echo=FALSE, warning=FALSE, message=FALSE}


plot_ano_temp_EOL_1822 <- ggplot(data = anomalies_EOL_1822, aes(x = datetime_fusion, y = Temperature_ano), na.rm=TRUE) +
  scale_x_datetime(date_breaks="1 year", labels = date_format("%Y")) +
  geom_point(colour="blue", na.rm=TRUE, size=0.65) + 
  geom_smooth(method=lm, colour="black", fill="grey", linewidth=0.6, na.rm=TRUE) +
  labs(title="Temperature anomalies + trend (2018-2022)",x="", y="Temp. (°C)") 

plot_temperature_EOL_1822 <- EOL_hourly_1822 %>% 
  ggplot() +
  aes(x=datetime, y=temperature) + 
  geom_line(size=0.7) +
  scale_x_datetime(name="") +
  scale_y_continuous(name="Temp. (°C)") +
  ggtitle("EOL time series observations of temperature (°C) 2018-2022")

plot_temperature_EOL_1822
plot_ano_temp_EOL_1822
```

### EOL : Plot of anomalies : Salinity (2018-2022)

*without interpolation*

```{r plot anomalies + trends - salinity - EOL 1822, echo=FALSE, warning=FALSE, message=FALSE}


plot_ano_sal_EOL_1822 <- ggplot(data = anomalies_EOL_1822, aes(x = datetime_fusion, y = Salinity_ano), na.rm=TRUE) +
  scale_x_datetime(date_breaks="1 year", date_minor_breaks="1 years", labels = date_format("%Y")) +
  geom_point(colour="blue", na.rm=TRUE, size=0.65) + 
  geom_smooth(method=lm, colour="black", fill="grey", linewidth=0.6, na.rm=TRUE) +
  labs(title="Salinity anomalies + trend (2018-2022)",x="", y="Salinity")

plot_salinity_EOL_1822 <- EOL_hourly_1822 %>% 
  ggplot() +
  aes(x=datetime, y=salinity) + 
  geom_point(size=1) +
  scale_x_datetime(name="") +
  scale_y_continuous(name="Salinity") +
  ggtitle("EOL time series observations of salinity  2018-2022")

plot_salinity_EOL_1822
plot_ano_sal_EOL_1822
```

### SOMLIT : Plot of anomalies : Temperature (°C) (2018-2022)

```{r plot anomalies + trends - temperature - SOMLIT 1822, echo=FALSE, warning=FALSE, message=FALSE}


plot_ano_temp_SOM_1822 <- ggplot(data = anomalies_SOM_1822, aes(x = datetime, y = Temperature_ano), 
                                 na.rm=TRUE) +
  scale_x_date(date_breaks="1 year", date_minor_breaks="1 years", labels = date_format("%Y")) +
  geom_point(colour="blue", na.rm=TRUE, size=0.65) + 
  geom_smooth(method=lm, colour="black", fill="grey", linewidth=0.6, na.rm=TRUE) +
  labs(title="Temperature anomalies + trend (2018-2022)",x="", y="Temp. (°C)")

plot_temperature_SOM_1822 <- SOMLIT_1822 %>% 
  ggplot() +
  aes(x=datetime, y=temperature) + 
  geom_point(size=1) +
  scale_x_date(name="") +
  scale_y_continuous(name="Temp. (°C)") +
  ggtitle("SOMLIT time series observations of temperature (°C) 2018-2022")

plot_temperature_SOM_1822
plot_ano_temp_SOM_1822
```

### SOMLIT : Plot of anomalies : Salinity (2018-2022)

```{r plot anomalies + trends - salinity - SOMLIT 1822, echo=FALSE, warning=FALSE, message=FALSE}


plot_ano_sal_SOM_1822 <- ggplot(data = anomalies_SOM_1822, aes(x = datetime, y = Salinity_ano), na.rm=TRUE) +
  scale_x_date(date_breaks="1 year", date_minor_breaks="1 years", labels = date_format("%Y")) +
  geom_point(colour="blue", na.rm=TRUE, size=0.65) + 
  geom_smooth(method=lm, colour="black", fill="grey", linewidth=0.6, na.rm=TRUE) +
  labs(title="Salinity anomalies + trend (2018-2022)",x="", y="Salinity")

plot_salinity_SOM_1822 <- SOMLIT_1822 %>% 
  ggplot() +
  aes(x=datetime, y=salinity) + 
  geom_point(size=1) +
  scale_x_date(name="") +
  scale_y_continuous(name="Salinity") +
  ggtitle("SOMLIT time series observations of salinity  2018-2022")

plot_salinity_SOM_1822
plot_ano_sal_SOM_1822
```
